---
title: "Tracking the Acquisition of Differential Object Marking in Children and Adults"
subtitle : "10th National Symposium of Spanish as a Heritage Language (Harvard University)"
author: "Patrick D. Thane"
institute: "Rutgers University / University of Massachusetts"
date: "April 14, 2023"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["rutgers", "rutgers-fonts"]
    nature:
      beforeInit: "http://www.jvcasillas.com/ru_xaringan/js/ru_xaringan.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, echo = FALSE, include = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 99)

library(tidyverse)
library(lme4)
library(emmeans)
library(knitr)
library(kableExtra)
library(xaringan)
library(patchwork)
library(jtools)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(here)
```

```{r, load-datasets}
# EPT
SDB_EPT <- read_csv(here("./CSV Files/SDB/SDB DOM EPT.csv"))

HSA_EPT <- read.csv(here("./CSV Files/Adult HS/Adult HS DOM EPT.csv")) %>% 
  mutate(Group = "Adult HS")

DLI78_EPT <- read.csv(here("./CSV Files/DLI-78/DLI-78 DOM EPT.csv"))
MLS78_EPT <- read.csv(here("./CSV Files/MLS-78/MLS-78 DOM EPT.csv"))

DLI5_EPT <- read.csv(here("./CSV Files/DLI-5/DLI-5 DOM EPT.csv"))
MLS5_EPT <- read.csv(here("./CSV Files/MLS-5/MLS-5 DOM EPT.csv"))


# FCT
SDB_FCT <- read_csv(here("./CSV Files/SDB/SDB DOM FCT.csv"))

HSA_FCT <- read.csv(here("./CSV Files/Adult HS/Adult HS DOM FCT.csv")) %>% 
  mutate(Group = "Adult HS")

DLI78_FCT <- read.csv(here("./CSV Files/DLI-78/DLI-78 DOM FCT.csv"))
MLS78_FCT <- read.csv(here("./CSV Files/MLS-78/MLS-78 DOM FCT.csv"))

DLI5_FCT <- read.csv(here("./CSV Files/DLI-5/DLI-5 DOM FCT.csv"))
MLS5_FCT <- read.csv(here("./CSV Files/MLS-5/MLS-5 DOM FCT.csv"))


# Join datasets by age group
Older_EPT <- rbind(DLI78_EPT, MLS78_EPT) %>% 
  mutate(Group = "7th/8th Grade HS")

Fifth_EPT <- rbind(DLI5_EPT, MLS5_EPT) %>% 
  mutate(Group = "5th Grade HS")

Older_FCT <- rbind(DLI78_FCT, MLS78_FCT) %>% 
  mutate(Group = "7th/8th Grade HS")

Fifth_FCT <- rbind(DLI5_FCT, MLS5_FCT) %>% 
  mutate(Group = "5th Grade HS")


# Bind datasets
EPT <- rbind(SDB_EPT, HSA_EPT, Fifth_EPT, Older_EPT)
EPT$Group <- factor(EPT$Group, levels = c("SDB", "Adult HS", "7th/8th Grade HS", "5th Grade HS"))

FCT <- rbind(SDB_FCT, HSA_FCT, Fifth_FCT, Older_FCT)
FCT$Group <- factor(FCT$Group, levels = c("SDB", "Adult HS", "7th/8th Grade HS", "5th Grade HS"))

Aggregate_HS <- rbind(HSA_EPT, Fifth_EPT, Older_EPT, HSA_FCT, Fifth_FCT, Older_FCT)
Aggregate_HS$Group <- factor(Aggregate_HS$Group, levels = c("HS Adults", "7th/8th Grade HS", "5th Grade HS"))
```

```{r, generate-participant-averages}
SDB_EPT_Structure <- aggregate(SDB_EPT$DOM_Use, list(SDB_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_EPT_Structure <- SDB_EPT_Structure %>% rename(Structure_Avg = x)
SDB_EPT_Structure <- left_join(SDB_EPT, SDB_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

SDB_FCT_Structure <- aggregate(SDB_FCT$DOM_Use, list(SDB_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_FCT_Structure <- SDB_FCT_Structure %>% rename(Structure_Avg = x)
SDB_FCT_Structure <- left_join(SDB_FCT, SDB_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HSA_EPT_Structure <- aggregate(HSA_EPT$DOM_Use, list(HSA_EPT$Structure), FUN = mean, na.rm = TRUE)
HSA_EPT_Structure <- HSA_EPT_Structure %>% rename(Structure_Avg = x)
HSA_EPT_Structure <- left_join(HSA_EPT, HSA_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HSA_FCT_Structure <- aggregate(HSA_FCT$DOM_Use, list(HSA_FCT$Structure), FUN = mean, na.rm = TRUE)
HSA_FCT_Structure <- HSA_FCT_Structure %>% rename(Structure_Avg = x)
HSA_FCT_Structure <- left_join(HSA_FCT, HSA_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

Older_EPT_Structure <- aggregate(Older_EPT$DOM_Use, list(Older_EPT$Structure), FUN = mean, na.rm = TRUE)
Older_EPT_Structure <- Older_EPT_Structure %>% rename(Structure_Avg = x)
Older_EPT_Structure <- left_join(Older_EPT, Older_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

Older_FCT_Structure <- aggregate(Older_FCT$DOM_Use, list(Older_FCT$Structure), FUN = mean, na.rm = TRUE)
Older_FCT_Structure <- Older_FCT_Structure %>% rename(Structure_Avg = x)
Older_FCT_Structure <- left_join(Older_FCT, Older_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

Fifth_EPT_Structure <- aggregate(Fifth_EPT$DOM_Use, list(Fifth_EPT$Structure), FUN = mean, na.rm = TRUE)
Fifth_EPT_Structure <- Fifth_EPT_Structure %>% rename(Structure_Avg = x)
Fifth_EPT_Structure <- left_join(Fifth_EPT, Fifth_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

Fifth_FCT_Structure <- aggregate(Fifth_FCT$DOM_Use, list(Fifth_FCT$Structure), FUN = mean, na.rm = TRUE)
Fifth_FCT_Structure <- Fifth_FCT_Structure %>% rename(Structure_Avg = x)
Fifth_FCT_Structure <- left_join(Fifth_FCT, Fifth_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

EPT_Structure <- rbind(SDB_EPT_Structure, HSA_EPT_Structure, Older_EPT_Structure, Fifth_EPT_Structure)
FCT_Structure <- rbind(SDB_FCT_Structure, HSA_FCT_Structure, Older_FCT_Structure, Fifth_FCT_Structure)

Aggregate <- rbind(EPT_Structure, FCT_Structure)
Aggregate$Group <- factor(Aggregate$Group, levels = c("SDB", "Adult HS", "7th/8th Grade HS", "5th Grade HS"))
```

```{r, generate-HS-data}
Aggregate_HS <- Aggregate %>% 
  filter(Group == c("Adult HS", "7th/8th Grade HS", "5th Grade HS"))
```

# Statement on Labor Negotiations at Rutgers


--
I stand in solidarity with my graduate student colleagues, as well as members of the AAUP unions that represent adjunct, teaching, tenured, and postdoctoral faculty. We also are representing members of Rutgers' administrative and custodial staff in an attempt to gain more favorable funding, stable healthcare, and job permanence. We are also fighting to obtain rent stability for undergraduate and graduate students, not only for the benefit of students but also to prevent gentrification that has already begun to displace members of the surrounding communities, a disproportionate amount of whom are Hispanic.

The union has not explicitly dissuaded graduate students from pursuing scholarly activity, including conference travel. It behooves those of us who are active in pursuing more equitable educational opportunities for Spanish heritage speakers to continue scholarly activity related to these topics. I dedicate today's presentation to the New Brunswick community, some of whom participated in this study, in our unified efforts to create a more inclusive Rutgers and surrounding region with a vision for more accessible public university education.

The outcomes of the strike at Rutgers, which began on April 10, will have minor economic impact on my salary, as I have accepted a postdoctoral position at the University of Massachusetts for the coming academic year. However, I am striking primarily in support of our joint unions' Bargaining for the Common Good initiatives (https://smlr.rutgers.edu/faculty-research-engagement/center-innovation-worker-organization-ciwo/bargaining-common-good).


---
# Introduction

- Differential object marking (DOM) is common to many Romance/Semitic languages (Bossong, 1992)


--
- Spanish is a language that marks direct objects with *a* when they are (Torrego, 1998):


--
  + Animate (this can include animals)


--
  + Specific


--
- DOM is an essential structure to maintain freedom of word order in Spanish:


--
  + **Ve Juana** (*Juan sees*, verb-subject)


--
  + **Ve *a* Juana** (*null subject + sees Juan*, subject-verb-object)


--
- There are additional considerations in the marking of animacy, such as the humanness of the object (Callen & Miller, 2021)


--
  + In this study, proper nouns were used as DOs


--
  + Avoids variability along animacy scale
  
  
--
  + Animacy scale (Aissen, 2013): animate, human objects are most likely to be marked


---
# DOM in Heritage Spanish


--
- Research in adults: Frequently omitted in expected contexts (Guijarro-Fuentes et al., 2015; Montrul, 2004; Montrul & Bowles, 2009; Montrul et al., 2015)


--
  + Variability in production is proficiency-dependent

--
  + Overacceptance of DOM omission occurs regardless of proficiency (Montrul & Bowles, 2009)


--
  + No effect of sequential versus simultaneous bilingualism (Montrul et al., 2015)


--
- DOM continues to develop over time, as children's production increases into adulthood (Montrul & Sánchez-Walker, 2013)


--
- HS are sensitive to the lexical frequency of verbs, especially at intermediate proficiency levels (Hur, 2020)


--
- Extreme degree of variability, ranging from 0% to 100% production in expected contexts


---
# Age Effects in HL Acquisition


- Age of onset of the L2 has received more consideration than developmental age


--
  + Age of onset represents an approximation of HL exposure prior to crosslinguistic influence


--
  + Age at time of testing allows us to trace the protracted developmental path of bilingual children


--
- There is evidence that HS children's command of inflectional morphology increases with age


--
  + Differential object marking (Montrul & Sánchez-Walker, 2013), children vs. adults


--
  + Preterit/imperfect contrast (Corbet & Domínguez, 2020; Cuza & Miller, 2015), young vs. older children


--
  + Gender agreement and assignment (Montrul & Potowski, 2007), young vs. older children


---
# Production-Comprehension Asymmetries


--
- Research on adult HS has shown that patterns of exposure condition asymmetrical knowledge of morphology/syntax (Giancaspro & Sánchez, 2021; Perez-Cortes et al., 2019; Sherkina-Lieber, 2015)


--
  + Interpretation/receptive knowledge > production


--
  + Has not been tested in older school-aged children


--
- Determining the origin of this variability has theoretical implications


--
  + When do children show adult-like tendencies?
  
  
--
  + Are these asymmetries a result of restructuring of HL grammars (increase over time) or are they typical of bilingual development (constant over time)? (See Montrul, 2013 and Putnam & Sánchez, 2013)


--
  + Are asymmetries proficiency-dependent?


---
# Research Questions and Hypotheses


--
**RQ1: What is the role of age in the development of differential object marking?**


Hypothesis: HS will produce and select DOM more frequently as they become older.


--
**RQ2: Do patterns of HL use and proficiency affect mastery of DOM?**


Hypothesis: both current frequency of use and proficiency will account for differences in DOM production.


--
**RQ3: Do HS show asymmetrical knowledge of DOM between production and selection?**


Hypothesis: Following previous research on adults, all participants will show asymmetrical knowledge favoring DOM selection over production.


---
# Participants


--
- 18 Spanish-dominant bilingual adults


--
  + Raised in 7 Spanish-speaking countries


--
  + Arrival in United States no earlier than age 12 (Ahn et al., 2017)


--
- 35 heritage speaker adults (university students, working professionals)


--
- 34 HS in 7th/8th grade (ages 12-14)


--
  + Two schools (one group graduated from K-5 immersion program, one in traditional school); no effect of school


--
  + Sequential bilinguals; primarily of Mexican descent


--
- 41 HS in 5th grade (ages 10-11)


--
  + Two schools (one in K-5 immersion program, one in traditional school); no effect of school


--
  + Sequential bilinguals; primarily of Mexican descent


---
# Tasks

--
- Background measures:


--
  + BESA proficiency (Peña et al., 2014)

--
  + DELE lexical/morphosyntactic test (Montrul & Slabakova, 2003) for adults only

--
  + Language use and exposure questionnaire (self-developed), 25-point scale with 5 contexts


--
- **Elicited Production Task (EPT)** sentence completion exercise


--
  + 10 items targeting DOM with transitive verbs in indicative subordinate clauses (14 for adults); animate/specific object

--
  + 6 additional intransitive distractors for all children

--
  + 31 additional distractors for adults (infinitivals, indicative sentences)


--
- **Forced Choice Task (FCT)** mood selection (receptive) measure

--
  + 8 sentences targeting differential object marking

--
  + 15 distractors evaluating mood selection for all participants

--
  + 32 additional distractors for adults


---
# EPT Example Sentences

--
**EPT:**

A veces Juanito se pone triste si sus hermanas dicen que no quieren hablar con él. ¿Qué quiere la mamá? Quiere que las hermanas _________ (LLAMAR) Juanito cada noche.


--
**FCT**

Juanito se pone feliz si las hermanas le prestan atención. ¿Qué quiere la mamá?
  + (A) Quiere que las hermanas llamen *a* Juanito cada día.
  + (B) Quiere que las hermanas llamen Juanito cada día.


---
# Data Analysis


--
- Binary dependent variable: production/selection (1) or omission (0) of DOM


--
- Independent variables:

--
  + Group (SDB as baseline, Adult HS, 7th/8th Grade HS, 5th Grade HS)

--
  + Task (EPT as baseline, FCT)

--
  + Frequency of use (sum of 5 1-5 Likert scales; 25-point continuum)
  
--
  + Proficiency (number of expected answers on 18-item BESA test; Peña et al., 2014)


--
- Two linear mixed effects models

--
  + Omnibus model: all data; group, task, and interaction

--
  + Heritage model: HS data; frequency of use, proficiency, and task; 2-way interactions with task
  
--
  + Random effects: participant and item


---
# Results

```{r, generate-plot}
Bar_Plot <- Aggregate %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
  labs(title = "Differential Object Marking by Group and Task", y = "Percentage of DOM Responses") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16))
```

```{r, print-bar-plot, fig.height = 8, fig.width = 16, fig.align = 'center', fig.retina = 2}
(Bar_Plot)
```

---
# Inferential Statistics: Omnibus Model

```{r, generate-stat-model}
Omnibus <- lmer(DOM_Use ~ + Group * Task +
                  (1 | Part_ID) + (1 | Item),
                data = Aggregate)
```

```{r, generate-omnibus-plot}
Omnibus_Plot <- plot_model(Omnibus, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
  scale_y_continuous(breaks = seq(-1.5, 1.5, 0.5),
                     limits = c(-1.5, 1.5)) +
  scale_x_discrete(labels = c("5th Grade HS : FCT", "7th/8th Grade HS : FCT", "Adult HS : FCT", "FCT", "5th Grade HS", "7th/8th Grade HS", "Adult HS", "(Intercept)")) +
  labs(title = "Results of Omnibus Model", y = "β Estimates") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-omnibus-plot, fig.height = 8, fig.width = 16, fig.align = 'center', fig.retina = 2}
(Omnibus_Plot)
```

---
# Group Comparisons

```{r, generate-tukey}
Omnibus_Pairwise <- emmeans(Omnibus, spec = "Group")
Omnibus_Tukey <- contrast(Omnibus_Pairwise, method = "pairwise")
Tukey_Table <- summary(Omnibus_Tukey)
```

```{r, show-Tukey-results}
kable(Tukey_Table[1:6], col.names = c("Contrast", "Estimate", "SE", "df", "t ratio", "p"), align = "lccccc")
```


--
**Differences significant at the *p* < .05 level:**

--
- SDB group was different from all HS groups


--
- HS adults were different from fifth grade group


---
# Inferential Statistics: Individual Variability Model

```{r, generate-heritage-model}
Heritage <- lmer(DOM_Use ~ BESA_Total_Std + Use_Joint_Std + Task + BESA_Total_Std:Task + Use_Joint_Std:Task +
                   (1 | Part_ID) + (1 | Item),
                 data = Aggregate_HS)
```

```{r, generate-heritage-plot}
Heritage_Plot <- plot_model(Heritage, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
  scale_y_continuous(breaks = seq(-1.5, 1.5, 0.5),
                     limits = c(-1.5, 1.5)) +
    scale_x_discrete(labels = c("Frequency of use : FCT", "Proficiency : FCT", "FCT", "Frequency of use", "Proficiency", "(Intercept)")) +
  labs(title = "Results of Heritage Individual Variability Model", y = "β Estimates") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-heritage-plot, fig.height = 8, fig.width = 16, fig.align = 'center', fig.retina = 2}
(Heritage_Plot)
```

---
# Summary of Findings


--
- Age modulates HS' knowledge of the HL (consistent with hypothesis 1)


--
- Proficiency and frequency of use had minimal roles on results (inconsistent with hypothesis 2)


--
  + BESA may not be an appropriate measure of proficiency for adults


--
  + No direct operationalization of exposure


--
- HS select more DOM than they produce this structure (consistent with hypothesis 3)


--
  + Age does not affect this process


--
  + Reassembly of features or differential exposure versus production?


--
- HS differ in the quantity of DOM production with animate and specific objects (proper nouns) when compared to Spanish-dominant participants


--
- HS do not categorically produce the *a* personal, but improve with age


--
  + Steady improvement across age groups


--
  + HS adults > 7th/8th graders > 5th graders (adult-5th grade contrast is significant)


--
  + Development of DOM differs from Spanish-dominant adults, even by adulthood


---
# Future Directions


--
- An even broader spectrum of age groups is necessary


--
  + Distinguishing protracted acquisition from attrition/restructuring


--
  + Better for plotting the developmental stages of bilingual development


--
- What is the role of COVID-19 in children's HL development?


---
# References

Aissen, J. (2003). Differential Object Marking: Iconicity vs. Economy. Natural Language & Linguistic Theory, 21(3), 435–483. https://doi.org/10.1023/A:1024109008573

Bossong, G. (1991). Differential object marking in Romance and beyond. In D. Wanner & D. A. Kibbee (Eds.), *Current issues in linguistic theory* (Vol. 69, p. 143). John Benjamins Publishing Company. https://doi.org/10.1075/cilt.69.14bos

Callen, M. C., & Miller, K. (2021). Children’s acquisition of variable differential object marking in Spanish. In Y. Gong & F. Kpogo (Eds.), *Proceedings of the 46th annual Boston University Conference on Language Development* (pp. 76–89). Cascadilla Proceedings Project.

Corbet, J., & Domínguez, L. (2020). The comprehension of tense–aspect morphology by Spanish heritage speakers in the United Kingdom. *Languages*, *5*(46). https://doi.org/10.3390/languages5040046

Cuza, A., & Miller, L. (2015). The protracted acquisition of past tense aspectual values in child heritage Spanish. In R. Klassen, J. M. Liceras, & E. Valenzuela (Eds.), *Issues in Hispanic and Lusophone linguistics* (Vol. 4, pp. 211–230). John Benjamins. https://doi.org/10.1075/ihll.4.11cuz

Giancaspro, D., & Sánchez, L. (2021). Me, mi, my: Innovation and variability in heritage speakers’ knowledge of inalienable possession. *Glossa: A Journal of General Linguistics*, *6*(1). https://doi.org/10.5334/gjgl.1240

---
# References

Guijarro-Fuentes, P., Pires, A., & Nediger, W. (2017). Delay in the acquisition of Differential Object Marking by Spanish monolingual and bilingual teenagers. *International Journal of Bilingualism*, *21*(2), 159–177. https://doi.org/10.1177/1367006915601249

Montrul, S. (2004). Subject and object expression in Spanish heritage speakers: A case of morphosyntactic convergence. *Bilingualism: Language and Cognition*, *7*(2), 125–142. https://doi.org/10.1017/S1366728904001464

Montrul, S. (2013). Incomplete L1 acquisition. In J. Herschensohn & M. Young-Scholten (Eds.), *The Cambridge handbook of second language acquisition* (pp. 353–371). Cambridge University Press. https://doi.org/10.1017/CBO9781139051729.022

Montrul, S., Bhatt, R., & Girju, R. (2015). Differential object marking in Spanish, Hindi, and Romanian as heritage languages. *Language*, *91*(3), 564–610. https://doi.org/10.1353/lan.2015.0035

Montrul, S., & Bowles, M. (2009). Back to basics: Incomplete knowledge of Differential Object Marking in Spanish heritage speakers. *Bilingualism: Language and Cognition*, *12*(3), 363–383. https://doi.org/10.1017/S1366728909990071

Montrul, S., & Potowski, K. (2007). Command of gender agreement in school-age Spanish-English bilingual Children. *International Journal of Bilingualism*, *11*(3), 301–328. https://doi.org/10.1177/13670069070110030301

---
# References

Montrul, S., & Sánchez-Walker, N. (2013). Differential object marking in child and adult Spanish heritage speakers. *Language Acquisition*, *20*(2), 109–132. https://doi.org/10.1080/10489223.2013.766741

Montrul, S., & Slabakova, R. (2003). Competence similarities between native and near-native speakers: An investigation of the preterite-imperfect contrast in Spanish. *Studies in Second Language Acquisition*, *25*(3), 351–398. https://doi.org/10.1017/S0272263103000159

Peña, E. D., Gutiérrez-Clellen, V. F., Iglesias, A., Goldstein, B. A., & Bedore, L. M. (2014). *Bilingual English Spanish Assessment (BESA).* AR Clinical Publishing.

Perez-Cortes, S., Putnam, M., & Sánchez, L. (2019). Differential access: Asymmetries in accessing features and building representations in heritage language grammars. *Languages*, *4*(81), 1–27. https://doi.org/10.3390/languages4040081

Putnam, M. T., & Sánchez, L. (2013). What’s so incomplete about incomplete acquisition?: A prolegomenon to modeling heritage language grammars. *Linguistic Approaches to Bilingualism*, *3*(4), 478–508. https://doi.org/10.1075/lab.3.4.04put

Rodríguez-Mondoñedo, M. (2008). The acquisition of Differential Object Marking in Spanish. *Probus*, *20*(1). https://doi.org/10.1515/PROBUS.2008.004

Sherkina-Lieber, M. (2015). Tense, aspect, and agreement in heritage Labrador Inuttitut: Do receptive bilinguals understand functional morphology? *Linguistic Approaches to Bilingualism*, *5*(1), 30–61. https://doi.org/10.1075/lab.5.1.02she

Torrego, E. (1998). *The dependencies of objects*. MIT Press.