---
title: "Immersion versus Non-Immersion: Exploring Pedagogical Practices and Research with K-12 Immersion Students"
subtitle : "10th National Symposium of Spanish as a Heritage Language (Harvard University)"
author: "Patrick D. Thane"
institute: "Rutgers University / University of Massachusetts"
date: "April 14, 2023"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["rutgers", "rutgers-fonts"]
    nature:
      beforeInit: "http://www.jvcasillas.com/ru_xaringan/js/ru_xaringan.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, echo = FALSE, include = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 99)

library(tidyverse)
library(lme4)
library(emmeans)
library(knitr)
library(kableExtra)
library(xaringan)
library(patchwork)
library(jtools)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
```

```{r, load-DOM-datasets}
# Load production data
SDB_DOM_EPT <- read_csv("../CSV Files/SDB/SDB DOM EPT.csv")

DLI78_DOM_EPT <- read.csv("../CSV Files/DLI-78/DLI-78 DOM EPT.csv") %>% 
  mutate(Group = "DLE-7/8")
  
MLS78_DOM_EPT <- read.csv("../CSV Files/MLS-78/MLS-78 DOM EPT.csv") %>% 
  mutate(Group = "MLS-7/8")

DLI5_DOM_EPT <- read.csv("../CSV Files/DLI-5/DLI-5 DOM EPT.csv") %>% 
  mutate(Group = "DLE-5")

MLS5_DOM_EPT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM EPT.csv") %>% 
  mutate(Group = "MLS-5")

# Load FCT data
SDB_DOM_FCT <- read_csv("../CSV Files/SDB/SDB DOM FCT.csv")

DLI78_DOM_FCT <- read.csv("../CSV Files/DLI-78/DLI-78 DOM FCT.csv") %>% 
  mutate(Group = "DLE-7/8")

MLS78_DOM_FCT <- read.csv("../CSV Files/MLS-78/MLS-78 DOM FCT.csv") %>% 
  mutate(Group = "MLS-7/8")

DLI5_DOM_FCT <- read.csv("../CSV Files/DLI-5/DLI-5 DOM FCT.csv") %>% 
  mutate(Group = "DLE-5")

MLS5_DOM_FCT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM FCT.csv") %>% 
  mutate(Group = "MLS-5")


# Bind datasets
EPT_DOM <- rbind(SDB_DOM_EPT, DLI78_DOM_EPT, MLS78_DOM_EPT, DLI5_DOM_EPT, MLS5_DOM_EPT)
EPT_DOM$Group <- factor(EPT_DOM$Group, levels = c("SDB", "DLE-7/8", "MLS-7/8", "DLE-5", "MLS-5"))

FCT_DOM <- rbind(SDB_DOM_FCT, DLI78_DOM_FCT, MLS78_DOM_FCT, DLI5_DOM_FCT, MLS5_DOM_FCT)
FCT_DOM$Group <- factor(FCT_DOM$Group, levels = c("SDB", "DLE-7/8", "MLS-7/8", "DLE-5", "MLS-5"))
```

```{r, generate-DOM-averages}
SDB_DOM_EPT_Structure <- aggregate(SDB_DOM_EPT$DOM_Use, list(SDB_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_DOM_EPT_Structure <- SDB_DOM_EPT_Structure %>% rename(Structure_Avg = x)
SDB_DOM_EPT_Structure <- left_join(SDB_DOM_EPT, SDB_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

SDB_DOM_FCT_Structure <- aggregate(SDB_DOM_FCT$DOM_Use, list(SDB_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_DOM_FCT_Structure <- SDB_DOM_FCT_Structure %>% rename(Structure_Avg = x)
SDB_DOM_FCT_Structure <- left_join(SDB_DOM_FCT, SDB_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_DOM_EPT_Structure <- aggregate(DLI78_DOM_EPT$DOM_Use, list(DLI78_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI78_DOM_EPT_Structure <- DLI78_DOM_EPT_Structure %>% rename(Structure_Avg = x)
DLI78_DOM_EPT_Structure <- left_join(DLI78_DOM_EPT, DLI78_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_DOM_FCT_Structure <- aggregate(DLI78_DOM_FCT$DOM_Use, list(DLI78_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI78_DOM_FCT_Structure <- DLI78_DOM_FCT_Structure %>% rename(Structure_Avg = x)
DLI78_DOM_FCT_Structure <- left_join(DLI78_DOM_FCT, DLI78_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_DOM_EPT_Structure <- aggregate(MLS78_DOM_EPT$DOM_Use, list(MLS78_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS78_DOM_EPT_Structure <- MLS78_DOM_EPT_Structure %>% rename(Structure_Avg = x)
MLS78_DOM_EPT_Structure <- left_join(MLS78_DOM_EPT, MLS78_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_DOM_FCT_Structure <- aggregate(MLS78_DOM_FCT$DOM_Use, list(MLS78_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS78_DOM_FCT_Structure <- MLS78_DOM_FCT_Structure %>% rename(Structure_Avg = x)
MLS78_DOM_FCT_Structure <- left_join(MLS78_DOM_FCT, MLS78_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_DOM_EPT_Structure <- aggregate(DLI5_DOM_EPT$DOM_Use, list(DLI5_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI5_DOM_EPT_Structure <- DLI5_DOM_EPT_Structure %>% rename(Structure_Avg = x)
DLI5_DOM_EPT_Structure <- left_join(DLI5_DOM_EPT, DLI5_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_DOM_FCT_Structure <- aggregate(DLI5_DOM_FCT$DOM_Use, list(DLI5_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI5_DOM_FCT_Structure <- DLI5_DOM_FCT_Structure %>% rename(Structure_Avg = x)
DLI5_DOM_FCT_Structure <- left_join(DLI5_DOM_FCT, DLI5_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_DOM_EPT_Structure <- aggregate(MLS5_DOM_EPT$DOM_Use, list(MLS5_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS5_DOM_EPT_Structure <- MLS5_DOM_EPT_Structure %>% rename(Structure_Avg = x)
MLS5_DOM_EPT_Structure <- left_join(MLS5_DOM_EPT, MLS5_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_DOM_FCT_Structure <- aggregate(MLS5_DOM_FCT$DOM_Use, list(MLS5_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS5_DOM_FCT_Structure <- MLS5_DOM_FCT_Structure %>% rename(Structure_Avg = x)
MLS5_DOM_FCT_Structure <- left_join(MLS5_DOM_FCT, MLS5_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

EPT_DOM_Structure <- rbind(SDB_DOM_EPT_Structure, DLI78_DOM_EPT_Structure, MLS78_DOM_EPT_Structure, DLI5_DOM_EPT_Structure, MLS5_DOM_EPT_Structure)

FCT_DOM_Structure <- rbind(SDB_DOM_FCT_Structure, DLI78_DOM_FCT_Structure, MLS78_DOM_FCT_Structure, DLI5_DOM_FCT_Structure, MLS5_DOM_FCT_Structure)

DOM <- rbind(EPT_DOM_Structure, FCT_DOM_Structure)
DOM$Group <- factor(DOM$Group, levels = c("SDB", "DLE-7/8", "MLS-7/8", "DLE-5", "MLS-5"))
```

```{r, load-subjunctive-datasets}
# Load production data
SDB_Subj_EPT <- read_csv("../CSV Files/SDB/SDB Subjunctive EPT.csv")

DLI78_Subj_EPT <- read.csv("../CSV Files/DLI-78/DLI-78 Subjunctive EPT.csv") %>% 
  mutate(Group = "DLE-7/8")

MLS78_Subj_EPT <- read.csv("../CSV Files/MLS-78/MLS-78 Subjunctive EPT.csv") %>% 
  mutate(Group = "MLS-7/8")

DLI5_Subj_EPT <- read.csv("../CSV Files/DLI-5/DLI-5 Subjunctive EPT.csv") %>% 
  mutate(Group = "DLE-5")

MLS5_Subj_EPT <- read.csv("../CSV Files/MLS-5/MLS-5 Subjunctive EPT.csv") %>% 
  mutate(Group = "MLS-5")


# Load FCT data
SDB_Subj_FCT <- read_csv("../CSV Files/SDB/SDB Subjunctive FCT.csv")

DLI78_Subj_FCT <- read.csv("../CSV Files/DLI-78/DLI-78 Subjunctive FCT.csv") %>% 
  mutate(Group = "DLE-7/8")

MLS78_Subj_FCT <- read.csv("../CSV Files/MLS-78/MLS-78 Subjunctive FCT.csv") %>% 
  mutate(Group = "MLS-7/8")

DLI5_Subj_FCT <- read.csv("../CSV Files/DLI-5/DLI-5 Subjunctive FCT.csv") %>% 
  mutate(Group = "DLE-5")

MLS5_Subj_FCT <- read.csv("../CSV Files/MLS-5/MLS-5 Subjunctive FCT.csv") %>% 
  mutate(Group = "MLS-5")
```

```{r, generate-subjunctive-averages}
# Calculate subjunctive rates by group
SDB_Subj_EPT_Structure <- aggregate(SDB_Subj_EPT$Mood_Use, list(SDB_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_EPT_Structure <- SDB_Subj_EPT_Structure %>% rename(Structure_Avg = x)
SDB_Subj_EPT_Structure <- left_join(SDB_Subj_EPT, SDB_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

SDB_Subj_FCT_Structure <- aggregate(SDB_Subj_FCT$Mood_Use, list(SDB_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_FCT_Structure <- SDB_Subj_FCT_Structure %>% rename(Structure_Avg = x)
SDB_Subj_FCT_Structure <- left_join(SDB_Subj_FCT, SDB_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_Subj_EPT_Structure <- aggregate(DLI78_Subj_EPT$Mood_Use, list(DLI78_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_EPT_Structure <- DLI78_Subj_EPT_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_EPT_Structure <- left_join(DLI78_Subj_EPT, DLI78_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_Subj_FCT_Structure <- aggregate(DLI78_Subj_FCT$Mood_Use, list(DLI78_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_FCT_Structure <- DLI78_Subj_FCT_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_FCT_Structure <- left_join(DLI78_Subj_FCT, DLI78_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_Subj_EPT_Structure <- aggregate(MLS78_Subj_EPT$Mood_Use, list(MLS78_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_EPT_Structure <- MLS78_Subj_EPT_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_EPT_Structure <- left_join(MLS78_Subj_EPT, MLS78_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_Subj_FCT_Structure <- aggregate(MLS78_Subj_FCT$Mood_Use, list(MLS78_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_FCT_Structure <- MLS78_Subj_FCT_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_FCT_Structure <- left_join(MLS78_Subj_FCT, MLS78_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_Subj_EPT_Structure <- aggregate(DLI5_Subj_EPT$Mood_Use, list(DLI5_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_EPT_Structure <- DLI5_Subj_EPT_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_EPT_Structure <- left_join(DLI5_Subj_EPT, DLI5_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_Subj_FCT_Structure <- aggregate(DLI5_Subj_FCT$Mood_Use, list(DLI5_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_FCT_Structure <- DLI5_Subj_FCT_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_FCT_Structure <- left_join(DLI5_Subj_FCT, DLI5_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_Subj_EPT_Structure <- aggregate(MLS5_Subj_EPT$Mood_Use, list(MLS5_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_EPT_Structure <- MLS5_Subj_EPT_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_EPT_Structure <- left_join(MLS5_Subj_EPT, MLS5_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_Subj_FCT_Structure <- aggregate(MLS5_Subj_FCT$Mood_Use, list(MLS5_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_FCT_Structure <- MLS5_Subj_FCT_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_FCT_Structure <- left_join(MLS5_Subj_FCT, MLS5_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


# Rejoin dataframes
EPT_Subj_Structure <- rbind(SDB_Subj_EPT_Structure, DLI78_Subj_EPT_Structure, MLS78_Subj_EPT_Structure, DLI5_Subj_EPT_Structure, MLS5_Subj_EPT_Structure)

FCT_Subj_Structure <- rbind(SDB_Subj_FCT_Structure, DLI78_Subj_FCT_Structure, MLS78_Subj_FCT_Structure, DLI5_Subj_FCT_Structure, MLS5_Subj_FCT_Structure)

Subj <- rbind(EPT_Subj_Structure, FCT_Subj_Structure)
Subj$Group <- factor(Subj$Group, levels = c("SDB", "DLE-7/8", "MLS-7/8", "DLE-5", "MLS-5"))
```

```{r, separate-subjunctive-types}
## Separate out polarity/intensional
DLI5_Subj_EPT_Intensional <- DLI5_Subj_EPT %>% filter(Property == "Intensional subjunctive")
DLI5_Subj_EPT_Polarity <- DLI5_Subj_EPT %>% filter(Property == "Polarity subjunctive")
DLI5_Subj_FCT_Intensional <- DLI5_Subj_FCT %>% filter(Property == "Intensional subjunctive")
DLI5_Subj_FCT_Polarity <- DLI5_Subj_FCT %>% filter(Property == "Polarity subjunctive")

MLS5_Subj_EPT_Intensional <- MLS5_Subj_EPT %>% filter(Property == "Intensional subjunctive")
MLS5_Subj_EPT_Polarity <- MLS5_Subj_EPT %>% filter(Property == "Polarity subjunctive")
MLS5_Subj_FCT_Intensional <- MLS5_Subj_FCT %>% filter(Property == "Intensional subjunctive")
MLS5_Subj_FCT_Polarity <- MLS5_Subj_FCT %>% filter(Property == "Polarity subjunctive")

DLI78_Subj_EPT_Intensional <- DLI78_Subj_EPT %>% filter(Property == "Intensional subjunctive")
DLI78_Subj_EPT_Polarity <- DLI78_Subj_EPT %>% filter(Property == "Polarity subjunctive")
DLI78_Subj_FCT_Intensional <- DLI78_Subj_FCT %>% filter(Property == "Intensional subjunctive")
DLI78_Subj_FCT_Polarity <- DLI78_Subj_FCT %>% filter(Property == "Polarity subjunctive")

MLS78_Subj_EPT_Intensional <- MLS78_Subj_EPT %>% filter(Property == "Intensional subjunctive")
MLS78_Subj_EPT_Polarity <- MLS78_Subj_EPT %>% filter(Property == "Polarity subjunctive")
MLS78_Subj_FCT_Intensional <- MLS78_Subj_FCT %>% filter(Property == "Intensional subjunctive")
MLS78_Subj_FCT_Polarity <- MLS78_Subj_FCT %>% filter(Property == "Polarity subjunctive")

SDB_Subj_EPT_Intensional <- SDB_Subj_EPT %>% filter(Property == "Intensional subjunctive")
SDB_Subj_EPT_Polarity <- SDB_Subj_EPT %>% filter(Property == "Polarity subjunctive")
SDB_Subj_FCT_Intensional <- SDB_Subj_FCT %>% filter(Property == "Intensional subjunctive")
SDB_Subj_FCT_Polarity <- SDB_Subj_FCT %>% filter(Property == "Polarity subjunctive")
```

```{r, generate-intensional-percentage}
# Generate intensional averages by group
SDB_Subj_EPT_Intensional_Structure <- aggregate(SDB_Subj_EPT_Intensional$Mood_Use, list(SDB_Subj_EPT_Intensional$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_EPT_Intensional_Structure <- SDB_Subj_EPT_Intensional_Structure %>% rename(Structure_Avg = x)
SDB_Subj_EPT_Intensional_Structure <- left_join(SDB_Subj_EPT_Intensional, SDB_Subj_EPT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

SDB_Subj_FCT_Intensional_Structure <- aggregate(SDB_Subj_FCT_Intensional$Mood_Use, list(SDB_Subj_FCT_Intensional$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_FCT_Intensional_Structure <- SDB_Subj_FCT_Intensional_Structure %>% rename(Structure_Avg = x)
SDB_Subj_FCT_Intensional_Structure <- left_join(SDB_Subj_FCT_Intensional, SDB_Subj_FCT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_Subj_EPT_Intensional_Structure <- aggregate(DLI78_Subj_EPT_Intensional$Mood_Use, list(DLI78_Subj_EPT_Intensional$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_EPT_Intensional_Structure <- DLI78_Subj_EPT_Intensional_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_EPT_Intensional_Structure <- left_join(DLI78_Subj_EPT_Intensional, DLI78_Subj_EPT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_Subj_FCT_Intensional_Structure <- aggregate(DLI78_Subj_FCT_Intensional$Mood_Use, list(DLI78_Subj_FCT_Intensional$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_FCT_Intensional_Structure <- DLI78_Subj_FCT_Intensional_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_FCT_Intensional_Structure <- left_join(DLI78_Subj_FCT_Intensional, DLI78_Subj_FCT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_Subj_EPT_Intensional_Structure <- aggregate(MLS78_Subj_EPT_Intensional$Mood_Use, list(MLS78_Subj_EPT_Intensional$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_EPT_Intensional_Structure <- MLS78_Subj_EPT_Intensional_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_EPT_Intensional_Structure <- left_join(MLS78_Subj_EPT_Intensional, MLS78_Subj_EPT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_Subj_FCT_Intensional_Structure <- aggregate(MLS78_Subj_FCT_Intensional$Mood_Use, list(MLS78_Subj_FCT_Intensional$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_FCT_Intensional_Structure <- MLS78_Subj_FCT_Intensional_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_FCT_Intensional_Structure <- left_join(MLS78_Subj_FCT_Intensional, MLS78_Subj_FCT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_Subj_EPT_Intensional_Structure <- aggregate(DLI5_Subj_EPT_Intensional$Mood_Use, list(DLI5_Subj_EPT_Intensional$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_EPT_Intensional_Structure <- DLI5_Subj_EPT_Intensional_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_EPT_Intensional_Structure <- left_join(DLI5_Subj_EPT_Intensional, DLI5_Subj_EPT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_Subj_FCT_Intensional_Structure <- aggregate(DLI5_Subj_FCT_Intensional$Mood_Use, list(DLI5_Subj_FCT_Intensional$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_FCT_Intensional_Structure <- DLI5_Subj_FCT_Intensional_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_FCT_Intensional_Structure <- left_join(DLI5_Subj_FCT_Intensional, DLI5_Subj_FCT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_Subj_EPT_Intensional_Structure <- aggregate(MLS5_Subj_EPT_Intensional$Mood_Use, list(MLS5_Subj_EPT_Intensional$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_EPT_Intensional_Structure <- MLS5_Subj_EPT_Intensional_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_EPT_Intensional_Structure <- left_join(MLS5_Subj_EPT_Intensional, MLS5_Subj_EPT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_Subj_FCT_Intensional_Structure <- aggregate(MLS5_Subj_FCT_Intensional$Mood_Use, list(MLS5_Subj_FCT_Intensional$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_FCT_Intensional_Structure <- MLS5_Subj_FCT_Intensional_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_FCT_Intensional_Structure <- left_join(MLS5_Subj_FCT_Intensional, MLS5_Subj_FCT_Intensional_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

# Rejoin dataframes
EPT_Intensional_Structure <- rbind(SDB_Subj_EPT_Intensional_Structure, DLI78_Subj_EPT_Intensional_Structure, MLS78_Subj_EPT_Intensional_Structure, DLI5_Subj_EPT_Intensional_Structure, MLS5_Subj_EPT_Intensional_Structure) %>% 
  mutate(Clause_Type = "Volitional Clauses")

FCT_Intensional_Structure <- rbind(SDB_Subj_FCT_Intensional_Structure, DLI78_Subj_FCT_Intensional_Structure, MLS78_Subj_FCT_Intensional_Structure, DLI5_Subj_FCT_Intensional_Structure, MLS5_Subj_FCT_Intensional_Structure) %>% 
  mutate(Clause_Type = "Volitional Clauses")

Intensional <- rbind(EPT_Intensional_Structure, FCT_Intensional_Structure)
```

```{r, generate-polarity-average}
# Generate polarity averages by group
SDB_Subj_EPT_Polarity_Structure <- aggregate(SDB_Subj_EPT_Polarity$Mood_Use, list(SDB_Subj_EPT_Polarity$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_EPT_Polarity_Structure <- SDB_Subj_EPT_Polarity_Structure %>% rename(Structure_Avg = x)
SDB_Subj_EPT_Polarity_Structure <- left_join(SDB_Subj_EPT_Polarity, SDB_Subj_EPT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

SDB_Subj_FCT_Polarity_Structure <- aggregate(SDB_Subj_FCT_Polarity$Mood_Use, list(SDB_Subj_FCT_Polarity$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_FCT_Polarity_Structure <- SDB_Subj_FCT_Polarity_Structure %>% rename(Structure_Avg = x)
SDB_Subj_FCT_Polarity_Structure <- left_join(SDB_Subj_FCT_Polarity, SDB_Subj_FCT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_Subj_EPT_Polarity_Structure <- aggregate(DLI78_Subj_EPT_Polarity$Mood_Use, list(DLI78_Subj_EPT_Polarity$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_EPT_Polarity_Structure <- DLI78_Subj_EPT_Polarity_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_EPT_Polarity_Structure <- left_join(DLI78_Subj_EPT_Polarity, DLI78_Subj_EPT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_Subj_FCT_Polarity_Structure <- aggregate(DLI78_Subj_FCT_Polarity$Mood_Use, list(DLI78_Subj_FCT_Polarity$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_FCT_Polarity_Structure <- DLI78_Subj_FCT_Polarity_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_FCT_Polarity_Structure <- left_join(DLI78_Subj_FCT_Polarity, DLI78_Subj_FCT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_Subj_EPT_Polarity_Structure <- aggregate(MLS78_Subj_EPT_Polarity$Mood_Use, list(MLS78_Subj_EPT_Polarity$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_EPT_Polarity_Structure <- MLS78_Subj_EPT_Polarity_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_EPT_Polarity_Structure <- left_join(MLS78_Subj_EPT_Polarity, MLS78_Subj_EPT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_Subj_FCT_Polarity_Structure <- aggregate(MLS78_Subj_FCT_Polarity$Mood_Use, list(MLS78_Subj_FCT_Polarity$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_FCT_Polarity_Structure <- MLS78_Subj_FCT_Polarity_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_FCT_Polarity_Structure <- left_join(MLS78_Subj_FCT_Polarity, MLS78_Subj_FCT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_Subj_EPT_Polarity_Structure <- aggregate(DLI5_Subj_EPT_Polarity$Mood_Use, list(DLI5_Subj_EPT_Polarity$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_EPT_Polarity_Structure <- DLI5_Subj_EPT_Polarity_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_EPT_Polarity_Structure <- left_join(DLI5_Subj_EPT_Polarity, DLI5_Subj_EPT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_Subj_FCT_Polarity_Structure <- aggregate(DLI5_Subj_FCT_Polarity$Mood_Use, list(DLI5_Subj_FCT_Polarity$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_FCT_Polarity_Structure <- DLI5_Subj_FCT_Polarity_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_FCT_Polarity_Structure <- left_join(DLI5_Subj_FCT_Polarity, DLI5_Subj_FCT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_Subj_EPT_Polarity_Structure <- aggregate(MLS5_Subj_EPT_Polarity$Mood_Use, list(MLS5_Subj_EPT_Polarity$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_EPT_Polarity_Structure <- MLS5_Subj_EPT_Polarity_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_EPT_Polarity_Structure <- left_join(MLS5_Subj_EPT_Polarity, MLS5_Subj_EPT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_Subj_FCT_Polarity_Structure <- aggregate(MLS5_Subj_FCT_Polarity$Mood_Use, list(MLS5_Subj_FCT_Polarity$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_FCT_Polarity_Structure <- MLS5_Subj_FCT_Polarity_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_FCT_Polarity_Structure <- left_join(MLS5_Subj_FCT_Polarity, MLS5_Subj_FCT_Polarity_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

# Rejoin dataframes
EPT_Polarity_Structure <- rbind(SDB_Subj_EPT_Polarity_Structure, DLI78_Subj_EPT_Polarity_Structure, MLS78_Subj_EPT_Polarity_Structure, DLI5_Subj_EPT_Polarity_Structure, MLS5_Subj_EPT_Polarity_Structure) %>% 
  mutate(Clause_Type = "Relative Clauses")

FCT_Polarity_Structure <- rbind(SDB_Subj_FCT_Polarity_Structure, DLI78_Subj_FCT_Polarity_Structure, MLS78_Subj_FCT_Polarity_Structure, DLI5_Subj_FCT_Polarity_Structure, MLS5_Subj_FCT_Polarity_Structure) %>% 
  mutate(Clause_Type = "Relative Clauses")

Polarity <- rbind(EPT_Polarity_Structure, FCT_Polarity_Structure)
```

```{r, rejoin-subjunctive-types}
EPT_Subj_Structure <- rbind(EPT_Intensional_Structure, EPT_Polarity_Structure)
FCT_Subj_Structure <- rbind(FCT_Intensional_Structure, FCT_Polarity_Structure)

Aggregate <- rbind(Polarity, Intensional)
Aggregate$Clause_Type <- factor(Aggregate$Clause_Type, levels = c("Volitional Clauses", "Relative Clauses"))
Aggregate$Group <- factor(Aggregate$Group, levels = c("SDB", "DLE-7/8", "MLS-7/8", "DLE-5", "MLS-5"))

```

# Statement on Labor Negotiations at Rutgers


--
I stand in solidarity with my graduate student colleagues, as well as members of the AAUP unions that represent adjunct, teaching, tenured, and postdoctoral faculty. We also are representing members of Rutgers' administrative and custodial staff in an attempt to gain more favorable funding, stable healthcare, and job permanence. We are also fighting to obtain rent stability for undergraduate and graduate students, not only for the benefit of students but also to counter gentrification that has already begun to displace members of the surrounding communities, a disproportionate amount of whom are Hispanic.

The union has not explicitly dissuaded graduate students from pursuing scholarly activity, including conference travel. It behooves those of us who are active in pursuing more equitable educational opportunities for Spanish heritage speakers to continue scholarly activity related to these topics. I dedicate today's presentation to the New Brunswick community, some of whom participated in this study, in our unified efforts to create a more inclusive Rutgers and surrounding region with a vision for more accessible public university education.

The outcomes of the strike at Rutgers, which began on April 10, will have minor economic impact on my salary, as I have accepted a postdoctoral position at the University of Massachusetts for the coming academic year. However, I am striking primarily in support of our joint unions' Bargaining for the Common Good initiatives (https://smlr.rutgers.edu/faculty-research-engagement/center-innovation-worker-organization-ciwo/bargaining-common-good).


---
# Introduction to Project


--
- Results from broad study on language acquisition in bilingual programs

--
  + Comparison of dual-language education (DLE) and monolingual school (MLS)

--
  + Exploration of long-term effects of immersion


--
- Study of 75 bilingual children and 18 Spanish-dominant adults in 2 educational contexts:

--
  + 18 Spanish-dominant bilingual adults from 7 countries (SDB)

--
  + 11 heritage speakers (HS) in 7th/8th who graduated from 50/50 DLE school (DLE-7/8)

--
  + 23 HS in 7th/8th grade from the monolingual school (MLS-7/8)

--
  + 19 HS in 5th grade at the end of the 50/50 DLE program (DLE-5)

--
  + 22 HS in 5th grade from the monolingual school (MLS-5)


--
- Schools were matched for demographic characteristics

--
  + 77% vs. 78% of students were reported as "economically disadvantaged"

--
  + Both schools reported that 86% of their population was Hispanic

--
  + 21% vs. 22% classified as English language learners

--
  + Students primarily of Mexican-American descent; sequential bilinguals


---
# Structures Tested


--
- Differential object marking (DOM): use of the "*a* personal" with animate, specific direct objects


--
  + *Ve Juana (Juana sees)* versus *Ve **a** Juana (She sees Juana)*


--
  + Shown to be highly variable in Spanish as a heritage language, in children and adults (e.g., Guijarro-Fuentes et al., 2017; Montrul & Sánchez-Walker, 2013)


--
- Subjunctive mood: use of verbal conjugations in multiple contexts to mark modality


--
  + Use in volitional clauses with verbs such as *querer*: *quiero que **vengas** a la fiesta*


--
  + Use in relative clauses to describe nouns whose entities are unknown: *busco una casa que **tenga** paredes invisibles*


--
  + Spanish HS exhibit different levels of subjunctive mood knowledge than monolinguals (Dracos & Requena, 2022; Giancaspro, 2019; Montrul, 2009)


--
- Production and selection task (indicative/subjunctive, DOM/no DOM)


---
# Comparing HS' Knowledge of DOM and Subjunctive Mood

```{r, generate-DOM-plot}
DOM_Plot <- DOM %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
  labs(title = "Differential Object Marking by Group and Task", y = "Percentage of DOM Responses") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16))
```

```{r, generate-subj-plot}
Subj_Plot <- Subj %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
  labs(title = "Subjunctive by Group and Task", y = "Percentage of Subjunctive Responses") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16))
```

```{r, print-structure-bar-plots, fig.height = 8, fig.width = 16, fig.align = 'center', fig.retina = 2}
(DOM_Plot)+(Subj_Plot)
```

---
# Comparing Subtypes of Subjunctive Mood

```{r, generate-subj-structure-plots}
Subj_Structure_Plot <- Aggregate %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_grid(cols = vars(Clause_Type)) +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
  labs(title = "Subjunctive Production by Task, Clause Type, and Group", y = "Percentage of Subjunctive Responses") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16),
        strip.text.x = element_text(size = 18, face = "bold"))
```

```{r, print-subj-bar-plots, fig.height = 8, fig.width = 16, fig.align = 'center', fig.retina = 2}
(Subj_Structure_Plot)
```

---
# Summarizing Results


--
- Generally, students in a 50/50 DLE program did not show advantages over age-matched peers in a monolingual program:


--
  + Students in 7th/8th grade showed higher productive/receptive command of differential object marking, but these students were no longer in DLE


--
  + No differences were detectable in the immersion students for subjunctive mood at any age group


--
- Exposure to the heritage language at school does not guarantee its acquisition


--
  + Overhearing the language does not guarantee strong receptive skills


--
  + Production and comprehension appear to be uncoordinated in HS' grammatical systems (Perez-Cortes et al., 2019)


--
  + In L2 acquisition, we know that output and input are both necessary ingredients (Swain, 2000)


--
  + Immersion students often respond far more in English, even if they hear Spanish daily (Potowski, 2004)


--
- Take-home message: language learning requires production, interaction, and structured development


---
# Pedagogical Translanguaging


--
- Cenoz (2017, p. 194): “Planned by the teacher inside the classroom and can refer to the use of different languages for input and output or to other planned strategies based on the use of students’ resources from the whole linguistic repertoire.”


--
- Balance between flexibility necessary for students to succeed and structure to leverage bilingual development


--
  + Teacher language use is planned and organized into instructional components


--
  + Student language use is flexible and fluid, but certain tasks occur in each language


--
- The happy medium


--
  + Tolerant of students' needs and of growing linguistic system


--
  + Leveraging of literacy strategies to develop and support heritage language growth


---
# Teaching for Biliteracy


--
- Lessons are broken into three components:


--
  + Target language zone: language use in the allocated language
  

--
  + **Bridge**: translanguaging space; contrastive analysis of predetermined linguistic structure or vocabulary relevant to the content lesson


--
  + Extension: use of target language to accomplish content tasks through the use of new grammar (also functions as transition to other content area through thematic units)


--
- Contrastive analysis built into language content objectives


--
- Research-based evidence in favor of this approach:


--
  + Leonet et al. (2020): fifth grade children showed stronger morphological awareness in 3 languages in translanguaging group than children in traditional group


--
  + Lyster (2013): second grade students benefit from contrastive analysis between L1 and L2 (advantage over children in language separation group)

---
# Processing Instruction


--
- Highly-researched teaching strategy in second language acquisition (see VanPatten, 1996, 2015)


--
  + Supported by generations of research in input processing theory


--
  + Techniques to draw attention towards a specific form that results in internalization of grammatical rules and paradigms
  

--
  + Useful for situating input activities within a content-based lesson


--
- Strategy for teaching verb inflections and word order


--
  + Brief explicit instruction with explanation of a specific component of grammar
  

--
  + Communicative tasks designed to carefully manipulate a specific structure


--
  + Students must complete simple comprehension activities whose answers depend on the interpretation of the structure in question


--
- It is critical to recognize that PI is effective for providing exposure to the language in question, but it is not the *only* strategy necessary to support development and interaction


--
  + More useful with some structures than others (e.g., those that can be fronted)


--
  + Does not concentrate on language output


---
# Making the Case for PI in K-12 Language Arts


--
- Evidence-based approach for multiple populations:


--
  + Adult L2 learners (see VanPatten, 2015 for review)


--
  + Child L2 learners (Benati & Lee, 2008)


--
  + Adult heritage learners (see Bowles, 2021, 2022 and contributions within)


--
  + *Child heritage learners?*
  

--
- Can easily be integrated into content-area lessons or during "bridging" activities


--
  + Structured input activities focus on a broader content objective (communication-focused)


--
  + Ideal opportunity to marry language learning objectives and content


--
  + Production activities in the heritage language are an ideal way to extend on classroom knowledge


--
  + PI could form part of the bridging sequence
  

---
# Takeaways


- Language development in immersion schools is not automatic or guaranteed


--
  + This result SHOULD NOT be interpreted as an argument against dual-language education


--
  + Specific pedagogies are necessary to guarantee successful development of multiple languages


--
- Pedagogical translanguaging offers a meaningful middle ground for HL instruction in immersion programs


--
  + Teachers are regimented in their language use, but students can translanguage freely with one another as a scaffolding tool


--
  + Supports biliteracy development in a structured way


--
- Teaching for biliteracy and Processing Instruction are useful tools for supporting HL development


--
  + Involvement of explicit teaching that involves a translanguaging component of all lessons/units


--
  + Strategic language use; communicative tasks to foster metalinguistic awareness


--
  + Cultivation of independent language learners


---
# References

Benati, A. G., & Lee, J. F. (2008). *Grammar acquisition and processing instruction: Secondary and cumulative effects*. Multilingual Matters.

Bowles, M. A. (2021). The interplay between heritage learners’ prior knowledge and instruction on a linguistically complex structure. *Heritage Language Journal*, *18*(3), 1–29. https://doi.org/10.1163/15507076-12340015

Bowles, M. A. (Ed.). (2022). *Outcomes of university Spanish heritage language instruction in the United States.* Georgetown University Press.

Cenoz, J. (2017). Translanguaging in school contexts: International perspectives. *Journal of Language, Identity & Education*, *16*(4), 193–198. https://doi.org/10.1080/15348458.2017.1327816

Cenoz, J., & Gorter, D. (2022). Pedagogical translanguaging and its application to language classes. *RELC Journal*, *53*(2), 342–354. https://doi.org/10.1177/00336882221082751

Cuza, A., & Miller, L. (2015). The protracted acquisition of past tense aspectual values in child heritage Spanish. In R. Klassen, J. M. Liceras, & E. Valenzuela (Eds.), *Issues in Hispanic and Lusophone linguistics* (Vol. 4, pp. 211–230). John Benjamins. https://doi.org/10.1075/ihll.4.11cuz

Dracos, M., & Requena, P. E. (2022). Child heritage speakers’ acquisition of the Spanish subjunctive in volitional and adverbial clauses. *Language Acquisition*, 1–28. https://doi.org/10.1080/10489223.2022.2071156

---
# References

García, O. (2009). *Bilingual education in the 21st century: A global perspective*. Wiley-Blackwell Pub.

Giancaspro, D. (2019). The late(r) bird gets the verb? Effects of age of acquisition of English on adult heritage speakers’ knowledge of subjunctive mood in Spanish. *Languages*, *4*(3), 69. https://doi.org/10.3390/languages4030069

Guijarro-Fuentes, P., Pires, A., & Nediger, W. (2017). Delay in the acquisition of Differential Object Marking by Spanish monolingual and bilingual teenagers. *International Journal of Bilingualism*, *21*(2), 159–177. https://doi.org/10.1177/1367006915601249

Leonet, O., Cenoz, J., & Gorter, D. (2020). Developing morphological awareness across languages: Translanguaging pedagogies in third language acquisition. *Language Awareness*, *29*(1), 41–59. https://doi.org/10.1080/09658416.2019.1688338

Lyster, R., Quiroga, J., & Ballinger, S. (2013). The effects of biliteracy instruction on morphological awareness. *Journal of Immersion and Content-Based Language Education*, *1*(2), 169–197. https://doi.org/10.1075/jicb.1.2.02lys

Montrul, S. (2009). Knowledge of tense-aspect and mood in Spanish heritage speakers. International Journal of Bilingualism, 13(2), 239–269. https://doi.org/10.1177/1367006909339816
Montrul, S., & Sánchez-Walker, N. (2013). Differential object marking in child and adult Spanish heritage speakers. *Language Acquisition*, *20*(2), 109–132. https://doi.org/10.1080/10489223.2013.766741

---
# References

Perez-Cortes, S., Putnam, M., & Sánchez, L. (2019). Differential access: Asymmetries in accessing features and building representations in heritage language grammars. *Languages*, *4*(81), 1–27. https://doi.org/10.3390/languages4040081

Potowski, K. (2004). Student Spanish use and investment in a dual immersion classroom: Implications for second language acquisition and heritage language maintenance. *The Modern Language Journal*, *88*(1), 75–101. https://doi.org/10.1111/j.0026-7902.2004.00219.x

Sánchez, M. T., García, O., & Solorza, C. (2018). Reframing language allocation policy in dual language bilingual education. *Bilingual Research Journal*, *41*(1), 37–51. https://doi.org/10.1080/15235882.2017.1405098

Swain, M. (2000). French immersion research in Canada: Recent contributions to SLA and applied linguistics. *Annual Review of Applied Linguistics*, *20*, 199–212. https://doi.org/10.1017/S0267190500200123

VanPatten, B. (2015). Foundations of processing instruction. *International Review of Applied Linguistics in Language Teaching*, *53*(2). https://doi.org/10.1515/iral-2015-0005


---
# Example of Processing Instruction in DLI


--
In a content lesson on history, students may need to learn to recognize the difference between habitual and one-time actions in Spanish. One of the language learning objectives in Spanish language arts is narrative development, so knowledge of the preterit/imperfect will be essential for both of these content areas.


--
During the class, the teacher may provide students with a reading that highlights the use of preterit and imperfect verbs in different colors. The teacher may also conduct a bridge activity comparing the use of past tense in English and Spanish, if s/he/they is following a TfB framework.


--
During language arts, students may receive a sheet where they must circle if the sentences they hear refer to a "one-time event" or a "repeated event." They need to listen to a series of sentences that are related to their reading (*not* taken directly from it) that they have heard.


--
Example sentences do not include adverbs or any information besides the verb to indicate the correct answer. These sentences could include:


--
- El pueblo elegía a los presidentes cada 4 años (repeated event/evento repetido)


--
- George Washington ganó la presidencia en 1789.

