---
title: "Empiricalizing Heritage Spanish Development in Dual-Language Immersion: Initial Insights"
author: "Patrick D. Thane, Ph.D. (él/he/him/his)<br/>pthane@umass.edu"
institute: "LLC XXVI: UMass College of Education<br/>https://tinyurl.com/4sathhsw"
date: "November 29, 2023"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r, setup, include = FALSE}
library(tidyverse)
library(lme4)
library(xaringan)
library(xaringanthemer)
library(xaringanExtra)
library(lmerTest)
library(kableExtra)
library(patchwork)
library(jtools)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(emmeans)
library(here)

options(htmltools.dir.version = FALSE)
options(scipen = 999)


knitr::opts_chunk$set(fig.width = 16, fig.height = 7, fig.retina = 3, fig.align = 'center', cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, hiline = TRUE)
```

```{r xaringan, include = FALSE, warning = FALSE}
style_duo_accent(
  primary_color = "#881C1C",
  secondary_color = "#212721",
  inverse_header_color = "#FFFDD0")
```

```{r xaringan-logo, echo = FALSE}
use_logo(
  image_url = "https://www.umass.edu/brand/sites/default/files/M-lockedup.png"
)
```

```{r, load-DOM-data}
# EPT
SDB_DOM_EPT <- read_csv(here("./CSV Files/SDB/SDB DOM EPT.csv")) %>% 
    filter(Item %in% c("EPT-02", "EPT-08", "EPT-12", "EPT-14", "EPT-20", "EPT-24", "EPT-26", "EPT-32", "EPT-38", "EPT-42")) %>% 
  mutate(Group = "SDBA",
         Task = "SCT",
         School = "SDBA",
         Age = "SDBA")

DLI78_DOM_EPT <- read_csv(here("./CSV Files/DLI-78/DLI-78 DOM EPT.csv")) %>% 
  mutate(Group = "DLI-7/8",
         Task = "SCT",
         School = "DLI",
         Age = "7th/8th")

MLS78_DOM_EPT <- read_csv(here("./CSV Files/MLS-78/MLS-78 DOM EPT.csv")) %>% 
  filter(!School == "GBCS") %>%
  mutate(Group = "MLS-7/8",
         Task = "SCT",
         School = "MLS",
         Age = "7th/8th")

DLI5_DOM_EPT <- read_csv(here("./CSV Files/DLI-5/DLI-5 DOM EPT.csv")) %>% 
  mutate(Group = "DLI-5",
         Task = "SCT",
         School = "DLI",
         Age = "5th")

MLS5_DOM_EPT <- read_csv(here("./CSV Files/MLS-5/MLS-5 DOM EPT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  mutate(Group = "MLS-5",
         Task = "SCT",
         School = "MLS",
         Age = "5th")


# FCT
SDB_DOM_FCT <- read_csv(here("./CSV Files/SDB/SDB DOM FCT.csv")) %>% 
  mutate(Group = "SDBA",
         Task = "MST",
         School = "SDBA",
         Age = "SDBA")

DLI78_DOM_FCT <- read_csv(here("./CSV Files/DLI-78/DLI-78 DOM FCT.csv")) %>% 
  mutate(Group = "DLI-7/8",
         Task = "MST",
         School = "DLI",
         Age = "7th/8th")

MLS78_DOM_FCT <- read_csv(here("./CSV Files/MLS-78/MLS-78 DOM FCT.csv")) %>% 
  filter(!School == "GBCS") %>%
  mutate(Group = "MLS-7/8",
         Task = "MST",
         School = "MLS",
         Age = "7th/8th")

DLI5_DOM_FCT <- read_csv(here("./CSV Files/DLI-5/DLI-5 DOM FCT.csv")) %>% 
  mutate(Group = "DLI-5",
         Task = "MST",
         School = "DLI",
         Age = "5th")

MLS5_DOM_FCT <- read_csv(here("./CSV Files/MLS-5/MLS-5 DOM FCT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  mutate(Group = "MLS-5",
         Task = "MST",
         School = "MLS",
         Age = "5th")


## Join datasets
DOM_EPT <- rbind(SDB_DOM_EPT, DLI78_DOM_EPT, MLS78_DOM_EPT, DLI5_DOM_EPT, MLS5_DOM_EPT)
DOM_FCT <- rbind(SDB_DOM_FCT, DLI78_DOM_FCT, MLS78_DOM_FCT, DLI5_DOM_FCT, MLS5_DOM_FCT)
```

```{r, load-subj-data}
# EPT
SDB_Subj_EPT <- read_csv(here("./CSV Files/SDB/SDB Subjunctive EPT.csv")) %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "SDBA",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "SDBA",
         Age = "SDBA")

DLI78_Subj_EPT <- read_csv(here("./CSV Files/DLI-78/DLI-78 Subjunctive EPT.csv")) %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-7/8",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "7th/8th")

MLS78_Subj_EPT <- read_csv(here("./CSV Files/MLS-78/MLS-78 Subjunctive EPT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-7/8",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "7th/8th")

DLI5_Subj_EPT <- read_csv(here("./CSV Files/DLI-5/DLI-5 Subjunctive EPT.csv"))%>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-5",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "5th")

MLS5_Subj_EPT <- read_csv(here("./CSV Files/MLS-5/MLS-5 Subjunctive EPT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-5",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "5th")


# FCT
SDB_Subj_FCT <- read_csv(here("./CSV Files/SDB/SDB Subjunctive FCT.csv")) %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "SDBA",
         Task = "MST",
         Mood = "Subjunctive",
         School = "SDBA",
         Age = "SDBA")

DLI78_Subj_FCT <- read_csv(here("./CSV Files/DLI-78/DLI-78 Subjunctive FCT.csv")) %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-7/8",
         Task = "MST",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "7th/8th")

MLS78_Subj_FCT <- read_csv(here("./CSV Files/MLS-78/MLS-78 Subjunctive FCT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-7/8",
         Task = "MST",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "7th/8th")

DLI5_Subj_FCT <- read_csv(here("./CSV Files/DLI-5/DLI-5 Subjunctive FCT.csv")) %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-5",
         Task = "MST",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "5th")

MLS5_Subj_FCT <- read_csv(here("./CSV Files/MLS-5/MLS-5 Subjunctive FCT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-5",
         Task = "MST",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "5th")


## Join datasets
Subj_EPT <- rbind(SDB_Subj_EPT, DLI78_Subj_EPT, MLS78_Subj_EPT, DLI5_Subj_EPT, MLS5_Subj_EPT)
Subj_FCT <- rbind(SDB_Subj_FCT, DLI78_Subj_FCT, MLS78_Subj_FCT, DLI5_Subj_FCT, MLS5_Subj_FCT)
```

```{r, generate-DOM-averages}
## SDB, SCT
SDB_DOM_EPT_Structure <- aggregate(SDB_DOM_EPT$DOM_Use, list(SDB_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_DOM_EPT_Structure <- SDB_DOM_EPT_Structure %>% rename(Structure_Avg = x)
SDB_DOM_EPT_Structure <- left_join(SDB_DOM_EPT, SDB_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## SDB, MST
SDB_DOM_FCT_Structure <- aggregate(SDB_DOM_FCT$DOM_Use, list(SDB_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_DOM_FCT_Structure <- SDB_DOM_FCT_Structure %>% rename(Structure_Avg = x)
SDB_DOM_FCT_Structure <- left_join(SDB_DOM_FCT, SDB_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, SCT
DLI78_DOM_EPT_Structure <- aggregate(DLI78_DOM_EPT$DOM_Use, list(DLI78_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI78_DOM_EPT_Structure <- DLI78_DOM_EPT_Structure %>% rename(Structure_Avg = x)
DLI78_DOM_EPT_Structure <- left_join(DLI78_DOM_EPT, DLI78_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, MST
DLI78_DOM_FCT_Structure <- aggregate(DLI78_DOM_FCT$DOM_Use, list(DLI78_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI78_DOM_FCT_Structure <- DLI78_DOM_FCT_Structure %>% rename(Structure_Avg = x)
DLI78_DOM_FCT_Structure <- left_join(DLI78_DOM_FCT, DLI78_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, SCT
MLS78_DOM_EPT_Structure <- aggregate(MLS78_DOM_EPT$DOM_Use, list(MLS78_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS78_DOM_EPT_Structure <- MLS78_DOM_EPT_Structure %>% rename(Structure_Avg = x)
MLS78_DOM_EPT_Structure <- left_join(MLS78_DOM_EPT, MLS78_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, MST
MLS78_DOM_FCT_Structure <- aggregate(MLS78_DOM_FCT$DOM_Use, list(MLS78_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS78_DOM_FCT_Structure <- MLS78_DOM_FCT_Structure %>% rename(Structure_Avg = x)
MLS78_DOM_FCT_Structure <- left_join(MLS78_DOM_FCT, MLS78_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, SCT
DLI5_DOM_EPT_Structure <- aggregate(DLI5_DOM_EPT$DOM_Use, list(DLI5_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI5_DOM_EPT_Structure <- DLI5_DOM_EPT_Structure %>% rename(Structure_Avg = x)
DLI5_DOM_EPT_Structure <- left_join(DLI5_DOM_EPT, DLI5_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, MST
DLI5_DOM_FCT_Structure <- aggregate(DLI5_DOM_FCT$DOM_Use, list(DLI5_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI5_DOM_FCT_Structure <- DLI5_DOM_FCT_Structure %>% rename(Structure_Avg = x)
DLI5_DOM_FCT_Structure <- left_join(DLI5_DOM_FCT, DLI5_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, SCT
MLS5_DOM_EPT_Structure <- aggregate(MLS5_DOM_EPT$DOM_Use, list(MLS5_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS5_DOM_EPT_Structure <- MLS5_DOM_EPT_Structure %>% rename(Structure_Avg = x)
MLS5_DOM_EPT_Structure <- left_join(MLS5_DOM_EPT, MLS5_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, MST
MLS5_DOM_FCT_Structure <- aggregate(MLS5_DOM_FCT$DOM_Use, list(MLS5_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS5_DOM_FCT_Structure <- MLS5_DOM_FCT_Structure %>% rename(Structure_Avg = x)
MLS5_DOM_FCT_Structure <- left_join(MLS5_DOM_FCT, MLS5_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## Join datasets
DOM_EPT_Final <- rbind(SDB_DOM_EPT_Structure, DLI78_DOM_EPT_Structure, MLS78_DOM_EPT_Structure, DLI5_DOM_EPT_Structure, MLS5_DOM_EPT_Structure) %>% 
  mutate(Structure = "DOM")

DOM_FCT_Final <- rbind(SDB_DOM_FCT_Structure, DLI78_DOM_FCT_Structure, MLS78_DOM_FCT_Structure, DLI5_DOM_FCT_Structure, MLS5_DOM_FCT_Structure) %>% 
  mutate(Structure = "DOM")
```

```{r, prepare-subj-averages}
## SDB, SCT
SDB_Subj_EPT_Structure <- aggregate(SDB_Subj_EPT$Mood_Use, list(SDB_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_EPT_Structure <- SDB_Subj_EPT_Structure %>% rename(Structure_Avg = x)
SDB_Subj_EPT_Structure <- left_join(SDB_Subj_EPT, SDB_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## SDB, MST
SDB_Subj_FCT_Structure <- aggregate(SDB_Subj_FCT$Mood_Use, list(SDB_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_FCT_Structure <- SDB_Subj_FCT_Structure %>% rename(Structure_Avg = x)
SDB_Subj_FCT_Structure <- left_join(SDB_Subj_FCT, SDB_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, SCT
DLI78_Subj_EPT_Structure <- aggregate(DLI78_Subj_EPT$Mood_Use, list(DLI78_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_EPT_Structure <- DLI78_Subj_EPT_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_EPT_Structure <- left_join(DLI78_Subj_EPT, DLI78_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, MST
DLI78_Subj_FCT_Structure <- aggregate(DLI78_Subj_FCT$Mood_Use, list(DLI78_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_FCT_Structure <- DLI78_Subj_FCT_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_FCT_Structure <- left_join(DLI78_Subj_FCT, DLI78_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, SCT
MLS78_Subj_EPT_Structure <- aggregate(MLS78_Subj_EPT$Mood_Use, list(MLS78_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_EPT_Structure <- MLS78_Subj_EPT_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_EPT_Structure <- left_join(MLS78_Subj_EPT, MLS78_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, MST
MLS78_Subj_FCT_Structure <- aggregate(MLS78_Subj_FCT$Mood_Use, list(MLS78_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_FCT_Structure <- MLS78_Subj_FCT_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_FCT_Structure <- left_join(MLS78_Subj_FCT, MLS78_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, SCT
DLI5_Subj_EPT_Structure <- aggregate(DLI5_Subj_EPT$Mood_Use, list(DLI5_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_EPT_Structure <- DLI5_Subj_EPT_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_EPT_Structure <- left_join(DLI5_Subj_EPT, DLI5_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, MST
DLI5_Subj_FCT_Structure <- aggregate(DLI5_Subj_FCT$Mood_Use, list(DLI5_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_FCT_Structure <- DLI5_Subj_FCT_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_FCT_Structure <- left_join(DLI5_Subj_FCT, DLI5_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, SCT
MLS5_Subj_EPT_Structure <- aggregate(MLS5_Subj_EPT$Mood_Use, list(MLS5_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_EPT_Structure <- MLS5_Subj_EPT_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_EPT_Structure <- left_join(MLS5_Subj_EPT, MLS5_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, MST
MLS5_Subj_FCT_Structure <- aggregate(MLS5_Subj_FCT$Mood_Use, list(MLS5_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_FCT_Structure <- MLS5_Subj_FCT_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_FCT_Structure <- left_join(MLS5_Subj_FCT, MLS5_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## Join datasets
Subj_EPT_Final <- rbind(SDB_Subj_EPT_Structure, DLI78_Subj_EPT_Structure, MLS78_Subj_EPT_Structure, DLI5_Subj_EPT_Structure, MLS5_Subj_EPT_Structure) %>% 
  mutate(Structure = "Subjunctive")

Subj_FCT_Final <- rbind(SDB_Subj_FCT_Structure, DLI78_Subj_FCT_Structure, MLS78_Subj_FCT_Structure, DLI5_Subj_FCT_Structure, MLS5_Subj_FCT_Structure) %>% 
  mutate(Structure = "Subjunctive")
```

```{r, generate-DOM-master}
## Calculate participant averages
DOM_EPT_Part_Avg <- aggregate(DOM_EPT_Final$DOM_Use, list(DOM_EPT_Final$Part_ID), FUN = sum, na.rm = TRUE)
DOM_EPT_Part_Avg <- DOM_EPT_Part_Avg %>% rename(Part_Avg = x)
DOM_EPT_Part_Avg <- left_join(DOM_EPT_Final, DOM_EPT_Part_Avg, by = c("Part_ID" = "Group.1"))

DOM_FCT_Part_Avg <- aggregate(DOM_FCT_Final$DOM_Use, list(DOM_FCT_Final$Part_ID), FUN = sum, na.rm = TRUE)
DOM_FCT_Part_Avg <- DOM_FCT_Part_Avg %>% rename(Part_Avg = x)
DOM_FCT_Part_Avg <- left_join(DOM_FCT_Final, DOM_FCT_Part_Avg, by = c("Part_ID" = "Group.1"))


## Form overall master
DOM_Master <- rbind(DOM_EPT_Part_Avg, DOM_FCT_Part_Avg)
DOM_Master$Group <- factor(DOM_Master$Group, levels = c("MLS-5", "DLI-5", "MLS-7/8", "DLI-7/8", "SDBA"))
DOM_Master$Task <- factor(DOM_Master$Task, levels = c("SCT", "MST"))


## Calculate HS master
DOM_Master_HS <- DOM_Master %>% 
  filter(!Group == "SDBA")
DOM_Master_HS$Group <- factor(DOM_Master_HS$Group, levels = c("MLS-5", "DLI-5", "MLS-7/8", "DLI-7/8"))
DOM_Master_HS$Task <- factor(DOM_Master_HS$Task, levels = c("SCT", "MST"))
DOM_Master_HS$Age <- factor(DOM_Master_HS$Age, levels = c("5th", "7th/8th"))
```

```{r, generate-subj-master}
## Calculate participant averages
Subj_EPT_Part_Avg <- aggregate(Subj_EPT_Final$Mood_Use, list(Subj_EPT_Final$Part_ID), FUN = sum, na.rm = TRUE)
Subj_EPT_Part_Avg <- Subj_EPT_Part_Avg %>% rename(Part_Avg = x)
Subj_EPT_Part_Avg <- left_join(Subj_EPT_Final, Subj_EPT_Part_Avg, by = c("Part_ID" = "Group.1"))

Subj_FCT_Part_Avg <- aggregate(Subj_FCT_Final$Mood_Use, list(Subj_FCT_Final$Part_ID), FUN = sum, na.rm = TRUE)
Subj_FCT_Part_Avg <- Subj_FCT_Part_Avg %>% rename(Part_Avg = x)
Subj_FCT_Part_Avg <- left_join(Subj_FCT_Final, Subj_FCT_Part_Avg, by = c("Part_ID" = "Group.1"))


## Form overall master
Subj_Master <- rbind(Subj_EPT_Part_Avg, Subj_FCT_Part_Avg)
Subj_Master$Group <- factor(Subj_Master$Group, levels = c("MLS-5", "DLI-5", "MLS-7/8", "DLI-7/8", "SDBA"))
Subj_Master$Task <- factor(Subj_Master$Task, levels = c("SCT", "MST"))


## Form HS master
Subj_Master_HS <- Subj_Master %>% 
  filter(!Group == "SDBA")
Subj_Master_HS$Group <- factor(Subj_Master_HS$Group, levels = c("MLS-5", "DLI-5", "MLS-7/8", "DLI-7/8"))
Subj_Master_HS$Task <- factor(Subj_Master_HS$Task, levels = c("SCT", "MST"))
Subj_Master_HS$Age <- factor(Subj_Master_HS$Age, levels = c("5th", "7th/8th"))
```

```{r, create-aggregate}
Aggregate <- rbind(DOM_Master, Subj_Master)
```

```{r, create-task-specific-dataframes}
DOM_EPT_HS <- DOM_Master_HS %>% 
  filter(Task == "SCT")

DOM_FCT_HS <- DOM_Master_HS %>% 
  filter(Task == "MST")

Subj_EPT_HS <- Subj_Master_HS %>% 
  filter(Task == "SCT")

Subj_FCT_HS <- Subj_Master_HS %>% 
  filter(Task == "MST")
```

# Abstract


Research on the grammars of heritage speakers of Spanish has become extensive in recent years, but the systematic exploration of how dual-language immersion influences Spanish-speaking children’s development of their heritage language is lacking. The present study presents data assessing middle school Spanish-speaking children’s knowledge of subjunctive mood and differential object marking by comparing participants in a dual-language school with those in a monolingual school. These two grammatical structures have been identified as challenging to acquire in extensive previous heritage language research with adults. Results from two experimental tasks (one testing production and one testing written receptive knowledge) show that children in the dual-language immersion program have an advantage in their knowledge of differential object marking over monolingually-educated peers, but this advantage does not extend to the subjunctive mood. Implications for teaching Spanish in dual-language immersion programs are discussed.

---
# Reviewing Dual-Language Immersion

--
- **Dual-language immersion (DLI)** is an equitable way to educate bilingual and bicultural children in U.S. public schools

--
  + English is taught for a portion of the school day

--
  + Spanish or a different language is taught for a different portion of the school day


--
- This is an inclusive method of education because:

--
  + Everyone goes through the process of learning a language
  
--
  + Languages that often have less social power are given equal status (in theory)

--
  + Children who come in with some proficiency in both languages are the "gold standard" (bilingualism as the resoltuion, not the problem)
  

---
# Exposing the Gap


--
There is a large body of research on dual-language immersion that shows benefits of many different sorts:


--
- Faster executive functioning skills (Nicolay & Poncelet, 2013; Garraffa et al., 2020)


--
- Greater sociocultural competence (Bearse & de Jong, 2008)


--
- Higher or equal standardized test scores in mathematics (Marian et al., 2013; Steele et al., 2017)


--
- Greater motivations to stay in school (Christian et al., 2004)


--
- Higher grade point averages (Thomas & Collier, 2002)


--
- Grade-level (or above grade-level) English language arts performance by the end of elementary school (Christian et al., 2004; Marian et al., 2013)


---
```{r, include-elephant, height = 10, fig.width = 16, fig.align = 'center', fig.retina = 2}
Elephant <- "http://www.elephantcapability.com/wp-content/uploads/2015/11/AdobeStock_94910721.jpeg"

knitr::include_graphics(Elephant)
```


---
# Revealing the Paradox


--
Heritage Spanish-speaking children…also continue developing their Spanish proficiency, particularly more formal registers, an opportunity not offered by the vast majority of United States elementary schools.


(Potowski, 2007b, p. 188)

--

**…but…**


--
"What about…a study that compares L1 heritage speakers in a dual immersion program with their counterparts in transitional bilingual education programs taught mostly in English?"


(Potowski, 2007a, p. 164).


--
**…so…**
--


More needs to be done to adequately document the impact of DLI on heritage language acquisition.


---
# A Role for Education?


--
- Bayram et al. (2017): literacy level affects children's knowledge of Turkish passives


--
- Kupisch & Rothman (2018): French HS educated in French in Germany > Italian HS in Germany


--
- Barnett et al. (2007): Spanish-speaking preschoolers in DLI have higher vocabulary scores than in monolingual programs


--
- To date, no systematic study comparing heritage speakers of Spanish in immersion programs to monolingually-educated HSs *during the school years*


--
- How does language exposure outside of school interact with the role of schooling? (Individual differences)


---
# The Role of Age in Previous Research


- Typically, language development increases with age


--
  + Monolingual children tend to master *most* aspects of their language by the school period
  
--
  + Bilingual children may need longer, but we still hope for an *upward* trajectory


--
- Some studies reveal *negative* age effects in HS children (e.g., Goebel-Mahrle & Shin, 2021; Merino, 1983)


--
- Other studies reveal *positive* age effects (e.g., Corbet & Domínguez, 2020; Cuza & Miller, 2015; Martinez Nieto & Restrepo, 2022; Montrul & Potowski, 2007)


--
- Many studies reveal no effect for age (e.g., Dracos & Requena, 2022; Guijarro-Fuentes et al., 2017)


--
- Do overall patterns of exposure to the language matter?


--
- Does age interact with education (e.g., education modulates the effect of age)?


---
# Structures Tested


--
- Volitional subjunctive:

--
  + Example: *quiero que VENGAS a la fiesta* (versus *quiero que VIENES a la fiesta*)

--
  + Late-acquired by monolingual Spanish-speaking children (Dracos et al., 2019)

--
  + Variable in child/adult HS (e.g., Dracos & Requena, 2022; Montrul, 2009)


--
- Differential object marking (DOM) or "personal *a*"

--
  + Example: mira Juan (Juan sees) versus mira *a* Juan (she sees Juan)

--
  + Stabilizes in monolingual acquisition by approx. age 3 (Rodríguez-Mondoñedo, 2008)
  
--
  + Variable in HS; adults > children (Montrul & Sánchez-Walker, 2013)

--
  + This structure can impact on meaning in Spanish


---

# Research Questions


--
**1. Do bilingual education and frequency of use of Spanish affect Spanish HS children's development of the subjunctive and differential object marking?**

--
  + Students in DLI will use the subjunctive and DOM more frequently than peers in monolingual schools.

--
  + Children who report using Spanish across contexts more frequently will use more subjunctive and DOM than peers.


--
**2. Do Spanish HS children recognize subjunctive and DOM on a receptive, comprehension-like task more than they produce them?**

--
  + Participants will recognize these structures more on a receptive task than they will produce them, because production is more challenging than comprehension (Perez-Cortes et al., 2019)


--
**3. Do Spanish HS children show improvement in their knowledge of subjunctive and differential object marking as they get older?**

--
  + Adolescent Spanish HSs will produce and select more subjunctive and DOM than younger children.


---
# Participants


--
- 17 Spanish-dominant bilingual adults (SBDA) raised in typically monolingual Spanish-speaking regions (descriptive stats only)


--
- 57 Spanish HS children (all children in two groups by age and school exposure); matched for SES

--
  + DLI-7/8: DLI students in 7th/8th grades (12-14 years old); 50% in Spanish through grade 5

--
  + MLS-7/8: students in a monolingual school in 7th/8th grades (12-14 years old); no Spanish at school

--
 + DLI-5: DLI students in 5th grade (10-11 years old); 50% in Spanish since Kindergarten

--
  + MLS-5: students in a monolingual school in 5th grade (11-12 years old); no Spanish at school


--
```{r, load-partipant-table}
Participants <- read_csv(here("./PoLaR/PoLaR Participant Table.csv"))
```

```{r, list-participant-table}
kable(Participants[1:4], align = "lccccc")
```

---
# Methods


--
- BESA proficiency test for bilingual children (Peña et al., 2014)


--
- Language questionnaire targeting self-reported frequency of use


--
  + 6 contexts of exposure

--
  + 1-5 Likert scales


--
- Productive and receptive domains tested:

--
  + Production: sentence completion task (SCT)

--
  + Receptive knowledge: morphology selection task (MST) between grammatical and non-grammatical options


---
# Production and Selection Rates
```{r, generate-DOM-bar-graph}
DOM_Bar_Plot <- DOM_Master %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
    facet_wrap(facets = vars(Structure)) +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
    geom_text(aes(label = round(Structure_Percentage)),
            position = position_dodge(width = .9),
            vjust = -0.5,
            size = 6) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of DOM responses") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, generate-subj-bar-graph}
Subj_Bar_Plot <- Subj_Master %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
    facet_wrap(facets = vars(Structure)) +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
    geom_text(aes(label = round(Structure_Percentage)),
            position = position_dodge(width = .9),
            vjust = -0.5,
            size = 6) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of subjunctive responses") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, print-DOM-bar-graph}
(DOM_Bar_Plot)+(Subj_Bar_Plot)
```


---
# DOM: Correlation

```{r, generate-DOM-overall-correlation}
DOM_Overall <- glmer(DOM_Use ~ School + Age + Task + Use_Joint_Std + School:Age + Task:Use_Joint_Std +
                       (1 | Part_ID) + (1 | Item),
                     data = DOM_Master_HS,
                     family = "binomial")
```

```{r, generate-DOM-overall-forest-plot}
DOM_Overall_Plot <- plot_model(DOM_Overall, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
  labs(title = "DOM Overall Model", y = "Estimates") +
                  scale_y_continuous(breaks = seq(-4, 4, 2),
                     limits = c(-4, 4)) +
              scale_x_discrete(labels = c("MST : Frequency of use", "MLS : 7th/8th grade", "Frequency of use", "Morph. Selection Task (MST)", "7th/8th grade age group", "Monolingual school (MLS) group", "(Intercept)")) +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-DOM-overall-forest-plot}
(DOM_Overall_Plot)
```


---
# DOM: Within-Group Comparisons


--
Group comparisons fitted through Tukey post-hoc comparisons with 4-way categorization collapsing school and age groups


--
```{r, generate-DOM-group-comparisons}
DOM_Group <- glmer(DOM_Use ~ Group +
                     (1 | Part_ID) + (1 | Item),
                   data = DOM_Master_HS,
                   family = "binomial")
```

```{r, generate-DOM-tukey}
DOM_Pairwise <- emmeans(DOM_Group, spec = "Group")
DOM_Tukey <- contrast(DOM_Pairwise, method = "pairwise")
DOM_Table <- summary(DOM_Tukey)
```

```{r, show-DOM-Tukey-results}
kable(DOM_Table[1:6], col.names = c("Contrast", "Estimate", "SE", "df", "t ratio", "p"), align = "lccccc")
```


--
- The following differences are significant at *p* > .05 level:


--
  + Between the DLI-7/8 and MLS-7/8 groups (the bilingually-educated children were more likely to use the subjunctive)

--
  + Between the DLI-7/8 and MLS-5 groups (older DLI students > younger, monolingually-educated students)


---
# DOM: Individual Analyses

```{r, calculate-DOM-ind-diffs}
DOM_EPT_Diffs <- DOM_EPT %>%
  filter(!is.na(DOM_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(DOM_Use))

DOM_FCT_Diffs <- DOM_FCT %>%
  filter(!is.na(DOM_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(DOM_Use))

DOM_Diffs = left_join(DOM_EPT_Diffs, DOM_FCT_Diffs, by = "Part_ID", "Group") %>%
  rename(Group = Group.x,
         Production = Production.x,
         Preference = Production.y) %>% 
  mutate(Total = Preference + Production) %>% 
  filter(!Group == "SDBA")

DOM_Diffs$Group <- factor(DOM_Diffs$Group, levels = c("MLS-5", "DLI-5", "MLS-7/8", "DLI-7/8", "SDBA"))
```

```{r, generate-DOM-diffs-plot}
DOM_Diffs_Plot <- DOM_Diffs %>% 
  ggplot(aes(x = Production, y = Preference)) +
  geom_jitter(mapping = aes(color = Group), size = 3) +
  scale_x_continuous(breaks = seq (0, 10, 2),
                     limits = c(-1, 11)) +
  scale_y_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  labs(x = "Number of sentences with DOM produced", y = "Number of sentences with DOM selected", title = "Individual Rates of DOM Selection and Production", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, print-DOM-diffs-plot}
(DOM_Diffs_Plot)
```


---
# DOM: Summary


--
- Age, but not bilingual education or their interaction, is significant at the *p* < .05 level


--
- Possible interaction between age and DLI work together because:

--
  + DLI-7/8 were greater than both monolingually-educated groups, but not different

--
  + DLI-7/8 were not different from DLI-5 (same school, different age)

--
  + DLI-5 were not different from MLS-5 (same age, different school)


--
- Heritage speakers recognized DOM on the MST (receptive task) more than they produced it on the SCT


--
- High degrees of individual variability; nobody *never* used DOM, and some were just like Spanish-dominant speakers


---
# Subjunctive: Correlation

```{r, generate-subj-overall-correlation}
Subj_Overall <- glmer(Mood_Use ~ School + Age + Task + Use_Joint_Std + School:Age + Task:Use_Joint_Std +
                       (1 | Part_ID) + (1 | Item),
                     data = Subj_Master_HS,
                     family = "binomial")
```

```{r, generate-subj-overall-forest-plot}
Subj_Overall_Plot <- plot_model(Subj_Overall, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
                  scale_y_continuous(breaks = seq(-4, 4, 2),
                     limits = c(-4, 4)) +
              scale_x_discrete(labels = c("MST : Frequency of use", "MLS : 7th/8th grade", "Frequency of use", "Morph. Selection Task (MST)", "7th/8th grade age group", "Monolingual school (MLS) group", "(Intercept)")) +
  labs(title = "Subjunctive Overall Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-Subj-overall-forest-plot}
(Subj_Overall_Plot)
```


---
# Subjunctive: Group Differences

```{r, generate-subj-group-comparisons}
Subj_Group <- glmer(Mood_Use ~ Group +
                     (1 | Part_ID) + (1 | Item),
                   data = Subj_Master_HS,
                   family = "binomial")
```

```{r, generate-subj-tukey}
Subj_Pairwise <- emmeans(Subj_Group, spec = "Group")
Subj_Tukey <- contrast(Subj_Pairwise, method = "pairwise")
Subj_Table <- summary(Subj_Tukey)
```

```{r, show-subj-Tukey-results}
kable(Subj_Table[1:6], col.names = c("Contrast", "Estimate", "SE", "df", "t ratio", "p"), align = "lccccc")
```

--
- No between-groups differences are significant at the *p* < .05 level


---
# Subjunctive: Individual Analyses

```{r, calculate-subj-ind-diffs}
Subj_EPT_Diffs <- Subj_EPT %>%
  filter(!is.na(Mood_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(Mood_Use))

Subj_FCT_Diffs <- Subj_FCT %>%
  filter(!is.na(Mood_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(Mood_Use))

Subj_Diffs = left_join(Subj_EPT_Diffs, Subj_FCT_Diffs, by = "Part_ID", "Group") %>%
  rename(Group = Group.x,
         Production = Production.x,
         Preference = Production.y) %>% 
  mutate(Total = Preference + Production) %>% 
  filter(!Group == "SDBA")

Subj_Diffs$Group <- factor(Subj_Diffs$Group, levels = c("MLS-5", "DLI-5", "MLS-7/8", "DLI-7/8", "SDBA"))
```

```{r, generate-subj-diffs-plot}
Subj_Diffs_Plot <- Subj_Diffs %>% 
  ggplot(aes(x = Production, y = Preference)) +
  geom_jitter(mapping = aes(color = Group), size = 3) +
  scale_x_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  scale_y_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  labs(x = "Number of sentences with subjunctive produced", y = "Number of sentences with subjunctive selected", title = "Individual Rates of Subjunctive Selection and Production", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, print-subj-diffs-plot}
(Subj_Diffs_Plot)
```

---
# Subjunctive: Summary


--
- Subjunctive knowledge increases with age, but not as a function of bilingual education


--
- No effect for education is found in the main model nor in the 4-way group comparison


--
- Children select more subjunctive on the MST than they produce it on the SCT


--
- Overall frequency of use matters in speakers' knowledge of this structure


--
- No child *never* uses the subjunctive, and some are within the range of the Spanish-dominant adults


---
# Research Question #1


--
**Do bilingual education and frequency of use of Spanish affect Spanish HS children's development of the subjunctive and differential object marking?**


--
- Prediction: both variables would affect children's knowledge of both structures


--
- Effect of bilingual education with older children in 4-way group analysis only


--
- Overall frequency of use subtly affected individual variability in subjunctive


--
- The hypothesis for RQ1 is partially supported and partially rejected


---
# Research Question #2


--
**Do Spanish HS children recognize subjunctive and DOM on a receptive, comprehension-like task more than they produce them?**


--
- Prediction: HS will use more DOM and subjunctive on a receptive task (MST) than in production (SCT)


--
- Results support this prediction, as well as current theories of HL acquisition (Putnam & Sánchez, 2013)


--
- Implication: students may understand more than they hear; **input is necessary but insufficient for language acquisition**

--
  + Long argued for second language acquisition (Swain, 1993, 1995)

--
  + Simply providing instruction in Spanish isn't enough!

--
  + We need children to be producing in immersion classrooms…


---
# Research Question #3


--
**Do Spanish HS children show improvement in their knowledge of subjunctive and differential object marking as they get older?**


--
- Prediction: HSs will show growth with age in their command of both DOM and subjunctive


--
- Age modulated rates of DOM and subjunctive production and selection


--
- Age *and* bilingual education may interact in the acquisition of DOM


--
- Hypothesis supported


---
# Implications


--
- Without creating specific efforts to integrate langauge and content, some of the more technical aspects of language might not develop "for free"

--
  + There is often not a systematic focus on language objectives across the content areas in DLI programs

--
  + Harley (1993): some explicit and some content-based instruction is necessary; they should go together

--
  + Babino & Stewart (2017), Potowski (2004): children often (or always) respond in English, even when they receive rich input in Spanish


--
- A space for translanguaging?

--
  + Translingual spaces are important and identity-validating spaces for bilingual children (Ibarra-Johnson et al., 2019)
  
--
  + Guerrero (2023): translanguaging may have negative implications in this instance (shift towards English)

--
  + Does translanguaging have detrimental consequences for the heritage langauge?

---
# Criticality and Dual-Language Immersion?


--
- These findings **do not** devalue the importance of DLI for multiple reasons


--
- Language may not come "for free"

--
  + Greater consideration of language learning objectives for all students

--
  + Sustained and intensive (90/10) DLI *across middle school*


--
- We want to find methods to support heritage language development across the childhood years in an *equitable* and *sustainable* way

--
  + Ibarra-Johnson et al. (2019): language separation with a dedicated translingual space

--
  + Urow & Beeman (2013): TfB and the Bridge; metalinguistic connections between languages and across content


---
# A Self-Critique


--
- Raciolinguistic ideologies (Rosa & Flores, 2017): is it truly important that students speak the way that we would expect them to speak?


--
  + A goal of DLI is to develop competence to interact with *a range of different speakers* (bilingual AND monolingual)

--
  + DOM has important implications for meaning in Spanish, but the subjunctive isn't as essential…


--
- Registers, metalinguistic awareness, and biliteracy development as important pieces of the puzzle


---
# Bridging the Gap

From Glenn Martínez (2003, p. 53)


--
**If our students walk into the class saying haiga and walk out saying haya, there has been, in my estimation, no value added. However, if they walk in saying haiga and walk out saying either haya or haiga and having the ability to defend their use of haiga if and when they see fit, then there has been value added. It is critical that we strive to allow students to develop this type of sociolinguistic sophistication in our endeavors as SHL [Spanish as a heritage language] educators.**


---
# References


Babino, A., & Stewart, M. A. (2017). “I like English better”: Latino dual language students’ investment in Spanish, English, and bilingualism. *Journal of Latinos and Education*, *16*(1), 18–29. https://doi.org/10.1080/15348431.2016.1179186

Bearse, C., & de Jong, E. (2008). Cultural and linguistic investment: Adolescents in a secondary two-way immersion program. *Equity & Excellence in Education*, *41*(3), 325–340. https://doi.org/10.1080/10665680802174817

Castilla-Earls, A., Francis, D., Iglesias, A., & Davidson, K. (2019). The impact of the Spanish-to-English proficiency shift on the grammaticality of English learners. Journal of Speech, Language, and Hearing Research, 62(6), 1739–1754. https://doi.org/10.1044/2018_JSLHR-L-18-0324

Christian, D., Genesee, F., Lindholm-Leary, K., & Howard, E. R. (2004). *Final progress report: CAL/CREDE study of two-way immersion education*. Center for Applied Linguistics. http://www.cal.org/twi/CREDEfinal.doc

Corbet, J., & Domínguez, L. (2020). The comprehension of tense–aspect morphology by Spanish heritage speakers in the United Kingdom. *Languages*, *5*(46), 1–20. https://doi.org/10.3390/languages5040046

Dracos, M., & Requena, P. E. (2022). Child heritage speakers’ acquisition of the Spanish subjunctive in volitional and adverbial clauses. *Language Acquisition*, *30*(1), 1–28. https://doi.org/10.1080/10489223.2022.2071156


---
# References


Dracos, M., Requena, P., & Miller, K. (2019). Acquisition of mood selection in Spanish-speaking children. *Language Acquisition*, *26*(1), 106–118. https://doi.org/10.1080/10489223.2018.1464006

Goebel-Mahrle, T., & Shin, N. L. (2020). A corpus study of child heritage speakers’ Spanish gender agreement. *International Journal of Bilingualism*, *24*(5–6), 1088–1104. https://doi.org/10.1177/1367006920935510

Guerrero, M. (2023). Gauging the adequacy of translanguaging allocation policy in two-way immersion programs in the U.S. *Journal of Latinos and Education*, *22*(4), 1427–1441. https://doi.org/10.1080/15348431.2021.1971086

Guijarro-Fuentes, P., Pires, A., & Nediger, W. (2017). Delay in the acquisition of Differential Object Marking by Spanish monolingual and bilingual teenagers. *International Journal of Bilingualism*, *21*(2), 159–177. https://doi.org/10.1177/1367006915601249

Harley, B. (1993). Instructional strategies and SLA in early French immersion. *Studies in Second Language Acquisition*, *15*(2), 245–259. https://doi.org/10.1017/S0272263100011980

Ibarra Johnson, S., García, O., & Seltzer, K. (2019). Biliteracy and translanguaging in dual-language bilingual education. In D. E. DeMatthews & E. Izquierdo (Eds.), *Dual language education: Teaching and leading in two languages* (pp. 119–132). Springer New York. https://doi.org/10.1007/978-3-030-10831-1_8

---
# References


Marian, V., Shook, A., & Schroeder, S. R. (2013). Bilingual two-way immersion programs benefit academic achievement. *Bilingual Research Journal*, *36*(2), 167–186. https://doi.org/10.1080/15235882.2013.818075

Martínez, G. (2003). Classroom based dialect awareness in heritage language instruction: A critical applied linguistic approach. *Heritage Language Journal*, *1*(1), 44–57. https://doi.org/10.46538/hlj.1.1.3

Martinez-Nieto, L., & Restrepo, M. A. (2022). Production and comprehension of grammatical gender by Spanish heritage speakers: Evidence from accusative clitic pronouns. *International Journal of Bilingualism*, 136700692110573. https://doi.org/10.1177/13670069211057318

Merino, B. J. (1983). Language loss in bilingual Chicano children. *Journal of Applied Developmental Psychology*, *4*(3), 277–294. https://doi.org/10.1016/0193-3973(83)90023-0

Montrul, S., & Potowski, K. (2007). Command of gender agreement in school-age Spanish-English bilingual children. *International Journal of Bilingualism*, *11*(3), 301–328. https://doi.org/10.1177/13670069070110030301

Montrul, S., & Sánchez-Walker, N. (2013). Differential object marking in child and adult Spanish heritage speakers. *Language Acquisition*, *20*(2), 109–132. https://doi.org/10.1080/10489223.2013.766741

---
# References


Peña, E. D., Gutiérrez-Clellen, V. F., Iglesias, A., Goldstein, B. A., & Bedore, L. M. (2014). *Bilingual English-Spanish Assessment (BESA)*. BROOKES Publishing CO. https://brookespublishing.com/wp-content/uploads/2021/01/BESA-excerpt.pdf

Perez-Cortes, S., Putnam, M., & Sánchez, L. (2019). Differential access: Asymmetries in accessing features and building representations in heritage language grammars. *Languages*, *4*(81), 1–27. https://doi.org/10.3390/languages4040081

Potowski, K. (2004). Student Spanish use and investment in a dual immersion classroom: Implications for second language acquisition and heritage language maintenance. *The Modern Language Journal*, *88*(1), 75–101. https://doi.org/10.1111/j.0026-7902.2004.00219.x

Potowski, K. (2007a). Characteristics of the Spanish grammar and sociolinguistic proficiency of dual immersion graduates. *Spanish in Context*, *4*(2), 187–216. https://doi.org/10.1075/sic.4.2.04pot

Potowski, K. (2007b). *Language and identity in a dual immersion school.* Multilingual Matters Limited.

Putnam, M. T., & Sánchez, L. (2013). What’s so incomplete about incomplete acquisition?: A prolegomenon to modeling heritage language grammars. *Linguistic Approaches to Bilingualism*, *3*(4), 478–508. https://doi.org/10.1075/lab.3.4.04put


---
# References


R Core Team. (2022). *R: A language and environment for statistical computing* [Computer software]. R Foundation for Statistical Computing. https://www.R-project.org/

Rosa, J., & Flores, N. (2017). Unsettling race and language: Toward a raciolinguistic perspective. *Language in Society*, *46*(5), 621–647. https://doi.org/10.1017/S0047404517000562

Steele, J. L., Slater, R. O., Zamarro, G., Miller, T., Li, J., Burkhauser, S., & Bacon, M. (2017). Effects of dual-language immersion programs on student achievement: Evidence From lottery data. *American Educational Research Journal*, *54*(1), 282S-306S. https://doi.org/10.3102/0002831216634463

Swain, M. (1993). The Output Hypothesis: Just speaking and writing aren’t enough. *The Canadian Modern Language Review*, *50*(1), 158–164. https://doi.org/10.3138/cmlr.50.1.158

Thomas, W. P., & Collier, V. P. (2002). *A national study of school effectiveness for language minority students’ long-term academic achievement.* Center for Research on Education, Diversity, and Excellence (CREDE). https://escholarship.org/uc/item/65j213pt

Urow, C., & Beeman, K. (2013). *Teaching for biliteracy: Strengthening bridges between languages*. Caslon Publishing.


---
# Sample Items from the Tasks


**Sample sentence from SCT**

A veces Juanito se pone triste si sus hermanas dicen que no quieren hablar con él. ¿Qué quiere la mamá? Quiere que las hermanas _________ (LLAMAR) Juanito cada noche.



**Sample sentences from MST**

La mamá sabe que Juanito tiene miedo de alturas. No quiere subir al tobogán en el parque de juegos solito. ¿Qué quiere la mamá?

  + Quiere que las hermanas lo lleven para subir al tobogán. 

  + *Quiere que las hermanas lo llevan para subir al tobogán. 


¿Qué tienen que hacer las hermanas? 
  + Tienen que llevar a Juanito para subir al tobogán. 

  + *Tienen que llevar Juanito para subir al tobogán
