---
title: "Education, Age, and Exposure: Conspiring Variables in the Acquisition of Spanish as a Heritage Language"
author: "Patrick D. Thane<br/>University of Massachusetts - Amherst"
institute: "Psycholinguistics of Language Representation (PoLaR) Lab"
date: "November 21, 2023"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r, setup, include = FALSE}
library(tidyverse)
library(lme4)
library(xaringan)
library(xaringanthemer)
library(xaringanExtra)
library(lmerTest)
library(kableExtra)
library(patchwork)
library(jtools)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(emmeans)

options(htmltools.dir.version = FALSE)
options(scipen = 999)


knitr::opts_chunk$set(fig.width = 16, fig.height = 7, fig.retina = 3, fig.align = 'center', cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, hiline = TRUE)
```

```{r xaringan, include = FALSE, warning = FALSE}
style_duo_accent(
  primary_color = "#881C1C",
  secondary_color = "#212721",
  inverse_header_color = "#FFFDD0")
```

```{r xaringan-logo, echo = FALSE}
use_logo(
  image_url = "https://www.umass.edu/brand/sites/default/files/M-lockedup.png"
)
```

```{r, load-DOM-data}
# EPT
SDB_DOM_EPT <- read_csv("../CSV Files/SDB/SDB DOM EPT.csv") %>% 
    filter(Item %in% c("EPT-02", "EPT-08", "EPT-12", "EPT-14", "EPT-20", "EPT-24", "EPT-26", "EPT-32", "EPT-38", "EPT-42")) %>% 
  mutate(Group = "SDBA",
         Task = "SCT",
         School = "SDBA",
         Age = "SDBA")

DLI78_DOM_EPT <- read.csv("../CSV Files/DLI-78/DLI-78 DOM EPT.csv") %>% 
  mutate(Group = "DLI-7/8",
         Task = "SCT",
         School = "DLI",
         Age = "7th/8th")

MLS78_DOM_EPT <- read.csv("../CSV Files/MLS-78/MLS-78 DOM EPT.csv") %>% 
  filter(!School == "GBCS") %>%
  mutate(Group = "MLS-7/8",
         Task = "SCT",
         School = "MLS",
         Age = "7th/8th")

DLI5_DOM_EPT <- read.csv("../CSV Files/DLI-5/DLI-5 DOM EPT.csv") %>% 
  mutate(Group = "DLI-5",
         Task = "SCT",
         School = "DLI",
         Age = "5th")

MLS5_DOM_EPT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM EPT.csv") %>% 
  filter(!School == "GBCS") %>% 
  mutate(Group = "MLS-5",
         Task = "SCT",
         School = "MLS",
         Age = "5th")


# FCT
SDB_DOM_FCT <- read_csv("../CSV Files/SDB/SDB DOM FCT.csv") %>% 
  mutate(Group = "SDBA",
         Task = "MST",
         School = "SDBA",
         Age = "SDBA")

DLI78_DOM_FCT <- read.csv("../CSV Files/DLI-78/DLI-78 DOM FCT.csv") %>% 
  mutate(Group = "DLI-7/8",
         Task = "MST",
         School = "DLI",
         Age = "7th/8th")

MLS78_DOM_FCT <- read.csv("../CSV Files/MLS-78/MLS-78 DOM FCT.csv") %>% 
  filter(!School == "GBCS") %>%
  mutate(Group = "MLS-7/8",
         Task = "MST",
         School = "MLS",
         Age = "7th/8th")

DLI5_DOM_FCT <- read.csv("../CSV Files/DLI-5/DLI-5 DOM FCT.csv") %>% 
  mutate(Group = "DLI-5",
         Task = "MST",
         School = "DLI",
         Age = "5th")

MLS5_DOM_FCT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM FCT.csv") %>% 
  filter(!School == "GBCS") %>% 
  mutate(Group = "MLS-5",
         Task = "MST",
         School = "MLS",
         Age = "5th")


## Join datasets
DOM_EPT <- rbind(SDB_DOM_EPT, DLI78_DOM_EPT, MLS78_DOM_EPT, DLI5_DOM_EPT, MLS5_DOM_EPT)
DOM_FCT <- rbind(SDB_DOM_FCT, DLI78_DOM_FCT, MLS78_DOM_FCT, DLI5_DOM_FCT, MLS5_DOM_FCT)
```

```{r, load-subj-data}
# EPT
SDB_Subj_EPT <- read_csv("../CSV Files/SDB/SDB Subjunctive EPT.csv") %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "SDBA",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "SDBA",
         Age = "SDBA")

DLI78_Subj_EPT <- read.csv("../CSV Files/DLI-78/DLI-78 Subjunctive EPT.csv") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-7/8",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "7th/8th")

MLS78_Subj_EPT <- read.csv("../CSV Files/MLS-78/MLS-78 Subjunctive EPT.csv") %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-7/8",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "7th/8th")

DLI5_Subj_EPT <- read.csv("../CSV Files/DLI-5/DLI-5 Subjunctive EPT.csv") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-5",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "5th")

MLS5_Subj_EPT <- read.csv("../CSV Files/MLS-5/MLS-5 Subjunctive EPT.csv") %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-5",
         Task = "SCT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "5th")


# FCT
SDB_Subj_FCT <- read_csv("../CSV Files/SDB/SDB Subjunctive FCT.csv") %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "SDBA",
         Task = "MST",
         Mood = "Subjunctive",
         School = "SDBA",
         Age = "SDBA")

DLI78_Subj_FCT <- read.csv("../CSV Files/DLI-78/DLI-78 Subjunctive FCT.csv") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-7/8",
         Task = "MST",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "7th/8th")

MLS78_Subj_FCT <- read.csv("../CSV Files/MLS-78/MLS-78 Subjunctive FCT.csv") %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-7/8",
         Task = "MST",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "7th/8th")

DLI5_Subj_FCT <- read.csv("../CSV Files/DLI-5/DLI-5 Subjunctive FCT.csv") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-5",
         Task = "MST",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "5th")

MLS5_Subj_FCT <- read.csv("../CSV Files/MLS-5/MLS-5 Subjunctive FCT.csv") %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-5",
         Task = "MST",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "5th")


## Join datasets
Subj_EPT <- rbind(SDB_Subj_EPT, DLI78_Subj_EPT, MLS78_Subj_EPT, DLI5_Subj_EPT, MLS5_Subj_EPT)
Subj_FCT <- rbind(SDB_Subj_FCT, DLI78_Subj_FCT, MLS78_Subj_FCT, DLI5_Subj_FCT, MLS5_Subj_FCT)
```

```{r, generate-DOM-averages}
## SDB, SCT
SDB_DOM_EPT_Structure <- aggregate(SDB_DOM_EPT$DOM_Use, list(SDB_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_DOM_EPT_Structure <- SDB_DOM_EPT_Structure %>% rename(Structure_Avg = x)
SDB_DOM_EPT_Structure <- left_join(SDB_DOM_EPT, SDB_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## SDB, MST
SDB_DOM_FCT_Structure <- aggregate(SDB_DOM_FCT$DOM_Use, list(SDB_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_DOM_FCT_Structure <- SDB_DOM_FCT_Structure %>% rename(Structure_Avg = x)
SDB_DOM_FCT_Structure <- left_join(SDB_DOM_FCT, SDB_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, SCT
DLI78_DOM_EPT_Structure <- aggregate(DLI78_DOM_EPT$DOM_Use, list(DLI78_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI78_DOM_EPT_Structure <- DLI78_DOM_EPT_Structure %>% rename(Structure_Avg = x)
DLI78_DOM_EPT_Structure <- left_join(DLI78_DOM_EPT, DLI78_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, MST
DLI78_DOM_FCT_Structure <- aggregate(DLI78_DOM_FCT$DOM_Use, list(DLI78_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI78_DOM_FCT_Structure <- DLI78_DOM_FCT_Structure %>% rename(Structure_Avg = x)
DLI78_DOM_FCT_Structure <- left_join(DLI78_DOM_FCT, DLI78_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, SCT
MLS78_DOM_EPT_Structure <- aggregate(MLS78_DOM_EPT$DOM_Use, list(MLS78_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS78_DOM_EPT_Structure <- MLS78_DOM_EPT_Structure %>% rename(Structure_Avg = x)
MLS78_DOM_EPT_Structure <- left_join(MLS78_DOM_EPT, MLS78_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, MST
MLS78_DOM_FCT_Structure <- aggregate(MLS78_DOM_FCT$DOM_Use, list(MLS78_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS78_DOM_FCT_Structure <- MLS78_DOM_FCT_Structure %>% rename(Structure_Avg = x)
MLS78_DOM_FCT_Structure <- left_join(MLS78_DOM_FCT, MLS78_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, SCT
DLI5_DOM_EPT_Structure <- aggregate(DLI5_DOM_EPT$DOM_Use, list(DLI5_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI5_DOM_EPT_Structure <- DLI5_DOM_EPT_Structure %>% rename(Structure_Avg = x)
DLI5_DOM_EPT_Structure <- left_join(DLI5_DOM_EPT, DLI5_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, MST
DLI5_DOM_FCT_Structure <- aggregate(DLI5_DOM_FCT$DOM_Use, list(DLI5_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI5_DOM_FCT_Structure <- DLI5_DOM_FCT_Structure %>% rename(Structure_Avg = x)
DLI5_DOM_FCT_Structure <- left_join(DLI5_DOM_FCT, DLI5_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, SCT
MLS5_DOM_EPT_Structure <- aggregate(MLS5_DOM_EPT$DOM_Use, list(MLS5_DOM_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS5_DOM_EPT_Structure <- MLS5_DOM_EPT_Structure %>% rename(Structure_Avg = x)
MLS5_DOM_EPT_Structure <- left_join(MLS5_DOM_EPT, MLS5_DOM_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, MST
MLS5_DOM_FCT_Structure <- aggregate(MLS5_DOM_FCT$DOM_Use, list(MLS5_DOM_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS5_DOM_FCT_Structure <- MLS5_DOM_FCT_Structure %>% rename(Structure_Avg = x)
MLS5_DOM_FCT_Structure <- left_join(MLS5_DOM_FCT, MLS5_DOM_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## Join datasets
DOM_EPT_Final <- rbind(SDB_DOM_EPT_Structure, DLI78_DOM_EPT_Structure, MLS78_DOM_EPT_Structure, DLI5_DOM_EPT_Structure, MLS5_DOM_EPT_Structure) %>% 
  mutate(Structure = "DOM")

DOM_FCT_Final <- rbind(SDB_DOM_FCT_Structure, DLI78_DOM_FCT_Structure, MLS78_DOM_FCT_Structure, DLI5_DOM_FCT_Structure, MLS5_DOM_FCT_Structure) %>% 
  mutate(Structure = "DOM")
```

```{r, prepare-subj-averages}
## SDB, SCT
SDB_Subj_EPT_Structure <- aggregate(SDB_Subj_EPT$Mood_Use, list(SDB_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_EPT_Structure <- SDB_Subj_EPT_Structure %>% rename(Structure_Avg = x)
SDB_Subj_EPT_Structure <- left_join(SDB_Subj_EPT, SDB_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## SDB, MST
SDB_Subj_FCT_Structure <- aggregate(SDB_Subj_FCT$Mood_Use, list(SDB_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_Subj_FCT_Structure <- SDB_Subj_FCT_Structure %>% rename(Structure_Avg = x)
SDB_Subj_FCT_Structure <- left_join(SDB_Subj_FCT, SDB_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, SCT
DLI78_Subj_EPT_Structure <- aggregate(DLI78_Subj_EPT$Mood_Use, list(DLI78_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_EPT_Structure <- DLI78_Subj_EPT_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_EPT_Structure <- left_join(DLI78_Subj_EPT, DLI78_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI78, MST
DLI78_Subj_FCT_Structure <- aggregate(DLI78_Subj_FCT$Mood_Use, list(DLI78_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI78_Subj_FCT_Structure <- DLI78_Subj_FCT_Structure %>% rename(Structure_Avg = x)
DLI78_Subj_FCT_Structure <- left_join(DLI78_Subj_FCT, DLI78_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, SCT
MLS78_Subj_EPT_Structure <- aggregate(MLS78_Subj_EPT$Mood_Use, list(MLS78_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_EPT_Structure <- MLS78_Subj_EPT_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_EPT_Structure <- left_join(MLS78_Subj_EPT, MLS78_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS78, MST
MLS78_Subj_FCT_Structure <- aggregate(MLS78_Subj_FCT$Mood_Use, list(MLS78_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS78_Subj_FCT_Structure <- MLS78_Subj_FCT_Structure %>% rename(Structure_Avg = x)
MLS78_Subj_FCT_Structure <- left_join(MLS78_Subj_FCT, MLS78_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, SCT
DLI5_Subj_EPT_Structure <- aggregate(DLI5_Subj_EPT$Mood_Use, list(DLI5_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_EPT_Structure <- DLI5_Subj_EPT_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_EPT_Structure <- left_join(DLI5_Subj_EPT, DLI5_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5, MST
DLI5_Subj_FCT_Structure <- aggregate(DLI5_Subj_FCT$Mood_Use, list(DLI5_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI5_Subj_FCT_Structure <- DLI5_Subj_FCT_Structure %>% rename(Structure_Avg = x)
DLI5_Subj_FCT_Structure <- left_join(DLI5_Subj_FCT, DLI5_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, SCT
MLS5_Subj_EPT_Structure <- aggregate(MLS5_Subj_EPT$Mood_Use, list(MLS5_Subj_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_EPT_Structure <- MLS5_Subj_EPT_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_EPT_Structure <- left_join(MLS5_Subj_EPT, MLS5_Subj_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5, MST
MLS5_Subj_FCT_Structure <- aggregate(MLS5_Subj_FCT$Mood_Use, list(MLS5_Subj_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS5_Subj_FCT_Structure <- MLS5_Subj_FCT_Structure %>% rename(Structure_Avg = x)
MLS5_Subj_FCT_Structure <- left_join(MLS5_Subj_FCT, MLS5_Subj_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## Join datasets
Subj_EPT_Final <- rbind(SDB_Subj_EPT_Structure, DLI78_Subj_EPT_Structure, MLS78_Subj_EPT_Structure, DLI5_Subj_EPT_Structure, MLS5_Subj_EPT_Structure) %>% 
  mutate(Structure = "Subjunctive")

Subj_FCT_Final <- rbind(SDB_Subj_FCT_Structure, DLI78_Subj_FCT_Structure, MLS78_Subj_FCT_Structure, DLI5_Subj_FCT_Structure, MLS5_Subj_FCT_Structure) %>% 
  mutate(Structure = "Subjunctive")
```

```{r, generate-DOM-master}
## Calculate participant averages
DOM_EPT_Part_Avg <- aggregate(DOM_EPT_Final$DOM_Use, list(DOM_EPT_Final$Part_ID), FUN = sum, na.rm = TRUE)
DOM_EPT_Part_Avg <- DOM_EPT_Part_Avg %>% rename(Part_Avg = x)
DOM_EPT_Part_Avg <- left_join(DOM_EPT_Final, DOM_EPT_Part_Avg, by = c("Part_ID" = "Group.1"))

DOM_FCT_Part_Avg <- aggregate(DOM_FCT_Final$DOM_Use, list(DOM_FCT_Final$Part_ID), FUN = sum, na.rm = TRUE)
DOM_FCT_Part_Avg <- DOM_FCT_Part_Avg %>% rename(Part_Avg = x)
DOM_FCT_Part_Avg <- left_join(DOM_FCT_Final, DOM_FCT_Part_Avg, by = c("Part_ID" = "Group.1"))


## Form overall master
DOM_Master <- rbind(DOM_EPT_Part_Avg, DOM_FCT_Part_Avg)
DOM_Master$Group <- factor(DOM_Master$Group, levels = c("SDBA", "DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
DOM_Master$Task <- factor(DOM_Master$Task, levels = c("SCT", "MST"))


## Calculate HS master
DOM_Master_HS <- DOM_Master %>% 
  filter(!Group == "SDBA")
DOM_Master_HS$Group <- factor(DOM_Master_HS$Group, levels = c("DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
DOM_Master_HS$Task <- factor(DOM_Master_HS$Task, levels = c("SCT", "MST"))
DOM_Master_HS$Age <- factor(DOM_Master_HS$Age, levels = c("5th", "7th/8th"))
```

```{r, generate-subj-master}
## Calculate participant averages
Subj_EPT_Part_Avg <- aggregate(Subj_EPT_Final$Mood_Use, list(Subj_EPT_Final$Part_ID), FUN = sum, na.rm = TRUE)
Subj_EPT_Part_Avg <- Subj_EPT_Part_Avg %>% rename(Part_Avg = x)
Subj_EPT_Part_Avg <- left_join(Subj_EPT_Final, Subj_EPT_Part_Avg, by = c("Part_ID" = "Group.1"))

Subj_FCT_Part_Avg <- aggregate(Subj_FCT_Final$Mood_Use, list(Subj_FCT_Final$Part_ID), FUN = sum, na.rm = TRUE)
Subj_FCT_Part_Avg <- Subj_FCT_Part_Avg %>% rename(Part_Avg = x)
Subj_FCT_Part_Avg <- left_join(Subj_FCT_Final, Subj_FCT_Part_Avg, by = c("Part_ID" = "Group.1"))


## Form overall master
Subj_Master <- rbind(Subj_EPT_Part_Avg, Subj_FCT_Part_Avg)
Subj_Master$Group <- factor(Subj_Master$Group, levels = c("SDBA", "DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
Subj_Master$Task <- factor(Subj_Master$Task, levels = c("SCT", "MST"))


## Form HS master
Subj_Master_HS <- Subj_Master %>% 
  filter(!Group == "SDBA")
Subj_Master_HS$Group <- factor(Subj_Master_HS$Group, levels = c("DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
Subj_Master_HS$Task <- factor(Subj_Master_HS$Task, levels = c("SCT", "MST"))
Subj_Master_HS$Age <- factor(Subj_Master_HS$Age, levels = c("5th", "7th/8th"))
```

```{r, create-aggregate}
Aggregate <- rbind(DOM_Master, Subj_Master)
```

```{r, create-task-specific-dataframes}
DOM_EPT_HS <- DOM_Master_HS %>% 
  filter(Task == "SCT")

DOM_FCT_HS <- DOM_Master_HS %>% 
  filter(Task == "MST")

Subj_EPT_HS <- Subj_Master_HS %>% 
  filter(Task == "SCT")

Subj_FCT_HS <- Subj_Master_HS %>% 
  filter(Task == "MST")
```

# Introduction


- Ph.D in Bilingualism and Second Language Acquisition, Rutgers University


--
- Researcher on heritage/second language acquisition of Spanish, particularly in children


--
- Postdoctoral Research Associate in Dual-Language Education, Equity, and Inclusion, College of Education, University of Massachusetts Amherst


--
- Dissertation on Spanish heritage language development in dual-language immersion schools


--
  + How do heritage speakers compare to one another under different contexts of schooling?

--
  + How do school-aged children contribute to current theoretical debates in HL acquisition more generally?


---

# Background

--
- School-aged children are the "missing link" (Montrul, 2018, p. 534) of HL acquisition research

--
  + What is the developmental trajectory between early childhood and adulthood?
  
--
  + What implications do data from these age ranges have for theories of HL acquisition?


--
- Product-oriented approaches to HL acquisition:

--
  + Incomplete acquisition: HL development "plateaus" and does not reach full competence, likely as a result of age of acquisition (Montrul, 2008, 2013)
  
--
  + Attrition: HS knowledge decreases over time due to lower exposure (e.g., Hicks & Domínguez, 2020)


--
- Process-oriented approach to HL acquisition:

--
  + HL knowledge is reassembled over time (Putnam & Sánchez, 2013)

--
  + Patterns of language use (activation) are necessary to conserve relationships between features

--
  + Decreasing activation leads to feature reassembly (following Lardiere, 2009)

--
  + Production is affected before receptive knowledge

---
# Revised Feature Reassembly Model

--
Quoted from Putnam & Sánchez (2013, pp. 489-490):


--
1. Transfer or re-assembly of some FFs (functional features) from the L2 grammar to L1 PF (phonological form) and semantic features which may coincide with the activation of L2 lexical items on a more frequent basis from the standpoint of linguistic production;


--
2. Transfer or re-assembly of massive sets of FFs from the L2 to L1 PF and semantic features, while concurrently showing significantly higher rates of activation of L2 lexical items than L1 lexical items for production purposes (i.e., they might code-switch more than bilinguals in the previous situation);


--
3. Exhibit difficulties in activating PF and semantic features (as well as other FFs) in the L1 for production purposes but are able to do so for comprehension of some high frequency lexical items; and,


--
4. Have difficulties activating PF features and semantic features (as well as other FFs) in the L1 for both production and comprehension purposes.


--
**What this implies is that this approach accounts for a *decrease* in use of Spanish (forms must have been acquired to be reassembled)**


---
# The Role of Age in Previous Research


--
- Putnam and Sánchez (2013) predict a reassembly of **HL** features to resemble the more-dominant **L2** (or **L1B**)


--
- To *re*assemble something, it must be "active" in the grammar in the first place

--
  + Some studies reveal *negative* age effects in HS children (e.g., Goebel-Mahrle & Shin, 2021; Merino, 1983)

--
  + Other studies reveal *positive* age effects (Corbet & Domínguez, 2020; Cuza & Miller, 2015; Martinez Nieto & Restrepo, 2022; Montrul & Potowski, 2007)


--
- Do children *re*assemble features, or do they exhibit protracted development?


---
# Pinning Down Activation


--
- Putnam and Sánchez's (2013) framework implies that *activation* affects command/maintenance of morphosyntactic structures

--
  + More-frequent use → more processing of input → stronger associations between FFs and other features
  
--
  + Less-frequent use → reassembly of features by those of the more-dominant language


--
- How to operationalize exposure

--
  + Patterns of language use on questionnaire (Dracos et al., 2019; Perez-Cortes, 2016)

--
  + Proficiency represents exposure to the HL (Giancaspro & Sánchez, 2021; López-Otero, 2022)

--
  + Age effects as a measure of *cumulative* exposure


--
- Proxies for the present study

--
  + Current exposure: proficiency and self-reported frequency of use
  
--
  + Cumulative exposure: age


---
# A Role for Education?


--
- Education in the heritage language likely provides greater *quantity* and *quality* of input for HSs


--
  + Quantitative component: more exposure across the school years, at a time when students begin to shift proficiency in English (Castilla-Earls et al., 2019)

--
  + Qualitative component: academic registers of the language likely expose learners to greater syntactic complexity and lexical diversity


--
- Previous evidence of educational impact on HL development:

--
  + Bayram et al. (2017): literacy level affects children's knowledge of Turkish passives

--
  + Kupisch & Rothman (2018): French HS educated in French in Germany > Italian HS in Germany

--
  + Barnett et al. (2007): Spanish-speaking preschoolers in DLI have higher vocabulary scores than in monolingual programs


--
- To date, no systematic study comparing heritage speakers of Spanish in immersion programs to monolingually-educated HSs *during the school years*


--
- Testable by evaluating HL development in **dual-language immersion (DLI)** schools in the U.S. that offer instruction in two languages

---
# Structures Tested

--
- Volitional subjunctive ("core" syntax through lexical selection)

--
  + Uninterpretable feature (instantiated differently from English mood)

--
  + Late-acquired in monolingual children (Dracos et al., 2019)

--
  + Variable in child/adult HS (e.g., Dracos & Requena, 2022; Montrul, 2009)

--
  + No age effects found in children (Dracos & Requena, 2022); underexplored in bilingual adolescents


--
- Differential object marking (syntax-semantics interface)

--
  + Interpretable feature (not present in English) used with [+animate] and [+specific] direct objects

--
  + Stabilizes in monolingual acquisition by approx. age 3 (Rodríguez-Mondoñedo, 2008)
  
--
  + Variable in HS; adults > children (Montrul & Sánchez-Walker, 2013)
  
--
  + No age effects found in young children (Cuza et al., 2019) or adolescents (Guijarro-Fuentes et al., 2017)


---
# "Competing" Features


--
- Comparing these features presents a unique opportunity regarding feature interpretability/interfaces:

--
  + Interpretability Hypothesis in SLA: uninterpretable, non-L1 features are not (fully) acquireable (Hawkins & Hattori, 2006; Tsimpli & Dimitrakopoulou, 2007)

--
  + Interface Hypothesis: core syntax > syntax-semantics > syntax-pragmatics (Tsimpli & Sorace, 2006)

--
  + Does feature interpretability or interface vulnerability win out?
  

--
- The FRH and the Interpretability Hypotheses may seem to exist in tension…


--
  + The FRH is a full access account

--
  + The Interpretability Hypothesis is a representational deficits account
  

--
- …But have been combined in past research on L2 acquisition of DOM (Guijarro-Fuentes, 2012)


--
- DOM: interpretable feature at the syntax-semantics interface


--
- Subjunctive: uninterpretable feature that is "interface-free"

---
# Research Questions


--
**1. Do frequency of use, morphosyntactic proficiency, DLI education affect individual HSs' knowledge of volitional subjunctive and DOM?**


--
- Greater use of subjunctive and DOM by individuals with more frequent use and more frequent of Spanish, and those who attend a DLI school


--
**2. Do school-aged Spanish HSs show increasing knowledge of volitional subjunctive and DOM with age?**


--
- Older HSs will produce/select more subjunctive mood than younger children


--
**3. Do school-aged Spanish HSs show asymmetrical productive and receptive knowledge of volitional subjunctive and DOM?**


--
- HSs will recognize both structures more frequently than they will produce them


--
**4. Does feature interpretability or interface variability have a greater effect on school-aged Spanish HSs knowledge of the structures tested?**


--
- If interpretability matters, DOM should be easier to acquire than volitional subjunctive

--
- If interface variability matters, volitional subjunctive should be easier to acquire than DOM


---
# Contributing to the Paradigm Shift


--
- Giancaspro et al. (2022): crossroads in HL research


--
  + Paradigm shift away from comparing HSs to other groups

--
  + Movement to understand the dynamic and pluralistic nature of HSs' competence


--
- None of the research questions concern between-group differences


--
- Shift in focus to questions concerning how to characterize variability between and within individuals


--
  + Between individuals: education, exposure, proficiency

--
  + Within-individuals: productive/receptive asymmetry, structure-to-structure comparisons

---
# Participants


- 17 Spanish-dominant bilingual adults (SBDA) raised in typically monolingual Spanish-speaking regions (descriptive stats only)


--
- 57 Spanish HS children (all children in two groups by age and school exposure); matched for SES

--
  + DLI-7/8: DLI students in 7th/8th grades (12-14 years old); 50% in Spanish through grade 5

--
  + MLS-7/8: students in a monolingual school in 7th/8th grades (12-14 years old); no Spanish at school

--
 + DLI-5: DLI students in 5th grade (10-11 years old); 50% in Spanish since Kindergarten

--
  + MLS-5: students in a monolingual school in 5th grade (11-12 years old); no Spanish at school


--
```{r, load-partipant-table}
Participants <- read_csv("../PoLaR/PoLaR Participant Table.csv")
```

```{r, list-participant-table}
kable(Participants[1:4], align = "lccccc")
```

---
# Methods


--
- BESA proficiency measure of morphosyntactic competence (Peña et al., 2014)


--
- Language questionnaire targeting self-reported frequency of use


--
  + 6 contexts of exposure

--
  + 1-5 Likert scales


--
- Productive and receptive domains tested:

--
  + Production: sentence completion task (SCT)

--
  + Receptive knowledge: morphology selection task (MST) between grammatical and non-grammatical options


---
# Production and Selection Rates
```{r, generate-DOM-bar-graph}
DOM_Bar_Plot <- DOM_Master %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
    facet_wrap(facets = vars(Structure)) +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
    geom_text(aes(label = round(Structure_Percentage)),
            position = position_dodge(width = .9),
            vjust = -0.5,
            size = 6) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of DOM responses") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, generate-subj-bar-graph}
Subj_Bar_Plot <- Subj_Master %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
    facet_wrap(facets = vars(Structure)) +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
    geom_text(aes(label = round(Structure_Percentage)),
            position = position_dodge(width = .9),
            vjust = -0.5,
            size = 6) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of subjunctive responses") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, print-DOM-bar-graph}
(DOM_Bar_Plot)+(Subj_Bar_Plot)
```


---
# Structure and Group Comparisons

```{r, create-structure-variable}
Subj_EPT_Combined <- Subj_EPT_HS %>% 
  mutate(Accuracy = Mood_Use,
         Structure = "Subjunctive")

Subj_FCT_Combined <- Subj_FCT_HS %>% 
  mutate(Accuracy = Mood_Use,
         Structure = "Subjunctive")

DOM_EPT_Combined <- DOM_EPT_HS %>% 
  mutate(Accuracy = DOM_Use,
         Structure = "DOM")

DOM_FCT_Combined <- DOM_FCT_HS %>% 
  mutate(Accuracy = DOM_Use,
         Structure = "DOM")

Combined <- rbind(Subj_EPT_Combined, Subj_FCT_Combined, DOM_EPT_Combined, DOM_FCT_Combined)
Combined$Group <- factor(Combined$Group, levels = c("DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
```

```{r, generate-combined-correlation}
Combined_Correlation <- glmer(Accuracy ~ Structure * Group +
                                (1 | Part_ID) + (1 | Item),
                              data = Combined,
                              family = "binomial")
```

```{r, generate-combined-correlation-forest-plot}
Combined_Plot <- plot_model(Combined_Correlation, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
    scale_y_continuous(breaks = seq(-4, 4, 2),
                     limits = c(-4, 4)) +
    scale_x_discrete(labels = c("Subjunctive : MLS-5", "Subjunctive : DLI-5", "Subjunctive : MLS-7/8", "MLS-5 group", "DLI-5 group", "MLS-7/8 group", "Subjunctive structure", "(Intercept)")) +
  labs(title = "Comparative Structures Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))

(Combined_Plot)
```

---
# Within-Group Comparisons

```{r, generate-tukey}
Combined_Pairwise <- emmeans(Combined_Correlation, spec = "Group")
Combined_Tukey <- contrast(Combined_Pairwise, method = "pairwise")
Tukey_Table <- summary(Combined_Tukey)
```

```{r, show-Tukey-results}
kable(Tukey_Table[1:6], col.names = c("Contrast", "Estimate", "SE", "df", "t ratio", "p"), align = "lccccc")
```

--
- Statistically significant differences at the *p* > .05 level


--
  + DLI-7/8 over DLI-5 (older immersion students stronger than younger immersion students)

--
  + DLI-7/8 over MLS-5 (contrasts in both age and educational experience)

---
# Structure and Group Data: Summary


--
- DLI-7/8 group (older participants with bilingual education) use DOM and the volitional subjunctive most out of the HS groups


--
- MLS-7/8 and MLS-5 groups are more likely to use the subjunctive than DOM


---
# DOM: Overall Correlation

```{r, generate-DOM-overall-correlation}
DOM_Overall <- glmer(DOM_Use ~ School + Age + Task + BESA_Total_Std + Use_Joint_Std + School:Age + Task:BESA_Total_Std + Task:Use_Joint_Std +
                       (1 | Part_ID) + (1 | Item),
                     data = DOM_Master_HS,
                     family = "binomial")
```

```{r, generate-DOM-overall-forest-plot}
DOM_Overall_Plot <- plot_model(DOM_Overall, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
      scale_x_discrete(labels = c("MST : Use", "MST : Proficiency", "MLS : 7th/8th", "Frequency of use", "Proficiency", "MST", "7th/8th grade group", "MLS group", "(Intercept)")) +
  labs(title = "DOM Overall Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-DOM-overall-forest-plot}
(DOM_Overall_Plot)
```


---
# DOM: Task-Specific Variability

```{r, generate-DOM-task-specific-correlations}
DOM_EPT_Correlation <- glmer(DOM_Use ~ School + Age + BESA_Other_Std + Use_Joint_Std + School:Age +
                               (1 | Part_ID) + (1 | Item),
                             data = DOM_EPT_HS,
                             family = "binomial")

DOM_FCT_Correlation <- glmer(DOM_Use ~ School + Age + BESA_Other_Std + Use_Joint_Std + School:Age +
                               (1 | Part_ID) + (1 | Item),
                             data = DOM_FCT_HS,
                             family = "binomial")
```

```{r, generate-DOM-EPT-forest-plot}
DOM_EPT_Plot <- plot_model(DOM_EPT_Correlation, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
        scale_y_continuous(breaks = seq(-6, 6, 2),
                     limits = c(-6, 6)) +
      scale_x_discrete(labels = c("MLS : 7th/8th", "Freq. of use", "Proficiency", "7th/8th grade", "MLS group", "(Intercept)")) +
  labs(title = "DOM Production Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, generate-DOM-FCT-forest-plot}
DOM_FCT_Plot <- plot_model(DOM_FCT_Correlation, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
          scale_y_continuous(breaks = seq(-6, 6, 2),
                     limits = c(-6, 6)) +
      scale_x_discrete(labels = c("MLS : 7th/8th", "Freq. of use", "Proficiency", "7th/8th grade", "MLS group", "(Intercept)")) +
  labs(title = "DOM Selection Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```


```{r, print-DOM-task-correlation-forest-plots}
(DOM_EPT_Plot) + (DOM_FCT_Plot)
```


---
# DOM: Individual Analyses

```{r, calculate-DOM-ind-diffs}
DOM_EPT_Diffs <- DOM_EPT %>%
  filter(!is.na(DOM_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(DOM_Use))

DOM_FCT_Diffs <- DOM_FCT %>%
  filter(!is.na(DOM_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(DOM_Use))

DOM_Diffs = left_join(DOM_EPT_Diffs, DOM_FCT_Diffs, by = "Part_ID", "Group") %>%
  rename(Group = Group.x,
         Production = Production.x,
         Preference = Production.y) %>% 
  mutate(Total = Preference + Production) %>% 
  filter(!Group == "SDBA")

DOM_Diffs$Group <- factor(DOM_Diffs$Group, levels = c("SDBA", "DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
```

```{r, generate-DOM-diffs-plot}
DOM_Diffs_Plot <- DOM_Diffs %>% 
  ggplot(aes(x = Production, y = Preference)) +
  geom_jitter(mapping = aes(color = Group), size = 3) +
  scale_x_continuous(breaks = seq (0, 10, 2),
                     limits = c(-1, 11)) +
  scale_y_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  labs(x = "Number of sentences with DOM produced", y = "Number of sentences with DOM selected", title = "Individual Rates of DOM Selection and Production", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, print-DOM-diffs-plot}
(DOM_Diffs_Plot)
```

---
# DOM: Summary


--
- Participants selected DOM more on the MST than they produced this structure on the SCT


--
- The older participants showed greater knowledge of DOM in the aggregate and the MST models (but not in the model evaluating production only)


--
- Participants with bilingual education produced more DOM than those in an English-only school


--
- Individual analyses provide additional insight into these findings:


--
  + 8 HSs produced DOM within the SDB range (16/18 or more)

--
  + None of the HS categorically omitted DOM (at least 2/18 instances, except for 1 participant)

--
  + Strong support for task asymmetries (selection > production)


---
# Subjunctive: Overall Correlation
  
```{r, generate-Subj-overall-correlation}
Subj_Overall <- glmer(Mood_Use ~ School + Age + Task + BESA_Total_Std + Use_Joint_Std + School:Age + Task:BESA_Total_Std + Task:Use_Joint_Std +
                       (1 | Part_ID) + (1 | Item),
                     data = Subj_Master_HS,
                     family = "binomial")
```

```{r, generate-Subj-overall-forest-plot}
Subj_Overall_Plot <- plot_model(Subj_Overall, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
            scale_y_continuous(breaks = seq(-3, 3, 1.5),
                     limits = c(-3, 3)) +
  scale_x_discrete(labels = c("MST : Use", "MST : Proficiency", "MLS : 7th/8th", "Frequency of use", "Proficiency", "MST", "7th/8th grade group", "MLS group", "(Intercept)")) +
  labs(title = "Subjunctive Overall Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-Subj-overall-forest-plot}
(Subj_Overall_Plot)
```


---
# Subjunctive: Task-Specific Variability

```{r, generate-subj-task-specific-correlations}
Subj_EPT_Correlation <- glmer(Mood_Use ~ School + Age + BESA_Other_Std + Use_Joint_Std + School:Age +
                               (1 | Part_ID) + (1 | Item),
                             data = Subj_EPT_HS,
                             family = "binomial")

Subj_FCT_Correlation <- glmer(Mood_Use ~ School + Age + BESA_Other_Std + Use_Joint_Std + School:Age +
                               (1 | Part_ID) + (1 | Item),
                             data = Subj_FCT_HS,
                             family = "binomial")
```

```{r, generate-subj-EPT-forest-plot}
Subj_EPT_Plot <- plot_model(Subj_EPT_Correlation, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
  scale_y_continuous(breaks = seq(-6, 6, 2),
                     limits = c(-6, 6)) +
  scale_x_discrete(labels = c("MLS : 7th/8th", "Freq. of use", "Proficiency", "7th/8th grade", "MLS group", "(Intercept)")) +
  labs(title = "Subjunctive Production Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, generate-subj-FCT-forest-plot}
Subj_FCT_Plot <- plot_model(Subj_FCT_Correlation, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
  scale_y_continuous(breaks = seq(-6, 6, 2),
                     limits = c(-6, 6)) +
  scale_x_discrete(labels = c("MLS : 7th/8th", "Freq. of use", "Proficiency", "7th/8th grade", "MLS group", "(Intercept)")) +
  labs(title = "Subjunctive Selection Model", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```


```{r, print-subj-task-correlation-forest-plots}
(Subj_EPT_Plot) + (Subj_FCT_Plot)
```

---
# Subjunctive: Individual Analyses

```{r, calculate-subj-ind-diffs}
Subj_EPT_Diffs <- Subj_EPT %>%
  filter(!is.na(Mood_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(Mood_Use))

Subj_FCT_Diffs <- Subj_FCT %>%
  filter(!is.na(Mood_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(Mood_Use))

Subj_Diffs = left_join(Subj_EPT_Diffs, Subj_FCT_Diffs, by = "Part_ID", "Group") %>%
  rename(Group = Group.x,
         Production = Production.x,
         Preference = Production.y) %>% 
  mutate(Total = Preference + Production) %>% 
  filter(!Group == "SDBA")

Subj_Diffs$Group <- factor(Subj_Diffs$Group, levels = c("SDBA", "DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
```

```{r, generate-subj-diffs-plot}
Subj_Diffs_Plot <- Subj_Diffs %>% 
  ggplot(aes(x = Production, y = Preference)) +
  geom_jitter(mapping = aes(color = Group), size = 3) +
  scale_x_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  scale_y_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  labs(x = "Number of sentences with subjunctive produced", y = "Number of sentences with subjunctive selected", title = "Individual Rates of Subjunctive Selection and Production", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, print-subj-diffs-plot}
(Subj_Diffs_Plot)
```


---
# Subjunctive: Summary


--
- Participants selected the subjunctive more on the MST than they produced it on the SCT


--
- Proficiency and frequency of use affected overall knowledge


--
  + Frequency of use emerged as significant in data from individual tasks

--
  + Proficiency was significant in the aggregate model only
  

--
- Older children showed greater use of the subjunctive, particularly on the MST


--
- Individual analyses support this claim


---
# Research Question 1


**Do the following experiential factors affect individual HSs' knowledge of volitional subjunctive and DOM: Frequency of use, morphosyntactic proficiency, DLI education?**


--
- Frequency of use affected HSs in overall subjunctive knowledge, and on the MST


--
- Proficiency affected overall rates of DOM and subjunctive use


--
- The effect for bilingual education was only evident for DOM in production


--
- Hypothesis #1 partially supported:

--
  + Each variable regressed onto certain areas (e.g., the production of DOM or receptive knowledge of subjunctive), but was not predictive of all outcomes

--
  + The predicted "strong" role for education in HL acquisition is not observed


---
# Research Question 2


**Do school-aged Spanish HSs show increasing knowledge of volitional subjunctive and DOM with age?**

--
- 7th/8th grade children produced/selected more DOM and subjunctive than 5th grade children


--
- The effect is particularly true for receptive knowledge of both structures


--
- Hypothesis #2 confirmed: cumulative exposure matters in the acquisition of these structures


---
# Research Question 3


**Do school-aged Spanish HSs show asymmetrical productive and receptive knowledge of volitional subjunctive and DOM?**


--
- HSs selected both structures more frequently on the MST than they produced them on the SCT


--
  + Drammatic assymetries are found for fifth graders (see descriptive statistics)
  
--
  + Individual analyses corrorborate this claim
  

--
- Hypothesis #3 confirmed: by this age, children seem adult-like with regards to Putnam and Sánchez's (2013) predictions of asymmetrical knowledge between these domains


---
# Research Question 4

**Does feature interpretability or interface variability have a greater effect on school-aged Spanish HSs knowledge of the structures tested?**


--
- Interface variability appears to matter more than interpretability


--
  + Subjunctive (uninterpretable feature of core syntax) > DOM (interpretable feature at the syn-sem interface)
  
--
  + Uninterpretable > interpretable (contra the Interpretability Hypothesis; Tsimpli & Dimitrakopoulou, 2007)

--
  + Core syntax > syntax-semantics interface (aligned with Interface Hypothesis; Tsimpli & Sorace, 2006)


--
- As stated previously, this question is an initial exploration into the role of different features in HL acquisition


--
  + A comparison of more than two features are needed

--
  + Other features such as gender could provide interesting insight into this question


---
# Theoretical Implications


--
- There is no evidence of altogether incomplete acquisition at the group level


--
  + HSs do show differential rates of DOM/volitional subjunctive production and selection
  
--
  + ALL HSs produced/selected DOM and the volitional subjunctive in at least two contexts

--
  + Growth continues with age: HSs protracted development may still be in progress (therefore, it can't be incomplete)


--
- No evidence whatsoever of attrition (longitudinal data would be helpful)
  

--
- Individual variability captures differences within HSs, which moves away from the need for group-level analyses or "labels" of acquisition altogether


---
# Theoretical Implications (continued)


- Mixed evidence for Putnam and Sánchez's (2013) framework


--
- Support for the feature-oriented approach

--
  + Language experience (proficiency, frequency of use, bilingual education) influences individual performance

--
  + Task asymmetries/differential access
  

--
- Counterevidence in the form of the directionality of age effects:

--
  + HSs' knowledge is not "reassembled" over time (if it were, we would see a decrease in the use of these structures)

--
  + Therefore, claims of reassembly in adults may not be due to changes during childhood


--
- Could the stages of Putnam and Sánchez's (2013) hypothesis run in reverse?


--
  + Evidence in the acquisition of DOM by heritage adults (Thane, Accepted)

--
  + Evidence in the acquisition of preterit aspect by L2 learners (Thane, 2023)


---
# The Paradigm Shift


- This project moves away from traditional group-level comparisons involving HSs (along with Giancaspro et al., 2022)


--
  + Shows that some HSs converge on the levels of SDBA, but does not require positing incomplete acquisition to show this

--
  + Development is modulated by factors at different levels of variability (within-group/between-speaker, within-speakers)


--
- Within-group/between-speaker variability


--
  + Education influences the development of Spanish as a heritage language, in some instances

--
  + Proficiency and frequency of use affect differences between speakers (current exposure)

--
  + Age also influences knowledge, particularly at the receptive level (cumulative exposure)


--
- Within-speaker variability


--
  + Receptive knowledge is stronger than production

--
  + Knowledge of certain structures is more robust than others (possibly due to interfaces)


---
# The Role of DLI Education in HL Development


- DLI does not *categorically* guarantee an advantage in the acquisition of Spanish as a heritage language


--
  + No influence on the development of the subjunctive

--
  + HSs in immersion programs continue to show (considerable) rates of optionality in their production of inflectional morphology


--
- Increases in quantity and quality of input *alone* do not guarantee an advantage in acquisition


--
- How can we support language development in immersion programs?


--
  + Promotion of linguistic equity within these programs (Babino & Stewart, 2017)

--
  + Strategic development of biliteracy (Urow & Beeman, 2013)

--
  + Implications for translanguaging and its implementation in bilingual schools

---
# References


Babino, A., & Stewart, M. A. (2017). “I like English better”: Latino dual language students’ investment in Spanish, English, and bilingualism. *Journal of Latinos and Education*, *16*(1), 18–29. https://doi.org/10.1080/15348431.2016.1179186

Barnett, W. S., Yarosz, D. J., Thomas, J., Jung, K., & Blanco, D. (2007). Two-way and monolingual English immersion in preschool education: An experimental comparison. *Early Childhood Research Quarterly*, *22*(3), 277–293. https://doi.org/10.1016/j.ecresq.2007.03.003

Bayram, F., Rothman, J., Iverson, M., Kupisch, T., Miller, D., Puig-Mayenco, E., & Westergaard, M. (2017). Differences in use without deficiencies in competence: Passives in the Turkish and German of Turkish heritage speakers in Germany. *International Journal of Bilingual Education and Bilingualism*, *22*(8), 919–939. https://doi.org/10.1080/13670050.2017.1324403

Castilla-Earls, A., Francis, D., Iglesias, A., & Davidson, K. (2019). The impact of the Spanish-to-English proficiency shift on the grammaticality of English learners. *Journal of Speech, Language, and Hearing Research*, *62*(6), 1739–1754. https://doi.org/10.1044/2018_JSLHR-L-18-0324

Corbet, J., & Domínguez, L. (2020). The comprehension of tense–aspect morphology by Spanish heritage speakers in the United Kingdom. *Languages*, *5*(46), 1–20. https://doi.org/10.3390/languages5040046

---
# References


Cuza, A., & Miller, L. (2015). The protracted acquisition of past tense aspectual values in child heritage Spanish. In R. Klassen, J. M. Liceras, & E. Valenzuela (Eds.), *Issues in Hispanic and Lusophone linguistics* (Vol. 4, pp. 211–230). John Benjamins. https://doi.org/10.1075/ihll.4.11cuz

Cuza, A., Miller, L., Pérez-Tattam, R., & Ortiz Vergara, M. (2019). Structure complexity effects in child heritage Spanish: The case of the Spanish personal *a*. *International Journal of Bilingualism*, *23*(6), 1333–1357. https://doi.org/10.1177/1367006918786467

Dracos, M., & Requena, P. E. (2022). Child heritage speakers’ acquisition of the Spanish subjunctive in volitional and adverbial clauses. *Language Acquisition*, *30*(1), 1–28. https://doi.org/10.1080/10489223.2022.2071156

Dracos, M., Requena, P., & Miller, K. (2019). Acquisition of mood selection in Spanish-speaking children. Language Acquisition, *26*(1), 106–118. https://doi.org/10.1080/10489223.2018.1464006

Giancaspro, D., Perez‐Cortes, S., & Higdon, J. (2022). (Ir)regular mood swings: Lexical variability in heritage speakers’ oral production of subjunctive mood. *Language Learning*, *72*(2), 456–496. https://doi.org/10.1111/lang.12489

Giancaspro, D., & Sánchez, L. (2021). Me, mi, my: Innovation and variability in heritage speakers’ knowledge of inalienable possession. *Glossa: A Journal of General Linguistics*, *6*(1). https://doi.org/10.5334/gjgl.1240

---
# References


Goebel-Mahrle, T., & Shin, N. L. (2020). A corpus study of child heritage speakers’ Spanish gender agreement. *International Journal of Bilingualism*, *24*(5–6), 1088–1104. https://doi.org/10.1177/1367006920935510

Guijarro-Fuentes, P. (2012). The acquisition of interpretable features in L2 Spanish: Personal *a*. *Bilingualism: Language and Cognition*, *15*(4), 701–720. https://doi.org/10.1017/S1366728912000144

Guijarro-Fuentes, P., Pires, A., & Nediger, W. (2017). Delay in the acquisition of Differential Object Marking by Spanish monolingual and bilingual teenagers. *International Journal of Bilingualism*, *21*(2), 159–177. https://doi.org/10.1177/1367006915601249

Hicks, G., & Domínguez, L. (2020). A model for L1 grammatical attrition. *Second Language Research*, *36*(2), 143–165. https://doi.org/10.1177/0267658319862011

Kupisch, T., & Rothman, J. (2018). Terminology matters! Why difference is not incompleteness and how early child bilinguals are heritage speakers. *International Journal of Bilingualism*, *22*(5), 564–582. https://doi.org/10.1177/1367006916654355

Lardiere, D. (2009). Some thoughts on the contrastive analysis of features in second language acquisition. *Second Language Research*, *25*(2), 173–227. https://doi.org/10.1177/0267658308100283

---
# References


López Otero, J. C. (2022). Lexical frequency effects on the acquisition of syntactic properties in heritage Spanish: A study on unaccusative and unergative predicates. *Heritage Language Journal*, *19*(1), 1–37. https://doi.org/10.1163/15507076-bja10011

Martinez-Nieto, L., & Restrepo, M. A. (2022). Production and comprehension of grammatical gender by Spanish heritage speakers: Evidence from accusative clitic pronouns. *International Journal of Bilingualism*, 136700692110573. https://doi.org/10.1177/13670069211057318

Merino, B. J. (1983). Language loss in bilingual Chicano children. *Journal of Applied Developmental Psychology*, *4*(3), 277–294. https://doi.org/10.1016/0193-3973(83)90023-0

Montrul, S. (2008). *Incomplete acquisition in bilingualism: Re-examining the age factor* (Vol. 39). John Benjamins Publishing Company. https://doi.org/10.1075/sibil.39

Montrul, S. (2009). Knowledge of tense-aspect and mood in Spanish heritage speakers. *International Journal of Bilingualism*, *13*(2), 239–269. https://doi.org/10.1177/1367006909339816

Montrul, S. (2013). Incomplete L1 acquisition. In J. Herschensohn & M. Young-Scholten (Eds.), *The Cambridge handbook of second language acquisition* (pp. 353–371). Cambridge University Press. https://doi.org/10.1017/CBO9781139051729.022

---
# References


Montrul, S. (2018). Heritage language development: Connecting the dots. *International Journal of Bilingualism*, *22*(5), 530–546. https://doi.org/10.1177/1367006916654368

Montrul, S., & Potowski, K. (2007). Command of gender agreement in school-age Spanish-English bilingual children. *International Journal of Bilingualism*, *11*(3), 301–328. https://doi.org/10.1177/13670069070110030301

Montrul, S., & Sánchez-Walker, N. (2013). Differential object marking in child and adult Spanish heritage speakers. *Language Acquisition*, *20*(2), 109–132. https://doi.org/10.1080/10489223.2013.766741

Peña, E. D., Gutiérrez-Clellen, V. F., Iglesias, A., Goldstein, B. A., & Bedore, L. M. (2014). *Bilingual English-Spanish Assessment (BESA)*. BROOKES Publishing CO. https://brookespublishing.com/wp-content/uploads/2021/01/BESA-excerpt.pdf

Perez-Cortes, S. (2016). *Acquiring obligatory and variable mood selection: Spanish heritage speakers’ and L2 learners’ performance in desideratives and reported speech contexts.* [Doctoral dissertation]. Rutgers University – New Brunswick.

Putnam, M. T., & Sánchez, L. (2013). What’s so incomplete about incomplete acquisition?: A prolegomenon to modeling heritage language grammars. *Linguistic Approaches to Bilingualism*, *3*(4), 478–508. https://doi.org/10.1075/lab.3.4.04put

---
# References

Rodríguez-Mondoñedo, M. (2008). The acquisition of Differential Object Marking in Spanish. *Probus*, *20*(1). https://doi.org/10.1515/PROBUS.2008.004

Thane, P. D. (2023). Frequency matters in second language acquisition too! Frequency effects in the production of preterit morphology with state verbs in Spanish second language learners and heritage speakers. *Lingua*, *295*, 1–20. https://doi.org/10.1016/j.lingua.2023.103611

Thane, P. D. (Accepted). Age effects in the heritage language acquisition of differential object marking in Spanish: Late childhood to adulthood. *International Journal of Bilingualism*.

Tsimpli, I. M., & Dimitrakopoulou, M. (2007). The Interpretability Hypothesis: Evidence from wh-interrogatives in second language acquisition. *Second Language Research*, *23*(2), 215–242. https://doi.org/10.1177/0267658307076546

Tsimpli, I., & Sorace, A. (2006). Differentiating interfaces-L2 performance in syntax-semantics and syntax-discourse phenomena. In D. Bamman, T. Magnitskaia, & C. Zaller (Eds.), *Proceedings of the 30th Annual Boston University Conference on Language Development* (pp. 653–664). Cascadilla Proceedings Project.

Urow, C., & Beeman, K. (2013). *Teaching for biliteracy: Strengthening bridges between languages*. Caslon Publishing.


---
# Appendix A: Sample Items


**Sample sentence from SCT**

A veces Juanito se pone triste si sus hermanas dicen que no quieren hablar con él. ¿Qué quiere la mamá? Quiere que las hermanas _________ (LLAMAR) Juanito cada noche.



**Sample sentences from MST**

La mamá sabe que Juanito tiene miedo de alturas. No quiere subir al tobogán en el parque de juegos solito. ¿Qué quiere la mamá?

  + Quiere que las hermanas lo lleven para subir al tobogán. 

  + *Quiere que las hermanas lo llevan para subir al tobogán. 


¿Qué tienen que hacer las hermanas? 
  + Tienen que llevar a Juanito para subir al tobogán. 

  + *Tienen que llevar Juanito para subir al tobogán.


---
# Appendix B: Proficiency Effects

```{r, select-individual-items for analysis}
# DOM
DOM_EPT_Single <- DOM_Master_HS %>% 
  filter(Item == "EPT-02") %>% 
  filter(!Group == "SDB")

DOM_FCT_Single <- DOM_Master_HS %>% 
  filter(Item == "FCT-03") %>% 
  filter(!Group == "SDB")

DOM_Single <- rbind(DOM_EPT_Single, DOM_FCT_Single)
DOM_Single$Group <- factor(DOM_Single$Group, levels = c("SDBA", "DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))


# Subjunctive
Subj_EPT_Single <- Subj_Master %>% 
  filter(Item == "EPT-02") %>% 
  filter(!Group == "SDB")

Subj_FCT_Single <- Subj_Master %>% 
  filter(Item == "FCT-02") %>% 
  filter(!Group == "SDB")

Subj_Single <- rbind(Subj_EPT_Single, Subj_FCT_Single)
Subj_Single$Group <- factor(Subj_Single$Group, levels = c("SDBA", "DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
```

```{r, generate-DOM-proficiency-plot}
DOM_Proficiency_Plot <- DOM_Single %>% 
  ggplot(aes(x = BESA_Total, y = Part_Avg)) + 
  geom_jitter(mapping = aes(color = Group), size = 3) +
  geom_smooth(method = glm) +
  facet_grid(rows = vars(Task)) +
  scale_y_continuous(breaks = seq (0, 10, 2),
                     limits = c(-1, 11)) +
  scale_x_continuous(breaks = seq (0, 18, 3),
                     limits = c(-1, 19)) +
  labs(x = "BESA score (with subjunctive items)", y = "Number of responses with DOM", title = "DOM", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, generate-subj-proficiency-plot}
Subj_Proficiency_Plot <- Subj_Single %>% 
  ggplot(aes(x = BESA_Other, y = Part_Avg)) + 
  geom_jitter(mapping = aes(color = Group), size = 3) +
  geom_smooth(method = glm) +
  facet_grid(rows = vars(Task)) +
  scale_y_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  scale_x_continuous(breaks = seq (0, 14, 2),
                     limits = c(-1, 15)) +
  labs(x = "BESA score (without subjunctive items)", y = "Number of responses with subjunctive",
       title = "Subjunctive", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, print-proficiency-plots, fig.height = 7, fig.width = 16, fig.align = 'center', fig.retina = 3}
(DOM_Proficiency_Plot)+(Subj_Proficiency_Plot)
```


---
# Appendix C: Frequency of Use Effects

```{r, generate-DOM-use-plot}
DOM_Use_Plot <- DOM_Single %>% 
  ggplot(aes(x = Use_Joint, y = Part_Avg)) + 
  geom_jitter(mapping = aes(color = Group), size = 3) +
  geom_smooth(method = glm) +
  facet_grid(rows = vars(Task)) +
  scale_y_continuous(breaks = seq (0, 10, 2),
                     limits = c(-1, 11)) +
  scale_x_continuous(breaks = seq (0, 30, 5),
                     limits = c(-1, 31)) +
  labs(x = "Self-reported frequency of use", y = "Number of responses with DOM", title = "DOM", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, generate-subj-use-plot}
Subj_Use_Plot <- Subj_Single %>% 
  ggplot(aes(x = Use_Joint, y = Part_Avg)) + 
  geom_jitter(mapping = aes(color = Group), size = 3) +
  geom_smooth(method = glm) +
  facet_grid(rows = vars(Task)) +
  scale_y_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  scale_x_continuous(breaks = seq (0, 14, 2),
                     limits = c(-1, 15)) +
  labs(x = "Self-reported frequency of use", y = "Number of responses with subjunctive", title = "Subjunctive", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, print-use-plots, fig.height = 7, fig.width = 16, fig.align = 'center', fig.retina = 3}
(DOM_Use_Plot)+(Subj_Use_Plot)
```
