---
title: "Towards a Unified Theory of Heritage Language Acquisition: Evidence from Spanish Differential Object Marking Across Childhood"
author: "Patrick D. Thane<br/>University of Massachusetts - Amherst"
institute: "17th Generative Approaches to Second Language Acquisition Conference"
date: "Friday, May 3, 2024"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r, setup, include = FALSE}
library(tidyverse)
library(lme4)
library(xaringan)
library(xaringanthemer)
library(xaringanExtra)
library(lmerTest)
library(kableExtra)
library(patchwork)
library(jtools)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(emmeans)

options(htmltools.dir.version = FALSE)
options(scipen = 999)


knitr::opts_chunk$set(fig.width = 16, fig.height = 7, fig.retina = 3, fig.align = 'center', cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, hiline = TRUE)
```

```{r xaringan, include = FALSE, warning = FALSE}
style_duo_accent(
  primary_color = "#881C1C",
  secondary_color = "#212721",
  inverse_header_color = "#FFFDD0")
```

```{r xaringan-logo, echo = FALSE}
use_logo(
  image_url = "https://www.umass.edu/university-relations/sites/default/files/styles/tile_max_250px/public/2023-12/seal_workmark_source-lrg.png?itok=DOrf20Xh"
)
```

```{r, load-DOM-data}
# EPT
SDB_EPT <- read.csv("../CSV Files/SDB/SDB DOM EPT.csv") %>% 
  mutate(Group = "SDBA",
         GradeGroup = "SDBA",
         ChildGroup = "N/A",
         Task = "EPT",
         School = "Adult")

HSA_EPT <- read.csv("../CSV Files/Adult HS/Adult HS DOM EPT.csv") %>% 
  mutate(Group = "HSA",
         GradeGroup = "N/A",
         ChildGroup = "N/A",
         Task = "EPT",
         School = "Adult")

DLI78_EPT <- read.csv("../CSV Files/DLI-78/DLI-78 DOM EPT.csv") %>% 
  mutate(Group = "HS7/8",
         GradeGroup = "7th/8th grade",
         ChildGroup = "DLI-7/8",
         Task = "EPT",
         School = "Dual language")

MLS78_EPT <- read.csv("../CSV Files/MLS-78/MLS-78 DOM EPT.csv") %>% 
  filter(!School == "GBCS") %>%
  mutate(Group = "HS7/8",
         GradeGroup = "7th/8th grade",
         ChildGroup = "MLS-7/8",
         Task = "EPT",
         School = "English-only")

DLI5_EPT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM EPT.csv") %>% 
  mutate(Group = "HS5",
         GradeGroup = "5th grade",
         ChildGroup = "DLI-5",
         Task = "EPT",
         School = "Dual language")

MLS5_EPT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM EPT.csv") %>% 
  filter(!School == "GBCS") %>% 
  mutate(Group = "HS5",
         GradeGroup = "5th grade",
         ChildGroup = "MLS-5",
         Task = "EPT",
         School = "English-only")


# FCT
SDB_FCT <- read.csv("../CSV Files/SDB/SDB DOM FCT.csv") %>% 
  mutate(Group = "SDBA",
         GradeGroup = "SDBA",
         ChildGroup = "N/A",
         Task = "MST",
         School = "Adult")

HSA_FCT <- read.csv("../CSV Files/Adult HS/Adult HS DOM FCT.csv") %>% 
  mutate(Group = "HSA",
         GradeGroup = "N/A",
         ChildGroup = "N/A",
         Task = "MST",
         School = "Adult")

DLI78_FCT <- read.csv("../CSV Files/DLI-78/DLI-78 DOM FCT.csv") %>% 
  mutate(Group = "HS7/8",
         GradeGroup = "7th/8th grade",
         ChildGroup = "DLI-7/8",
         Task = "MST",
         School = "Dual language")

MLS78_FCT <- read.csv("../CSV Files/MLS-78/MLS-78 DOM FCT.csv") %>% 
  filter(!School == "GBCS") %>%
  mutate(Group = "HS7/8",
         GradeGroup = "7th/8th grade",
         ChildGroup = "MLS-7/8",
         Task = "MST",
         School = "English-only")

DLI5_FCT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM FCT.csv") %>% 
  mutate(Group = "HS5",
         GradeGroup = "5th grade",
         ChildGroup = "DLI-5",
         Task = "MST",
         School = "Dual language")

MLS5_FCT <- read.csv("../CSV Files/MLS-5/MLS-5 DOM FCT.csv") %>% 
  filter(!School == "GBCS") %>% 
  mutate(Group = "HS5",
         GradeGroup = "5th grade",
         ChildGroup = "MLS-5",
         Task = "MST",
         School = "English-only")

## Join datasets
HS8_EPT <- rbind(DLI78_EPT, MLS78_EPT)
HS8_FCT <- rbind(DLI78_FCT, MLS78_FCT)

HS5_EPT <- rbind(DLI5_EPT, MLS5_EPT)
HS5_FCT <- rbind(DLI5_FCT, MLS5_FCT)
```

```{r, generate-averages}
# Adult data
## SDB
SDB_EPT_Structure <- aggregate(SDB_EPT$DOM_Use, list(SDB_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_EPT_Structure <- SDB_EPT_Structure %>% rename(Structure_Avg = x)
SDB_EPT_Structure <- left_join(SDB_EPT, SDB_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

SDB_FCT_Structure <- aggregate(SDB_FCT$DOM_Use, list(SDB_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_FCT_Structure <- SDB_FCT_Structure %>% rename(Structure_Avg = x)
SDB_FCT_Structure <- left_join(SDB_FCT, SDB_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## HSA
HSA_EPT_Structure <- aggregate(HSA_EPT$DOM_Use, list(HSA_EPT$Structure), FUN = mean, na.rm = TRUE)
HSA_EPT_Structure <- HSA_EPT_Structure %>% rename(Structure_Avg = x)
HSA_EPT_Structure <- left_join(HSA_EPT, HSA_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HSA_FCT_Structure <- aggregate(HSA_FCT$DOM_Use, list(HSA_FCT$Structure), FUN = mean, na.rm = TRUE)
HSA_FCT_Structure <- HSA_FCT_Structure %>% rename(Structure_Avg = x)
HSA_FCT_Structure <- left_join(HSA_FCT, HSA_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


# Children by age group
## HS7/8
HS8_EPT_Structure <- aggregate(HS8_EPT$DOM_Use, list(HS8_EPT$Structure), FUN = mean, na.rm = TRUE)
HS8_EPT_Structure <- HS8_EPT_Structure %>% rename(Structure_Avg = x)
HS8_EPT_Structure <- left_join(HS8_EPT, HS8_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HS8_FCT_Structure <- aggregate(HS8_FCT$DOM_Use, list(HS8_FCT$Structure), FUN = mean, na.rm = TRUE)
HS8_FCT_Structure <- HS8_FCT_Structure %>% rename(Structure_Avg = x)
HS8_FCT_Structure <- left_join(HS8_FCT, HS8_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## HS5
HS5_EPT_Structure <- aggregate(HS5_EPT$DOM_Use, list(HS5_EPT$Structure), FUN = mean, na.rm = TRUE)
HS5_EPT_Structure <- HS5_EPT_Structure %>% rename(Structure_Avg = x)
HS5_EPT_Structure <- left_join(HS5_EPT, HS5_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HS5_FCT_Structure <- aggregate(HS5_FCT$DOM_Use, list(HS5_FCT$Structure), FUN = mean, na.rm = TRUE)
HS5_FCT_Structure <- HS5_FCT_Structure %>% rename(Structure_Avg = x)
HS5_FCT_Structure <- left_join(HS5_FCT, HS5_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


# Children by school group
## DLI7/8
DLI78_EPT_Structure <- aggregate(DLI78_EPT$DOM_Use, list(DLI78_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI78_EPT_Structure <- DLI78_EPT_Structure %>% rename(Structure_Avg = x)
DLI78_EPT_Structure <- left_join(DLI78_EPT, DLI78_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI78_FCT_Structure <- aggregate(DLI78_FCT$DOM_Use, list(DLI78_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI78_FCT_Structure <- DLI78_FCT_Structure %>% rename(Structure_Avg = x)
DLI78_FCT_Structure <- left_join(DLI78_FCT, DLI78_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS7/8
MLS78_EPT_Structure <- aggregate(MLS78_EPT$DOM_Use, list(MLS78_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS78_EPT_Structure <- MLS78_EPT_Structure %>% rename(Structure_Avg = x)
MLS78_EPT_Structure <- left_join(MLS78_EPT, MLS78_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS78_FCT_Structure <- aggregate(MLS78_FCT$DOM_Use, list(MLS78_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS78_FCT_Structure <- MLS78_FCT_Structure %>% rename(Structure_Avg = x)
MLS78_FCT_Structure <- left_join(MLS78_FCT, MLS78_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## DLI5
DLI5_EPT_Structure <- aggregate(DLI5_EPT$DOM_Use, list(DLI5_EPT$Structure), FUN = mean, na.rm = TRUE)
DLI5_EPT_Structure <- DLI5_EPT_Structure %>% rename(Structure_Avg = x)
DLI5_EPT_Structure <- left_join(DLI5_EPT, DLI5_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

DLI5_FCT_Structure <- aggregate(DLI5_FCT$DOM_Use, list(DLI5_FCT$Structure), FUN = mean, na.rm = TRUE)
DLI5_FCT_Structure <- DLI5_FCT_Structure %>% rename(Structure_Avg = x)
DLI5_FCT_Structure <- left_join(DLI5_FCT, DLI5_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## MLS5
MLS5_EPT_Structure <- aggregate(MLS5_EPT$DOM_Use, list(MLS5_EPT$Structure), FUN = mean, na.rm = TRUE)
MLS5_EPT_Structure <- MLS5_EPT_Structure %>% rename(Structure_Avg = x)
MLS5_EPT_Structure <- left_join(MLS5_EPT, MLS5_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

MLS5_FCT_Structure <- aggregate(MLS5_FCT$DOM_Use, list(MLS5_FCT$Structure), FUN = mean, na.rm = TRUE)
MLS5_FCT_Structure <- MLS5_FCT_Structure %>% rename(Structure_Avg = x)
MLS5_FCT_Structure <- left_join(MLS5_FCT, MLS5_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)
```

```{r, rejoin-datasets}
# By grade
EPT_Age <- rbind(SDB_EPT_Structure, HSA_EPT_Structure, HS8_EPT_Structure, HS5_EPT_Structure) %>% 
  mutate(Structure = "DOM")

FCT_Age <- rbind(SDB_FCT_Structure, HSA_FCT_Structure, HS8_FCT_Structure, HS5_FCT_Structure) %>% 
  mutate(Structure = "DOM")

Master_Age <- rbind(EPT_Age, FCT_Age)
Master_Age$Group <- factor(Master_Age$Group, levels = c("SDBA", "HSA", "HS7/8", "HS5"))
Master_Age$Task <- factor(Master_Age$Task, levels = c("EPT", "MST"))
```

```{r, generate-school-group}
EPT_School <- rbind(DLI78_EPT_Structure, MLS78_EPT_Structure, DLI5_EPT_Structure, MLS5_EPT_Structure)
FCT_School <- rbind(DLI78_FCT_Structure, MLS78_FCT_Structure, DLI5_FCT_Structure, MLS5_FCT_Structure)

Master_School <- rbind(EPT_School, FCT_School)

Master_School$GradeGroup <- factor(Master_School$GradeGroup, levels = c("7th/8th grade", "5th grade"))
Master_School$ChildGroup <- factor(Master_School$ChildGroup, levels = c("DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
Master_School$School <- factor(Master_School$School, levels = c("Dual language", "English-only"))
Master_School$Task <- factor(Master_School$Task, levels = c("EPT", "MST"))
```

```{r, generate-participant-part-avg}
EPT_Part_Avg <- aggregate(EPT_Age$DOM_Use, list(EPT_Age$Part_ID), FUN = sum, na.rm = TRUE)
EPT_Part_Avg <- EPT_Part_Avg %>% rename(Part_Avg = x)
EPT_Part_Avg <- left_join(EPT_Age, EPT_Part_Avg, by = c("Part_ID" = "Group.1")) %>% 
  filter(Item == "EPT-02") %>%
  filter(!Group == "SDBA")

FCT_Part_Avg <- aggregate(FCT_Age$DOM_Use, list(FCT_Age$Part_ID), FUN = sum, na.rm = TRUE)
FCT_Part_Avg <- FCT_Part_Avg %>% rename(Part_Avg = x)
FCT_Part_Avg <- left_join(FCT_Age, FCT_Part_Avg, by = c("Part_ID" = "Group.1")) %>% 
  filter(Item == "FCT-03") %>%
  filter(!Group == "SDBA")

Master_Part_Avg <- rbind(EPT_Part_Avg, FCT_Part_Avg)
Master_Part_Avg$Group <- factor(Master_Part_Avg$Group, levels = c("HSA", "HS7/8", "HS5"))
```

```{r, generate-boxplot-averages-by-age}
EPT_Age_Averages <- rbind(SDB_EPT, HSA_EPT, HS8_EPT, HS5_EPT) %>% 
  group_by(Part_ID, Group) %>%
  summarize(Total_DOM = sum(DOM_Use, na.rm = TRUE),
            Total_Responses = sum(!is.na(DOM_Use)),
            Ratio = Total_DOM/Total_Responses) %>% 
  mutate(Task = "EPT",
         Ratio = (Ratio*100))

FCT_Age_Averages <- rbind(SDB_FCT, HSA_FCT, HS8_FCT, HS5_FCT) %>% 
  group_by(Part_ID, Group) %>%
  summarize(Total_DOM = sum(DOM_Use, na.rm = TRUE),
            Total_Responses = sum(!is.na(DOM_Use)),
            Ratio = Total_DOM/Total_Responses) %>% 
  mutate(Task = "MST",
         Ratio = (Ratio*100))

Master_Age_Averages <- rbind(EPT_Age_Averages, FCT_Age_Averages)
```

```{r, generate-boxplot-averages-by-education}
EPT_School_Averages <- rbind(DLI78_EPT, MLS78_EPT, DLI5_EPT, MLS5_EPT) %>% 
  group_by(Part_ID, GradeGroup, School, ChildGroup) %>%
  summarize(Total_DOM = sum(DOM_Use, na.rm = TRUE),
            Total_Responses = sum(!is.na(DOM_Use)),
            Ratio = Total_DOM/Total_Responses) %>% 
  mutate(Task = "EPT",
         Ratio = (Ratio*100))

FCT_School_Averages <- rbind(DLI78_FCT, MLS78_FCT, DLI5_FCT, MLS5_FCT) %>% 
  group_by(Part_ID, GradeGroup, School, ChildGroup) %>%
  summarize(Total_DOM = sum(DOM_Use, na.rm = TRUE),
            Total_Responses = sum(!is.na(DOM_Use)),
            Ratio = Total_DOM/Total_Responses) %>% 
  mutate(Task = "MST",
         Ratio = (Ratio*100))

Master_School_Averages <- rbind(EPT_School_Averages, FCT_School_Averages)
Master_School_Averages$ChildGroup <- factor(Master_School_Averages$ChildGroup, levels = c("DLI-7/8", "MLS-7/8", "DLI-5", "MLS-5"))
```

# Introduction


--
- There has been extensive theorizing concerning HL acquisition and maintenance


--
  + Incomplete acquisition/divergent development (e.g., Montrul, 2008, 2013)

--
  + Protracted development
  
--
  + Reassembly of features (Putnam and Sánchez, 2013)


--
- School-aged heritage speakers are the "missing link" in heritage language acquisition (Montrul, 2018, p. 534)

--
  + Longitudinal data are necessary to definitively distinguish between end-state differences in development and language attrition

--
  + In the absence of such data, comparing age groups of children and adult HSs allows for a "quasi-longitudinal" approach to distinguish between theories
  

--
- López-Beltrán Forcada (2021, pp. 4-5): "achieving a unified theory of HL acquisition and processing requires a certain degree of common ground between all of these approaches."


---
# Identifying Layers of Variability


--
- Incomplete acquisition, attrition, protracted development: HSs diverge from other native speaker populations due to:

--
  + Insufficient input

--
  + Early onset of acquisition of an additional language

--
  + Changes in exposure leading to morphosyntactic optionality ("reactivation" of the inference engine, c.f. Hicks & Dominguez, 2020)


--
- Feature reassembly approach (Putnam & Sánchez): HSs differ from one another and at the within-speaker level

--
  + Patterns of activation (frequency of use and/or proficiency) affect the maintenance of HL features
  
--
  + Lower frequency of use ~ greater crosslinguistic influence and feature reassembly (between-speakers)

--
  + Asymmetrical performance between syntax (receptive tasks) vs. morphology (production tasks) due to mapping challenges and crosslinguistic influence (intraspeaker variability)

--
  + Reassembly begins with features bundled within less-frequent lexical items and in production before underlying knowledge (intraspeaker variability)


--
- Can we accomplish a theory that focuses on all of these levels (between-groups, between-speakers, within-speakers) and that can be bidirectional (e.g., acquisition and attrition) in child language acquisition?


---
# Education Effects in HL Acquisition


--
- Previous studies in European contexts have shown that literacy level and/or bilingual education affects acquisition and maintenance of HL competence (Bayram et al., 2017; Kupisch & Rothman, 2018; Torregrossa et al., 2023)


--
- Dual language immersion (DLI): opportunity in the U.S. for children to learn at least 50% of instruction in Spanish (or a different non-English language)


--
- Previous research on the acquisition of inflectional morphology on DLI has shown:

--
  + HSs show stronger competence than L2 learners (e.g., Montrul & Potowski, 2007; Potowski, 2007)

--
  + Spanish-dominant children show stronger knowledge of morphosyntax than HSs (e.g., Potowski, 2007)
  

--
- Does DLI confer an advantage relative to non-immersion environments in HL acquisition/maintenance?


---
# Differential Object Marking in Spanish


--
- Example of structure at the syntax-semantics interface in Spanish


--
- Spanish is one of 300+ languages to mark certain direct objects, particularly those that are [+animate] and [+specific] (Bossong, 1991)

--
  + Other entailments are involved with DOM, but only these two "core" considerations are discussed here

--
  + DOM-marked objects are preceded by the marker *a* (morphological realization of DOM)


--
- In Spanish, [+animate] and [+specific] direct objects raise from VP-internal to check a D-feature in _spec,vP_

--
  + Example of inherent case marking

--
  + The *a* marking is also present with all indirect objects to mark dative case regardless of semantics

---
# Why DOM?


--
- DOM is abundantly studied in HL acquisition, and appears to be prone to:

--
  + Patterns of HL exposure (Montrul & Sánchez-Walker, 2013)

--
  + Morphosyntactic proficiency (Arechabaleta-Regulez & Montrul, 2023; Guijarro-Fuentes et al., 2017; Montrul & Bowles, 2009)


--
- The role of age in DOM acquisition remains unclear:

--
  + Protracted development comparing children to adults (Montrul & Sánchez-Walker, 2013)

--
  + No age effects within children (Cuza et al., 2019) or adolescents (Guijarro-Fuentes et al., 2017)


--
- DOM attrition has even been attested in Spanish-dominant bilingual adults (Montrul & Sánchez-Walker, 2013)

--
- To date, there is not yet a single study on DOM that:

--
  + Has both productive and receptive tasks

--
  + Tests the role of proficiency and frequency of exposure

--
  + Traces development from early adolescence to adulthood


---
# Research Questions and Hypotheses


--
** 1. At the group and individual levels, do English-dominant HSs converge on the Spanish-dominant DOM system with animate and specific direct objects?**

--
  + Hypotheses: (A) Spanish-dominant bilingual adults will produce/select more DOM with animate and specific direct objects than HSs; (B) some HSs will converge on Spanish-dominant adults.


--
** 2. Do HSs develop stronger knowledge of DOM with animate and specific direct objects with age? **

--
  + Hypothesis: HSs will show protracted development of DOM from 5th grade through adulthood.


--
** 3. Do proficiency, frequency of HL exposure, and bilingual education influence HSs' command of DOM with animate and specific direct objects? **

--
  + Hypotheses: (A) HSs with greater proficiency and exposure will produce/select more DOM; (B) HSs who attend a dual-language immersion program will produce/select more DOM


--
** 4. Do HSs show stronger receptive knowledge of DOM than they produce this structure?**

--
  + Hypothesis: HSs will show stronger receptive knowledge than they exhibit in production.


---
# Participants


--
- 127 total participants

--
  + 41 5th grade Spanish HS, ages 10-11 (HS5)

--
  + 34 7th/8th grade Spanish HS, ages 12-14 (HS7/8)

--
  + 35 HS adults (HSA)

--
  + 18 Spanish-dominant bilingual adults (SDBA)


--
- HS5 and HS7/8 were further subdivided into different groups:

--
  + DLI students (received 50% of elementary education in Spanish + Spanish HL course in middle school)
  
--
  + Other students received a traditional monolingual English education

--
  + Both schools were selected because of their similar student bodies (socioeconomic status, home Spanish speakers, number of English language learners)


--
- Extensive diversity in Spanish spoken among bilingual adults:

--
  + HSAs spoke varieties from 13 total Spanish-speaking countries
  
--
  + SDBAs were raised in 9 Spanish-speaking countries and did not live in the USA until at least age 12


--
- Most children were of Mexican descent


---
# Background Tasks

--
- Children completed the Bilingual English-Spanish Assessment (BESA; Peña et al., 2018)

--
  + 18 questions evaluating multiple morphosyntactic structures

--
  + Number of expected answers used as proficiency score in statistical modeling


--
- Language questionnaire evaluating linguistic variables (children completed tasks per validity of child self-reports; Castilla-Earls et al., 2022):

--
  + Number of Spanish-speaking parents
  
--
  + Languages spoken at home

--
  + Frequency of use (1-5 Likert scale) across 6 contexts (continuous variable for statistics)


---
# Linguistic Tasks


--
- Production and selection tasks designed to tap [+animate] and [+specific] direct objects

--
  + Communicative context about a mother's wishes for her children at summer camp

--
  + Also used to test subjunctive mood morphology


--
- **Elicited production task (EPT)**: participants completed sentences targeting DOM orally

--
  + 10 sentences tapping knowledge of DOM with [+animate] and [+specific] direct objects

--
  + 3rd person plural verb (ending in -an) to increase salience of *a*

--
  + 6 distractors for children, 37 distractors for adults (following Montrul & Sánchez-Walker, 2013)


--
- **Morphology selection task (MST)**: written multiple choice between sentences with and without DOM

--
  + 8 sentences tapping DOM (one choice with *a* and one choice without any object marking)

--
  + DOM occurred only after 3rd person plural verbs (-an inflection) to increase salience of *a*

--
  + 14 distractors for children, 47 for adults


--
- DOM was only included preceding the proper noun _Juanito_ where no dialectal variability of DOM is attested


---
# DOM Production/Selection by Group

```{r, generate-age-boxplot}
Age_Boxplot <- Master_Age_Averages %>% 
  ggplot(aes(x = Group, y = Ratio, fill = Task)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq (0, 100, 20),
                     limits = c(0, 100)) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of responses with DOM", title = "DOM by Age Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, generate-school-boxplot}
School_Boxplot <- Master_School_Averages %>% 
  ggplot(aes(x = School, y = Ratio, fill = Task)) +
  geom_boxplot() +
  facet_grid(facets = vars(GradeGroup)) +
  scale_y_continuous(breaks = seq (0, 100, 20),
                     limits = c(0, 100)) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "School", y = "Percentage of responses with DOM", title = "DOM by School, Age, and Task") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(face = "bold", size = 18))
```

```{r, print-boxplots}
(Age_Boxplot)+(School_Boxplot)
```


---
# Individual Variability

```{r, generate-DOM-individual-values}
EPT_Diffs <- EPT_Age %>%
  filter(!is.na(DOM_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(DOM_Use))

FCT_Diffs <- FCT_Age %>%
  filter(!is.na(DOM_Use)) %>%
  group_by(Part_ID, Group) %>%
  summarize(Production = sum(DOM_Use))

Diffs = left_join(EPT_Diffs, FCT_Diffs, by = "Part_ID", "Group") %>%
  rename(Group = Group.x,
         Production = Production.x,
         Preference = Production.y) %>% 
  mutate(Total = Preference + Production) %>% 
  filter(!Group == "SDBA")

Diffs$Group <- factor(Diffs$Group, levels = c("HSA", "HS7/8", "HS5"))
```

```{r, generate-DOM-diffs-plot}
Diffs_Plot <- Diffs %>% 
  ggplot(aes(x = Production, y = Preference)) +
  geom_jitter(mapping = aes(color = Group), size = 3) +
  scale_x_continuous(breaks = seq (0, 10, 2),
                     limits = c(-1, 11)) +
  scale_y_continuous(breaks = seq (0, 8, 2),
                     limits = c(-1, 9)) +
  labs(x = "Number of sentences with DOM produced", y = "Number of sentences with DOM selected", title = "Individual Rates of DOM Selection and Production", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, print-DOM-diffs-plot}
(Diffs_Plot)
```

---
# Inferential Statistics (All Groups)

```{r, DOM-master-correlation}
DOM_Correlation <- glmer(DOM_Use ~ Group + BESA_Total_Std + Use_Joint_Std + Task + BESA_Total_Std:Task + Use_Joint_Std:Task +
                           (1 | Part_ID) + (1 | Item),
                         data = Master_Age,
                         family = "binomial")
```

```{r, generate-DOM-correlation-forest-plot}
DOM_Forest_Plot <- plot_model(DOM_Correlation, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
  labs(title = "GLMM Summary (All Participant Groups)", y = "Estimates") +
      scale_y_continuous(breaks = seq(-9, 9, 3),
                     limits = c(-9, 9)) +
    scale_x_discrete(labels = c("Use : MST", "BESA : MST", "MST", "Frequency of use", "BESA score", "Group (HS5)", "Group (HS7/8)", "Group (HSA)", "(Intercept)")) +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-DOM-correlation-forest-plot}
(DOM_Forest_Plot)
```


---
# Group Comparisons (All Groups)

```{r, generate-DOM-correlation-tukey}
DOM_Correlation_Pairwise <- emmeans(DOM_Correlation, spec = "Group")
DOM_Correlation_Tukey <- contrast(DOM_Correlation_Pairwise, method = "pairwise")
DOM_Correlation_Table <- summary(DOM_Correlation_Tukey)
```

```{r, show-DOM-correlation-Tukey-results}
kable(DOM_Correlation_Table[1:6], col.names = c("Contrast", "Estimate", "SE", "df", "t ratio", "p"), align = "lccccc")
```

--
* SDBA have higher production/selection rates of DOM than all groups of HSs (speaker group-level difference)


--
* HSA have significantly higher percentages of DOM production/selection than HS5 (protracted development over time)


--
* No differences emerge as significant when comparisons involve the "intermittent" HS7/8 group (development is slow and gradual)


---
# Effects of School and Age (HS Children)
```{r, DOM-school-correlation}
DOM_School <- glmer(DOM_Use ~ GradeGroup * School + 
                      (1 | Part_ID) + (1 | Item),
                    data = Master_School,
                    family = "binomial")
```

```{r, generate-DOM-child-forest-plot}
DOM_School_Plot <- plot_model(DOM_School, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
      scale_y_continuous(breaks = seq(-9, 9, 3),
                     limits = c(-9, 9)) +
    scale_x_discrete(labels = c("HS5 : English-only school", "English-only school", "HS5", "(Intercept)")) +
  labs(title = "GLMM Summary (Children)", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-DOM-child-forest-plot}
(DOM_School_Plot)
```


---
# Group Comparisons (HS Children)

```{r, generate-child-group-correlation}
DOM_Children <- glmer(DOM_Use ~ + ChildGroup +
                        (1 | Part_ID) + (1 | Item),
                      data = Master_School,
                      family = "binomial")
```

```{r, generate-DOM-child-tukey}
DOM_Children_Pairwise <- emmeans(DOM_Children, spec = "ChildGroup")
DOM_Children_Tukey <- contrast(DOM_Children_Pairwise, method = "pairwise")
DOM_Children_Table <- summary(DOM_Children_Tukey)
```

```{r, show-DOM-child-Tukey-results}
kable(DOM_Children_Table[1:6], col.names = c("Contrast", "Estimate", "SE", "df", "t ratio", "p"), align = "lccccc")
```


--
- Children enrolled in the dual language immersion school in 7th/8th grade used significantly more DOM than all other groups of children


--
- Suggests clear interaction between age and school (supported by statistics)


--
- Age and bilingual education work together


---
# Research Question #1


--
- **At the group and individual levels, do English-dominant HSs converge on the Spanish-dominant DOM system with animate and specific direct objects?**


--
- Hypotheses:

--
  + (A) Spanish-dominant bilingual adults will produce/select more DOM with animate and specific direct objects than HSs

--
  + (B) some HSs will converge on Spanish-dominant adults.


--
- SDBAs do not show variability in producing/selecting DOM with animate and specific direct objects


--
- HSs show more heterogeneity / variability in DOM production and selection than SDBAs


--
- HSs *can* perform within the range of SDBAs (19/109 HSs were within SDBA range)


--
- Hypotheses 1A and 1B upheld


---
# Research Question #2


--
- **Do HSs develop stronger knowledge of DOM with animate and specific direct objects with age? **


--
- Hypothesis: HSs will show protracted development of DOM from 5th grade through adulthood.


--
- HSA > HS5 in statistical modeling; gradual increases across age groups in descriptive statistics


--
- Hypothesis #2 upheld


--
- Joins a growing body of work showing continuation of HL acquisition late into childhood and even into adulthood in multiple languages:

--
  + Greek (Daskalaki et al., 2022)

--
  + Mandarin (Jia & Paradis, 2020)

--
  + Portuguese (Flores & Barbosa, 2014; Flores et al., 2017)
  
--
  + Spanish (e.g., Corbet & Domínguez, 2020; Cuza & Miller, 2015; Montrul & Sánchez-Walker, 2013)


---
# Research Question #3


--
- **Do proficiency, frequency of HL exposure, and bilingual education influence HSs' command of DOM with animate and specific direct objects?**


--
- Hypotheses:


--
  + (A) HSs with greater proficiency and exposure will produce/select more DOM

--
  + (B) HSs who attend a dual-language immersion program will produce/select more DOM


--
- Proficiency affected HSs' production/selection of DOM


---
# Research Question #4


--
- **Do HSs show stronger receptive knowledge of DOM than they produce this structure?**


--
- Hypothesis: HSs will show stronger receptive knowledge than they exhibit in production.


--
- HSs' performance on the MST was stronger than on the EPT; all HSs showed *some* use of DOM


--
- Hypothesis #4 upheld


--
- Supports research arguing for a dissociation between morphology and syntax in HL acquisition


---
# Theoretical Discussion


--
- HSs diverge as a group for SDBAs (supporting incomplete acquisition at the between-group level)


--
- Putnam and Sánchez (2013) discuss feature _re_ assembly, yet some of their predictions can be adapted to account for results at the within-speaker level

--
  + Proficiency and bilingual education affect the acquisition of DOM

--
  + Receptive knowledge stronger than production (syntactic representation intact; morphological realization more variable)

--
  + No evidence of _re_ assembly of features, but rather protracted development


--
- Is it possible that Putnam and Sánchez's (2013) approach runs in reverse in initial protracted HL acquisition

--
  + According to descriptive statistics, comprehension stabilizes before production

--
  + Exposure appears to modulate the rate and ultimate point of development


--
- Incomplete acquisition *and* Putnam and Sánchez's (2013) approach work together at distinct levels

--
  + Incomplete acquisition: differences between HSs and other groups of native speakers

--
  + Putnam & Sánchez's (2013) activation approach: proficiency modulates differences between and within individual HSs


---
# Limitations


--
- Only evaluated DOM in "core" contexts such as [+animate] and [+specific] direct objects


--
- Broader age range to make a clearer set of predictions regarding protracted development: how would younger bilinguals perform?


--
- Did not control for age of acquisition using statistical modeling (only requested number of bilingual parents)


---
# References


Arechabaleta Regulez, B., & Montrul, S. (2023). Production, acceptability, and online comprehension of Spanish differential object marking by heritage speakers and L2 learners. *Frontiers in Psychology*, *14*, 1106613. https://doi.org/10.3389/fpsyg.2023.1106613

Bayram, F., Rothman, J., Iverson, M., Kupisch, T., Miller, D., Puig-Mayenco, E., & Westergaard, M. (2017). Differences in use without deficiencies in competence: Passives in the Turkish and German of Turkish heritage speakers in Germany. *International Journal of Bilingual Education and Bilingualism*, *22*(8), 919–939. https://doi.org/10.1080/13670050.2017.1324403

Bossong, G. (1991). Differential object marking in Romance and beyond. In D. Wanner & D. A. Kibbee (Eds.), *Current issues in linguistic theory* (Vol. 69, p. 143). John Benjamins Publishing Company. https://doi.org/10.1075/cilt.69.14bos

Castilla-Earls, A., Ronderos, J., & Fitton, L. (2022). Can bilingual children self-report their bilingual experience and proficiency? The Houston Questionnaire. *Journal of Speech, Language, and Hearing Research*, *65*(10), 3835–3853. https://doi.org/10.1044/2022_JSLHR-21-00675

Corbet, J., & Domínguez, L. (2020). The comprehension of tense–aspect morphology by Spanish heritage speakers in the United Kingdom. *Languages*, *5*(46), 1–20. https://doi.org/10.3390/languages5040046


---
# References

Cuza, A., & Miller, L. (2015). The protracted acquisition of past tense aspectual values in child heritage Spanish. In R. Klassen, J. M. Liceras, & E. Valenzuela (Eds.), *Issues in Hispanic and Lusophone linguistics* (Vol. 4, pp. 211–230). John Benjamins. https://doi.org/10.1075/ihll.4.11cuz

Cuza, A., Miller, L., Pérez-Tattam, R., & Ortiz Vergara, M. (2019). Structure complexity effects in child heritage Spanish: The case of the Spanish personal a. *International Journal of Bilingualism*, *23*(6), 1333–1357. https://doi.org/10.1177/1367006918786467

Daskalaki, E., Chondrogianni, V., & Blom, E. (2023). Path and rate of development in child heritage speakers: Evidence from Greek subject/object form and placement. *International Journal of Bilingualism*, *27*(5), 634–662. https://doi.org/10.1177/13670069221111648

Flores, C., & Barbosa, P. (2014). When reduced input leads to delayed acquisition: A study on the acquisition of clitic placement by Portuguese heritage speakers. *International Journal of Bilingualism*, *18*(3), 304–325. https://doi.org/10.1177/1367006912448124

Flores, C., Santos, A. L., Jesus, A., & Marques, R. (2017). Age and input effects in the acquisition of mood in Heritage Portuguese. *Journal of Child Language*, *44*(4), 795–828. https://doi.org/10.1017/S0305000916000222


---
# References


Guijarro-Fuentes, P., Pires, A., & Nediger, W. (2017). Delay in the acquisition of Differential Object Marking by Spanish monolingual and bilingual teenagers. *International Journal of Bilingualism*, *21*(2), 159–177. https://doi.org/10.1177/1367006915601249

Hicks, G., & Domínguez, L. (2020). A model for L1 grammatical attrition. *Second Language Research*, *36*(2), 143–165. https://doi.org/10.1177/0267658319862011

Jia, R., & Paradis, J. (2020). The acquisition of relative clauses by Mandarin heritage language children. *Linguistic Approaches to Bilingualism*, *10*(2), 153–183. https://doi.org/10.1075/lab.16015.jia

Kupisch, T., & Rothman, J. (2018). Terminology matters! Why difference is not incompleteness and how early child bilinguals are heritage speakers. *International Journal of Bilingualism*, *22*(5), 564–582. https://doi.org/10.1177/1367006916654355

López-Beltrán Forcada, P. (2021). *Heritage speakers’ online processing of the Spanish subjunctive: A comprehensive usage-based study* [Doctoral dissertation]. The Pennsylvania State University.

Montrul, S. (2008). *Incomplete acquisition in bilingualism: Re-examining the age factor* (Vol. 39). John Benjamins Publishing Company. https://doi.org/10.1075/sibil.39


---
# References

Montrul, S. (2013). Incomplete L1 acquisition. In J. Herschensohn & M. Young-Scholten (Eds.), *The Cambridge handbook of second language acquisition* (pp. 353–371). Cambridge University Press. https://doi.org/10.1017/CBO9781139051729.022

Montrul, S. (2018). Heritage language development: Connecting the dots. *International Journal of Bilingualism*, *22*(5), 530–546. https://doi.org/10.1177/1367006916654368

Montrul, S., & Bowles, M. (2009). Back to basics: Incomplete knowledge of Differential Object Marking in Spanish heritage speakers. *Bilingualism: Language and Cognition*, *12*(3), 363–383. https://doi.org/10.1017/S1366728909990071

Montrul, S., & Potowski, K. (2007). Command of gender agreement in school-age Spanish-English bilingual children. *International Journal of Bilingualism*, *11*(3), 301–328. https://doi.org/10.1177/13670069070110030301

Montrul, S., & Sánchez-Walker, N. (2013). Differential object marking in child and adult Spanish heritage speakers. *Language Acquisition*, *20*(2), 109–132. https://doi.org/10.1080/10489223.2013.766741

Peña, E. D., Gutiérrez-Clellen, V. F., Iglesias, A., Goldstein, B. A., & Bedore, L. M. (2018). *Bilingual English Spanish Assessment (BESA)*. Brookes.


---
# References


Potowski, K. (2007). Characteristics of the Spanish grammar and sociolinguistic proficiency of dual immersion graduates. *Spanish in Context*, *4*(2), 187–216. https://doi.org/10.1075/sic.4.2.04pot

Putnam, M. T., & Sánchez, L. (2013). What’s so incomplete about incomplete acquisition?: A prolegomenon to modeling heritage language grammars. *Linguistic Approaches to Bilingualism*, *3*(4), 478–508. https://doi.org/10.1075/lab.3.4.04put

Torregrossa, J., Flores, C., & Rinke, E. (2023). What modulates the acquisition of difficult structures in a heritage language? A study on Portuguese in contact with French, German and Italian. *Bilingualism: Language and Cognition*, *26*(1), 179–192. https://doi.org/10.1017/S1366728922000438


---
# Appendix A: Percentages of DOM Used

```{r, bar-graph-by-age}
Master_Age_Plot <- Master_Age %>% 
  ggplot(aes(x = Group, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 105)) +
  geom_text(aes(label = round(Structure_Percentage)),
            position = position_dodge(width = .9),
            vjust = -0.5,
            size = 3.5) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of responses with DOM", fill = "Task", title = "DOM by Age Group and Task") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(face = "bold", size = 18))
```

```{r, bar-graph-by-school}
School_Age_Plot <- Master_School %>% 
  ggplot(aes(x = GradeGroup, y = Structure_Percentage, fill = Task)) +
  facet_grid(facets = vars(School)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 105)) +
  geom_text(aes(label = round(Structure_Percentage)),
            position = position_dodge(width = .9),
            vjust = -0.5,
            size = 3.5) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Child age group", y = "Percentage of responses with DOM", fill = "Task", title = "DOM by School Group and Task") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(face = "bold", size = 18))
```

```{r, print-bar-graphs}
(Master_Age_Plot)+(School_Age_Plot)
```


---
# Appendix B: DOM by Proficiency/Freq. of Use

```{r, generate-proficiency-plot}
Proficiency_Plot <- Master_Part_Avg %>% 
  ggplot(aes(x = BESA_Total, y = Part_Avg)) + 
  geom_jitter(mapping = aes(color = Group), size = 3) +
  geom_smooth(method = glm) +
  facet_grid(rows = vars(Task)) +
  scale_y_continuous(breaks = seq (0, 10, 2),
                     limits = c(-1, 11)) +
  scale_x_continuous(breaks = seq (0, 18, 3),
                     limits = c(3, 18)) +
  labs(x = "BESA proficiency score", y = "Number of DOM responses per participant", 
       title = "DOM by BESA Proficiency and Task") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(face = "bold", size = 18))
```

```{r, generate-use-plot}
Use_Plot <- Master_Part_Avg %>% 
  ggplot(aes(x = Use_Joint, y = Part_Avg)) + 
  geom_jitter(mapping = aes(color = Group), size = 3) +
  geom_smooth(method = glm) +
  facet_grid(rows = vars(Task)) +
  scale_y_continuous(breaks = seq (0, 10, 2),
                     limits = c(-1, 11)) +
  scale_x_continuous(breaks = seq (0, 30, 5),
                     limits = c(2.5, 31)) +
  labs(x = "Self-reported frequency of use of Spanish", y = "Number of DOM responses per participant", 
       title = "DOM by Frequency of Use and Task") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(face = "bold", size = 18))
```

```{r, print-use-and-proficiency-plots}
(Proficiency_Plot)+(Use_Plot)
```