---
title: "Behind the Scenes: Constructing Manuscripts to Build Heritage Language Theory"
author: "Patrick D. Thane (he/him/his)"
institute: "University of Houston Dual Lab"
date: "February 4, 2025"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r, setup, include = FALSE}
library(tidyverse)
library(lme4)
library(xaringan)
library(xaringanthemer)
library(xaringanExtra)
library(lmerTest)
library(kableExtra)
library(patchwork)
library(jtools)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(emmeans)
library(here)

options(htmltools.dir.version = FALSE)
options(scipen = 999)


knitr::opts_chunk$set(fig.width = 16, fig.height = 7, fig.retina = 3, fig.align = 'center', cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, hiline = TRUE)
```

```{r xaringan, include = FALSE, warning = FALSE}
style_duo_accent(
  primary_color = "#C8102E",
  secondary_color = "#888B8D",
  inverse_header_color = "#FFF9D9")
```

```{r xaringan-logo, echo = FALSE}
use_logo(
  image_url = "https://uh.edu/brand/_img/uh_red.png"
)
```

```{r, load-subj-data}
# EPT
SDB_EPT <- read_csv(here("./CSV Files/SDB/SDB Subjunctive EPT.csv")) %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "SDBA",
         Task = "EPT",
         Mood = "Subjunctive",
         School = "SDBA",
         Age = "SDBA")

HSA_EPT <- read_csv(here("./CSV Files/Adult HS/Adult HS Subjunctive EPT.csv")) %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "HSA",
         Task = "EPT",
         Mood = "Subjunctive",
         School = "HSA",
         Age = "HSA")

DLI78_EPT <- read_csv(here("./CSV Files/DLI-78/DLI-78 Subjunctive EPT.csv")) %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-7/8",
         Task = "EPT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "HS7/8")

MLS78_EPT <- read_csv(here("./CSV Files/MLS-78/MLS-78 Subjunctive EPT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-7/8",
         Task = "EPT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "HS7/8")

DLI5_EPT <- read_csv(here("./CSV Files/DLI-5/DLI-5 Subjunctive EPT.csv")) %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-5",
         Task = "EPT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "HS5")

MLS5_EPT <- read_csv(here("./CSV Files/MLS-5/MLS-5 Subjunctive EPT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-5",
         Task = "EPT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "HS5")


# FCT
SDB_FCT <- read_csv(here("./CSV Files/SDB/SDB Subjunctive FCT.csv")) %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "SDBA",
         Task = "FCT",
         Mood = "Subjunctive",
         School = "SDBA",
         Age = "SDBA")

HSA_FCT <- read_csv(here("./CSV Files/Adult HS/Adult HS Subjunctive FCT.csv")) %>%
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "HSA",
         Task = "FCT",
         Mood = "Subjunctive",
         School = "HSA",
         Age = "HSA")

DLI78_FCT <- read_csv(here("./CSV Files/DLI-78/DLI-78 Subjunctive FCT.csv")) %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-7/8",
         Task = "FCT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "HS7/8")

MLS78_FCT <- read_csv(here("./CSV Files/MLS-78/MLS-78 Subjunctive FCT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-7/8",
         Task = "FCT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "HS7/8")

DLI5_FCT <- read_csv(here("./CSV Files/DLI-5/DLI-5 Subjunctive FCT.csv")) %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "DLI-5",
         Task = "FCT",
         Mood = "Subjunctive",
         School = "DLI",
         Age = "HS5")

MLS5_FCT <- read_csv(here("./CSV Files/MLS-5/MLS-5 Subjunctive FCT.csv")) %>% 
  filter(!School == "GBCS") %>% 
  filter(Short_Prop == "Intensional") %>% 
  mutate(Group = "MLS-5",
         Task = "FCT",
         Mood = "Subjunctive",
         School = "MLS",
         Age = "HS5")


## Join datasets by age group
HS5_EPT <- rbind(DLI5_EPT, MLS5_EPT)
HS78_EPT <- rbind(DLI78_EPT, MLS78_EPT)

HS5_FCT <- rbind(DLI5_FCT, MLS5_FCT)
HS78_FCT <- rbind(DLI78_FCT, MLS78_FCT)


## Join all age groups
EPT <- rbind(SDB_EPT, HSA_EPT, HS78_EPT, HS5_EPT)
FCT <- rbind(SDB_FCT, HSA_FCT, HS78_FCT, HS5_FCT)
```


```{r, prepare-subj-averages}
## SDBA
SDB_EPT_Structure <- aggregate(SDB_EPT$Mood_Use, list(SDB_EPT$Structure), FUN = mean, na.rm = TRUE)
SDB_EPT_Structure <- SDB_EPT_Structure %>% rename(Structure_Avg = x)
SDB_EPT_Structure <- left_join(SDB_EPT, SDB_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

SDB_FCT_Structure <- aggregate(SDB_FCT$Mood_Use, list(SDB_FCT$Structure), FUN = mean, na.rm = TRUE)
SDB_FCT_Structure <- SDB_FCT_Structure %>% rename(Structure_Avg = x)
SDB_FCT_Structure <- left_join(SDB_FCT, SDB_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## HSA
HSA_EPT_Structure <- aggregate(HSA_EPT$Mood_Use, list(HSA_EPT$Structure), FUN = mean, na.rm = TRUE)
HSA_EPT_Structure <- HSA_EPT_Structure %>% rename(Structure_Avg = x)
HSA_EPT_Structure <- left_join(HSA_EPT, HSA_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HSA_FCT_Structure <- aggregate(HSA_FCT$Mood_Use, list(HSA_FCT$Structure), FUN = mean, na.rm = TRUE)
HSA_FCT_Structure <- HSA_FCT_Structure %>% rename(Structure_Avg = x)
HSA_FCT_Structure <- left_join(HSA_FCT, HSA_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## HS7/8
HS78_EPT_Structure <- aggregate(HS78_EPT$Mood_Use, list(HS78_EPT$Structure), FUN = mean, na.rm = TRUE)
HS78_EPT_Structure <- HS78_EPT_Structure %>% rename(Structure_Avg = x)
HS78_EPT_Structure <- left_join(HS78_EPT, HS78_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HS78_FCT_Structure <- aggregate(HS78_FCT$Mood_Use, list(HS78_FCT$Structure), FUN = mean, na.rm = TRUE)
HS78_FCT_Structure <- HS78_FCT_Structure %>% rename(Structure_Avg = x)
HS78_FCT_Structure <- left_join(HS78_FCT, HS78_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## 5th, EPT
HS5_EPT_Structure <- aggregate(HS5_EPT$Mood_Use, list(HS5_EPT$Structure), FUN = mean, na.rm = TRUE)
HS5_EPT_Structure <- HS5_EPT_Structure %>% rename(Structure_Avg = x)
HS5_EPT_Structure <- left_join(HS5_EPT, HS5_EPT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)

HS5_FCT_Structure <- aggregate(HS5_FCT$Mood_Use, list(HS5_FCT$Structure), FUN = mean, na.rm = TRUE)
HS5_FCT_Structure <- HS5_FCT_Structure %>% rename(Structure_Avg = x)
HS5_FCT_Structure <- left_join(HS5_FCT, HS5_FCT_Structure, by = c("Structure" = "Group.1")) %>%  
  mutate(Structure_Percentage = Structure_Avg * 100)


## Join datasets
EPT_Final <- rbind(SDB_EPT_Structure, HSA_EPT_Structure, HS78_EPT_Structure, HS5_EPT_Structure) %>% 
  mutate(Structure = "Subjunctive")

FCT_Final <- rbind(SDB_FCT_Structure, HSA_FCT_Structure, HS78_FCT_Structure, HS5_FCT_Structure) %>% 
  mutate(Structure = "Subjunctive")
```

```{r, merge-subj-averages-for-boxplot}
## Merge group-level averages from above
EPT_Group_Avg <- aggregate(EPT_Final$Mood_Use, list(EPT_Final$Part_ID), FUN = sum, na.rm = TRUE)
EPT_Group_Avg <- EPT_Group_Avg %>% rename(Part_Avg = x)
EPT_Group_Avg <- left_join(EPT_Final, EPT_Group_Avg, by = c("Part_ID" = "Group.1"))

FCT_Group_Avg <- aggregate(FCT_Final$Mood_Use, list(FCT_Final$Part_ID), FUN = sum, na.rm = TRUE)
FCT_Group_Avg <- FCT_Group_Avg %>% rename(Part_Avg = x)
FCT_Group_Avg <- left_join(FCT_Final, FCT_Group_Avg, by = c("Part_ID" = "Group.1"))


## Form overall master file for boxplot
Master <- rbind(EPT_Group_Avg, FCT_Group_Avg)
Master$Age <- factor(Master$Age, levels = c("HS5", "HS7/8", "HSA", "SDBA"))
```

```{r, reorder-groups-for-correlation}
Master_Correlation <- Master
Master_Correlation$Age <- factor(Master_Correlation$Age, levels = c("SDBA", "HS5", "HS7/8", "HSA"))
```

```{r, generate-participant-averages-for-boxplot}
EPT_Participant_Averages <- rbind(SDB_EPT, HSA_EPT, HS78_EPT, HS5_EPT) %>% 
  group_by(Part_ID, Age) %>%
  summarize(Total_Subj = sum(Mood_Use, na.rm = TRUE),
            Total_Responses = sum(!is.na(Mood_Use)),
            Ratio = Total_Subj/Total_Responses) %>% 
  mutate(Task = "EPT",
         Structure_Percentage = (Ratio*100))

FCT_Participant_Averages <- rbind(SDB_FCT, HSA_FCT, HS78_FCT, HS5_FCT) %>% 
  group_by(Part_ID, Age) %>%
  summarize(Total_Subj = sum(Mood_Use, na.rm = TRUE),
            Total_Responses = sum(!is.na(Mood_Use)),
            Ratio = Total_Subj/Total_Responses) %>% 
  mutate(Task = "FCT",
         Structure_Percentage = (Ratio*100))

Participant_Averages <- rbind(EPT_Participant_Averages, FCT_Participant_Averages)
Participant_Averages$Age <- factor(Participant_Averages$Age, levels = c("HS5", "HS7/8", "HSA", "SDBA"))

```

```{r, create-HS-specific-dataframes}
EPT_HS <- Master %>% 
  filter(Task == "EPT")

FCT_HS <- Master %>% 
  filter(Task == "FCT")

HS <- rbind(EPT_HS, FCT_HS) %>% 
  filter(!Group == "SDBA")
HS$Age <- factor(HS$Age, levels = c("HS5", "HS7/8", "HSA"))
```


# Road Map

1. Brief introduction


--
2. Quick summary of BLC article


--
3. Q&A about BLC article


--
4. Open data practices


--
5. Any other questions about…anything


---
# Introduction


- Background:


--
  + Graduated from Rutgers (Bilingualism/SLA) in 2023
  
--
  + Dissertation on heritage language (HL) acquisition in dual language schools
  
--
  + Current postdoctoral fellow in Dual Lab, University of Houston

--
  + Incoming asst. professor at UT Austin


--
- My research areas

--
  + Morphosyntactic development in bilingual Spanish

--
  + Language acquisition in dual language schools

--
  + Typical and atypical morphological development


--
- Research aims:


--
  + Bring together theoretical and applied research in language acquisition

--
  + Create a holistic perspective of heritage language development, particularly in children

--
  + Challenge deficit notions of bilingualism through novel perspectives on the language faculty

---
# Summing Up

Link to article: https://www.cambridge.org/core/journals/bilingualism-language-and-cognition/article/acquiring-morphology-through-adolescence-in-spanish-as-a-heritage-language-the-case-of-subjunctive-mood/674D0A3F8DAAAE5388E728FD98C07D15


--
* Theories within generative heritage language (HL) acquisition differ in multiple dimensions (see Table 3)

--
  + Language acquisition vs. language loss
  
--
  + Focus on group-level or individual-level differences
  
--
  + Process vs. product of acquisition (end result or continuing shifts?)

--
* Understanding bilingual development in childhood might help us to achieve a "unified" acquisition theory


--
* Most existing theories are needed in order to account for results at three levels:

--
  + Group-level differences (why do HSs differ from other types of bilinguals?)

--
  + Individual-level differences (why do some HSs differ from others?)

--
  + Within-speaker differences (what factors characterize the "gray area" in why HSs show very gradient knowledge?)


---
# Unpacking Putnam and Sánchez (2013)


- You've already read about incomplete acquisition (see work by S. Montrul), which proposes:

--
  + Language development is interrupted due to a reduction in input

--
  + HSs differ as a group from other groups of speakers as a result


--
- Putnam and Sánchez discuss *feature reassembly:*

--
  + In generative linguistics, language is made up of bunches of features

--
  + These features get *reassembled* if exposure to the HL is low (prediction of language loss)

--
  + Producing the HL is harder than interpreting it (explains scenario of balanced bilinguals)

--
  + Language acquisition as a process; explains differences between/within speakers
  

--
- These theories do NOT make conflicting predictions, BUT…

--
  + …Putnam & Sánchez (2013) is proposed as an alternative to incomplete acquisition

--
  + …They do differ on philosophical grounds (should we study group-level differences at all?)


---
# Shifting Our Methods


--
- The way we craft our studies can be limiting in the contributions we make!


--
- To test individual differences, we need detailed and precise questionnaire data to gather information about:

--
  + Patterns of exposure and use

--
  + Age of acquisition


--
- In order to evaluate within-speaker differences, we need to:

--
  + Study both production and receptive knowledge (separate tasks)

--
  + Conduct longitudinal studies, whenever possible
  
--
  + Test multiple linguistic structures or areas


--
- We almost always design studies whose questions focus on only 1-2 of three levels mentioned previously:


--
  + Group-level

--
  + Individual-level

--
  + Within-speaker


---
# Results


- Spanish-dominant bilingual adults produce and select subjunctive more than all *groups* of heritage speakers


--
- Individual differences paint a different picture:

--
  + 25/109 HSs produced subjunctive within the range of Spanish-dominant adults

--
  + All participants used (produced + selected) subjunctive in at least 2 instances

--
  + HSs are either gradient or categorical in using the subjunctive (but never in omitting it), so group-level differences are quite misleading


--
- Proficiency and use of Spanish modulated individual subjunctive production and selection rates


--
- Older children and adults use similar rates of subjunctive, more than the fifth grade children


--
- Greater selection of subjunctive than production (in line with Putnam and Sánchez, 2013)


--
- Sets the stage for creating a comprehensive theory of HL acquisition by children:


--
  + Nuances at the between-groups, individual, and within-speaker levels

--
  + (Indirectly) evaluates continued development across the childhood years

--
  + Along with partner studies, requires bridging theoretical linguistics with the education

--
  + Achieves a new perspective that HSs' grammars are *more* complex than those of monolinguals


---
# Pausing, Reflecting, and Transitioning


--
- Any questions, theoretical or otherwise?


--
- What are the implications of this work for our view of heritage bilingualism?


--
- How can we make these data transparent and reproducible to contribute to open science?


--
- How do we make critical decisions in R that influence our analyses?


---
# Reproducible Research in R Using GitHub


--
- Anonymized data without identifiers (name, DOB, any sort of ID number, etc.) are not human subjects data and can be shared publicly


--
- The data and code from this study (and all studies from my dissertation) are available on GitHub, and links to the repository are in my manuscripts (https://github.com/pthane/DLI-Morphosyntax-2023)


--
- GitHub is a place to store codes and other documents related to your research such as:


--
  + *ANONYMIZED* CSV files

--
  + Scripts used for coding
  
--
  + Text files explaining the structure of your repository (usually RTF files)

--
  + Presentations and HTML documents prepared in RMarkdown


--
- This allows others to download and elaborate on your work for free


--
  + GitHub membership is free and unlimited with an @.edu email address
  
--
  + Public repositories are visible to the entire Internet
  
--
  + R is free, and RTF files do not require Microsoft Word (accessible to all)


--
- This satisfies open materials requirements for publications along with a list of stimuli or interview questions


---
# Reproducible Coding in R


--
- Using the `here` function helps others who download your work to reproduce it

--
  + Use this when loading or creating CSV or Excel files

--
  + This creates a _relative path_ so that anyone who downloads your code can use it

--

```{r, here-example, echo = TRUE}
## Install `here` package
install.packages("here", repos = "https://cloud.r-project.org/")


## Load `here` package on every script
library(here)


## Load a CSV file with `here`
## (Notice the use of parentheses. The name of the file begins in the same place where the .Rproj
## file is stored. We will use the adult HSs' production data as an example.)
My_Data <- read_csv(here("CSV Files", "Adult HS", "Adult HS Subjunctive EPT.csv"))
```


---
# Presentations in R with `xaringan`


- The `xaringan` package allows you to turn an RMarkdown file into slides that contain code

--
  + This and all presentations in my GitHub repository were created using `xaringan`

--
  + `xaringan` allows you to use any code as you would in regular R scripts to create presentations


--
- Using `xaringan` has a number of advantages:

--
  + Great for conference presentations that you want others to be able to see after via GitHub

--
  + Great for projects with ongoing data collection

--
  + Perfect for showing a reproducible product to general audiences

--
  + You can post code in your slides, allowing others to learn


--
- Tutorial for using `xaringan`: https://www.youtube.com/watch?v=RPFh3y9UAX4

--
  + Includes 2 hours of instruction and practice activities

--
  + You don't need all 2 hours to get started and get familiar


---
# Sharing `xaringan` RSlides Using GitHub


- All data can be downloaded and presentations can be shared using stable URLs


--
- You must create a subfolder in your RProject with an RMarkdown file named index.Rmd in it, where you will put all of your code to create your slides


--
- After you knit your RMarkdown file with all of the desired code for the presentation, you can upload it to GitHub.


--
- Every GitHub repository has settings to make it into a GitHub site:

--
  + Settings > Pages (on left menu)

--
  + In _Pages_, select 'main' under 'Branches' (this is only possible after you have uploaded files)
  
--
  + Remember to save your settings!


--
- When uploaded to a GitHub repository, the resulting index.html file will automatically turn into a public version of your slides after a few minutes


--
- Review of how to create HTML page: https://mbounthavong.com/blog/2022/7/30/hosting-a-r-markdown-html-file-on-a-github-page


---
# Group-Level Plots from BLC Paper

```{r, generate-subj-bar-graph}
Bar_Graph <- Master %>% 
  ggplot(aes(x = Age, y = Structure_Percentage, fill = Task)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     limits = c(0, 100)) +
    geom_text(aes(label = round(Structure_Percentage)),
            position = position_dodge(width = .9),
            vjust = -0.5,
            size = 6) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of subjunctive responses", title = "Group-Level Average Subjunctive Use") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18),
        strip.text = element_text(size = 20, face = "bold"))
```

```{r, generate-subj-boxplot}
Boxplot <- Participant_Averages %>% 
  ggplot(aes(x = Age, y = Structure_Percentage, fill = Task)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq (0, 100, 20),
                     limits = c(0, 100)) +
  scale_fill_manual(values = c("#56B4E9", "#F0E442")) +
  labs(x = "Group", y = "Percentage of subjunctive responses", title = "Group-Level Summary of Subjunctive Use") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, print-plots}
(Bar_Graph)+(Boxplot)
```


---
# Individual Differences Plot from BLC Paper

```{r, calculate-subj-ind-diffs}
Diffs = left_join(EPT_Participant_Averages, FCT_Participant_Averages, by = "Part_ID", "Age") %>%
  rename(Age = Age.x) %>% 
  select(Part_ID, Age, Structure_Percentage.x, Structure_Percentage.y) %>% 
  rename(Production = Structure_Percentage.x,
         Preference = Structure_Percentage.y) %>% 
  filter(!Age == "SDBA")

Diffs$Age <- factor(Diffs$Age, levels = c("HS5", "HS7/8", "HSA"))
```

```{r, generate-subj-diffs-plot}
Diffs_Plot <- Diffs %>% 
  ggplot(aes(x = Production, y = Preference)) +
  geom_jitter(mapping = aes(color = Age), size = 3) +
  scale_x_continuous(breaks = seq (0, 100, 20),
                     limits = c(0, 100)) +
  scale_y_continuous(breaks = seq (0, 100, 20),
                     limits = c(0, 100)) +
  labs(x = "Percentage of sentences subjunctive produced", y = "Percentage with subjunctive selected", title = "Individual Rates of Subjunctive Selection and Production", color = "Group") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18))
```

```{r, print-subj-diffs-plot}
(Diffs_Plot)
```


---
# GLMM Forest Plot (Result of Correlations)

```{r, generate-subj-overall-correlation}
Correlation <- glmer(Mood_Use ~ Age + Task + Use_Joint_Std + Task:Use_Joint_Std +
                       (1 | Part_ID) + (1 | Item),
                     data = Master_Correlation,
                     family = "binomial")
```

```{r, generate-subj-overall-forest-plot}
Correlation_Plot <- plot_model(Correlation, show.values = TRUE, show.intercept = TRUE, value.offset = .3, transform = NULL, vline.color = "black") +
                scale_y_continuous(breaks = seq(-8, 8, 2),
                     limits = c(-8, 8)) +
              scale_x_discrete(labels = c("Frequency of use : FCT", "Frequency of use", "FCT", "HSA", "HS7/8", "HS5", "(Intercept)")) +
  labs(title = "Subjunctive GLMM", y = "Estimates") +
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 20, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 24, face = "bold"))
```

```{r, print-Subj-overall-forest-plot}
(Correlation_Plot)
```


---
# Tying It All Together


- The BLC paper was prepared using *fully* open science:


--
  + Published open access in BLC (free for you while at OSU: https://www.cambridge.org/core/services/open-access-policies/read-and-publish-agreements/oa-agreement-ohiolink)
  
--
  + All anonymized data are on GitHub and visible to anyone

--
  + `here` and `xaringan` make it easy to share these results with others


--
- Theoretical contributions to language acquisition theory *should* be reproducible because:

--
  + Language is a human right (and bilingualism therefore is too)

--
  + Our ability to make claims about the way that language works depends on the analyses we perform
  
--
  + Sometimes, others may have different ideas on how to make analyses more robust, and sharing our approach is a necessary first step


---
# Closing Questions


- Please always feel free to reach out! pdthane@central.uh.edu

- Time for additional questions about anything you might find useful to know:


--
  + More questions about the broader scope of the article

--
  + How coding decisions impact our analyses and perceptions

--
  + The job market in bilingualism